\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, May 7, 2022}

 \textit{Vespers today is based on the following Sunday, which is the third after Easter.
 All is said as in the Eastertide booklet until the Little Chapter.}

\section*{Little Chapter (1 Pet. 2:11)}

\begin{latinenglishsection}

 \latinenglish{
	 Caríssimi: Obsecro vos tamquam ádvenas et peregrínos,~\GreDagger\
	 abstinére vos a carnálibus desidériis,~*
	 quæ mílitant advérsus ánimam.
 	\Rbar.~Deo gratias.
 }{
 	Beloved: I beseech you as strangers and pilgrims,
 	to refrain yourselves from carnal desires which war against the soul.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\begin{rubricbox}

The hymn \textit{Ad cenam agni providi} and versicle \textit{\Rbar.~Mane nobiscum\dots} are said as given in the booklet for Eastertide.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../hymns/ad-coenam-agni}

{\itshape
	1. The Lamb's high banquet we await
	in snow-white robes of royal state:
	and now, the Red Sea's channel past,
	to Christ our Prince we sing at last. 
	
	2. Upon the Altar of the Cross
	His Body hath redeemed our loss:
	and tasting of his roseate Blood,
	our life is hid with Him in God. 
	
	3. That Paschal Eve God's arm was bared,
	the devastating Angel spared:
	by strength of hand our hosts went free
	from Pharaoh's ruthless tyranny. 
	
	4. Now Christ, our Paschal Lamb, is slain,
	the Lamb of God that knows no stain,
	the true Oblation offered here,
	our own unleavened Bread sincere. 
	
	5. O Thou, from whom hell's monarch flies,
	O great, O very Sacrifice,
	Thy captive people are set free,
	and endless life restored in Thee.
	
	6. For Christ, arising from the dead,
	from conquered hell victorious sped,
	and thrust the tyrant down to chains,
	and Paradise for man regains. 
	
	7. We pray Thee, King with glory decked,
	in this our Paschal joy, protect
	from all that death would fain effect
	Thy ransomed flock, Thine own elect.
	
	8. To Thee who, dead, again dost live,
	all glory Lord, Thy people give;
	all glory, as is ever meet,
	to Father and to Paraclete. Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ma(h)ne(h) no(h)bís(h)cum(h) Dó(h)mi(h)ne,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Quó(h)ni(h)am(h) ad(h)ves(h)pe(h)rá(h)scit,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Stay with us, Lord, alleluia.
\Rbar.~For it is getting towards evening, alleluia.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-6}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} A little while, and you shall see Me no longer, saith the Lord; 
and again a little while, and you shall see Me, because I go to the Father, alleluia, alleluia.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui errántibus, ut in viam possint redíre justítiæ, veritátis tuæ lumen osténdis:
	da cunctis, qui christiána professióne censéntur, et illa respúere, quæ huic inimíca sunt nómini;
	et ea, quæ sunt apta, sectári.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty God, Who showest to them that be in error the light of thy truth,
	to the intent that they may return into the way of righteousness,
	grant unto all them that are admitted into the fellowship of Christ's Religion,
	that they may eschew those things that are contrary to their profession,
	and follow all such things as are agreeable to the same. 
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Stanislaus, Bishop and Martyr (May 7)}

\textit{\textnormal{Ant.} Saints and just rejoice in the Lord, alleluia, the Lord had chosen you for his inheritance, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--sancti_et_justi--solesmes_1961}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Pretiósa in conspéctu Dómini, allelúja.\\
	\Rbar.~ Mors Sanctórum ejus, allelúja.
	
	Orémus.
Deus, pro cujus honóre gloriósus Póntifex Stanisláus gládiis impiórum occúbuit: præsta, qu\'{\ae}sumus; ut omnes, qui ejus implórant auxílium, petitiónis suæ salutárem consequántur efféctum.
}{
	\Vbar.~Precious in the sight of the Lord, alleluia.
	\Rbar.~Is the death of his saints, alleluia.
	
	Let us pray.
O God, for Whose glory the noble Bishop Stanislaw fell by the swords of sinful men, grant, we beseech thee, that all whosoever ask his help, may find such answer to their petition as may profit them to the everlasting salvation of their souls.
}
\end{latinenglishsection}

\section*{Commemoration of the Apparition of St. Michael the Archangel (May 8)}

\textit{\textnormal{Ant.} While as John was gazing upon the holy mystery, the Archangel Michael sounded his trumpet: Forgive us, O Lord our God, Which openest the book, and loosest the seals thereof. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--dum_sacrum_mysterium--solesmes_1961}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Stetit Angelus juxta aram templi, allelúja.\\
	\Rbar.~Habens thuríbulum áureum in manu sua, allelúja.
	
	Orémus.
Deus, qui, miro órdine, Angelórum ministéria hominúmque dispénsas: concéde propítius; ut, a quibus tibi ministrántibus in cælo semper assístitur, ab his in terra vita nostra muniátur.
}{
	\Vbar.~An Angel stood at the Altar of the Temple. Alleluia.
	\Rbar.~Having a golden censer in his hand. Alleluia.
	
	Let us pray.
O God, Who hast ordained and constituted the services of angels and men in a wonderful order, mercifully grant that as thy holy angels alway do thee service in heaven, so, by thy appointment, they may succour and defend us on earth.
	Through Our Lord.
}
\end{latinenglishsection}

\pagebreak

\section*{Commemoration of the Octave of St. Joseph}

\textit{\textnormal{Ant.} When as Mary the Mother of Jesus, was espoused to Joseph, before they came together, she was found with child of the Holy Ghost. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--cum_esset_desponsata_..._allel.--solesmes}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Constítuit eum dóminum domus suæ.\\
	\Rbar.~Et príncipem omnis possessiónis suæ.
	
	Orémus.
Deus, qui ineffábili providéntia beátum Joseph sanctíssimæ Genitrícis tuæ sponsum elígere dignátus es: præsta, qu\'{\ae}sumus; ut quem protectórem venerámur in terris, intercessórem habére mereámur in cælis:
	Qui vivis et regnas cum Deo Patre, in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	\Vbar.~He made him lord of His house. Alleluia.
	\Rbar.~And ruler of all His substance. Alleluia.
	
	Let us pray.
God, Who in thine unspeakable foreknowledge didst choose thy blessed servant Joseph to be the husband of thine Own most holy Mother; mercifully grant that now that he is in heaven with thee, we who on earth do reverence him for our Defender, may worthily be holpen by the succour of his prayers to thee on our behalf;
	Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain with us always. \Rbar.~Amen.
}

\end{latinenglishsection}

\end{document}