\documentclass{../vespers-sheet}
\usepackage{multicol}

\begin{document}

\chapter*{St. John at the Latin Gate}

\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.}  Thy Saints, O Lord, shall flourish like the lily, alleluia;
and they shall be before Thee as the odor of balm, alleluia.}

 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-8}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Sancti tui\dots}.

\end{rubricbox}


\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} The dwelling of Saints is in heavenly kingdoms, alleluia, and their rest is everlasting, alleluia.}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-7}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{In cælestibus\dots}.

\end{rubricbox}

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} Thy Saints, O Lord, cried out within the veil, alleluia, alleluia, alleluia.} 

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-2}
	
}{
	\input{../psalms/english/ps111}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{In velamento\dots}.

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} O ye spirits and souls of the just, sing a hymn to our God, alleluia, alleluia.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-8}
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Spiritus\dots}.

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} The just will shine forth like the sun in the presence of God, alleluia.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps116-2}
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Fulgebunt justi\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Wisdom 5:1)}

\begin{latinenglishsection}

\latinenglish{
	Stabunt justi in magna constántia~\GreDagger\
	advérsus eos qui se angustiavérunt,~*
	et qui abstulérunt labóres eórum.
	\Rbar.~Deo grátias.
}{
	The just shall stand with great constancy against those that have afflicted them, and taken away their labours.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/tristes-erant-apostoli}

{\itshape

1. The Apostles' hearts were full of pain
for their dear Lord so lately slain:
that Lord His servants' wicked train
with bitter scorn had dared arraign.

2. With gentle voice the Angel gave
the women tidings at the grave;
``Forthwith your Master shall you see:
he goes before to Galilee.''

3. And while with fear and joy they pressed
to tell these tidings to the rest,
their Lord, their living Lord they meet,
and see His form, and kiss his feet.

4. The Eleven, when they hear, with speed
to Galilee forthwith proceed:
that there they may behold once more
the Lord's dear face, as oft before.

5. We pray Thee, King with glory decked,
in this our Paschal joy, protect
from that death would fain effect
Thy ransomed flock, Thine own elect.

6. To Thee Who dead, again dost live,
all glory, Lord, Thy people give;
All glory, as is ever meet,
to Father and to Paraclete. Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. San(h)cti(h) et(h) jus(h)ti(h) in(h) Dó(h)mi(h)no(h) gau(h)dé(h)te,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Vos(h) e(h)lé(h)git(h) De(h)us(h) in(h) he(h)re(h)di(h)tá(h)tem(h) si(h)bi,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

%% TODO
\textit{\Vbar.~O ye holy and just ones, rejoice in the Lord, alleluia.
\Rbar.~God hath chosen you unto Himself for an inheritance, alleluia.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.}  When he was sent into a cauldron of boiling oil, the apostle John, by the protection of divine grace, came out unharmed.} 

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-4}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon: \textit{In ferventis\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui cónspicis quia nos úndique mala nostra pertúrbant:
	præsta, qu\'{\ae}sumus; ut beáti Joánnis Apóstoli tui et Evangelístæ intercéssio gloriósa nos prótegat.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, who seest that on every side we are afflicted by our evils: deny us not, we entreat Thee, the safeguard of the glorious intercession of blessed John Thine apostle and evangelist. Through our Lord.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\section*{Commemoration of St. Pope Pius V}

\textit{\textnormal{Ant.} The Lord loved him, and adorned him; He clothed him with a robe of glory, and crowned him at the gates of Paradise, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-pont-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas, alleluia.\\
	\Rbar.~Et osténdit illi regnum Dei, alleluia.
	
	Orémus.
	Deus, qui ad conteréndos Ecclésiæ tuæ hostes, et ad divínum cultum reparándum,
	beátum Pium Pontíficem Máximum elígere dignátus es:
	fac nos ipsíus deféndi præsídiis, et ita tuis inhærére obséquiis;
	ut ómnium hóstium superátis insídiis, perpétua pace lætémur.
	Per Dóminum.
}{
	The Lord led the just in right paths, alleluia.
	\Rbar.~And showed him the kingdom of God, alleluia.
	
	Let us pray.
	O God, who for the overthrowing of the enemies of the Thy church and for the restoring of the beauty of divine worship,
	didst choose blessed Pius as supreme Pontiff: grant that we may be defended by his patronage and so cleave unto Thy service,
	that overcoming all the snares of our enemies, we may rejoice in Thy eternal peace.
	Through Our Lord.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-double}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}