\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, July 23}

 \textit{Vespers today is based on the following Sunday, which is the Seventh Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 11:33)}

\begin{latinenglishsection}

 \latinenglish{
 	O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\
 	quam incomprehensibília sunt judícia ejus, et investigábiles viæ ejus!
	 
 	\Rbar.~Deo grátias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h), Dó(h)mi(f)ne(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a(f.). (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Nathan the Prophet anointed Solomon King in Gihon, and they went up with joy, and cried, saying: May the King live for ever.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, cujus providéntia in sui dispositióne non fállitur: te súpplices exorámus; ut nóxia cuncta submóveas, et ómnia nobis profutúra concédas.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Whose never-failing Providence ordereth all things both in heaven and earth, we humbly beseech thee to put away from us all hurtful things, and to give us those things which be profitable for us.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Apollinaris, Bishop and Martyr (July 23)}

\textit{\textnormal{Ant.} If any man will come after me, let him deny himself and take up his cross daily and follow me.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--qui_vult_venire--solesmes_1961}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Justus ut palma florébit.\\
	\Rbar.~Sicut cedrus Líbani multiplicábitur.
	
	Orémus.
	Deus, fidélium remunerátor animárum, qui hunc diem beáti Apollináris Sacerdótis tui martýrio consecrásti: tríbue nobis, qu\'{\ae}sumus, fámulis tuis; ut, cujus venerándam celebrámus festivitátem, précibus ejus indulgéntiam consequámur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	\Vbar.~The righteous shall flourish like the palm-tree.
	\Rbar.~He shall grow like a cedar in Lebanon.
	
	Let us pray.
	O God, Who art thyself the exceeding great Reward of all faithful souls, and in Whose sight this day is holy because thy blessed Priest Apollinaris did hereon lift up his last earthly testimony, we beseech thee to grant unto us thy servants, who do keep his worshipful Feast-day, to obtain by his prayers thy gracious remission for our offences.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}