\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of Pentecost}

\section*{Beginning of the Office}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently.
Then, when the Officiant stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} When the days of Pentecost were drawing to a close, they were all saying together, alleluia.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-3}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Dum compleréntur\dots}}

\end{rubricbox}


\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} The Spirit of the Lord has filled the whole world, alleluia.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}
 
 \begin{rubricbox}

{\color{red}All \textbf{stand} as the Cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following Psalms.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-8}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Spiritus Domini\dots}}

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} They were all filled with the Holy Spirit, and began to speak, alleluia.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-8}
	
}{
	\input{../psalms/english/ps111}
}

\end{latinenglishsection}
\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Repleti sunt omnes\dots}}

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} Ye fountains, and all things that move in the waters, sing a hymn to God, alleluia.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-1}
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Fontes\dots}}

\end{rubricbox}

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} The Apostles were speaking with diverse tongues the wonderful works of God, alleluia, alleluia, alleluia.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps116-7}
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Loquebantur\dots}}

\end{rubricbox}

\section*{Little Chapter (Acts 2:1-2)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	 Cum compleréntur dies Pentecóstes, erant omnes discípuli pariter in eódem loco:~\GreDagger\
	 et factus est repénte de c{\oe}lo sonus, tamquam adveniéntis spíritus veheméntis:~*
	 et replévit totam domum, ubi erant sedentes.
	\Rbar.~Deo grátias.
}{
 	When the days of the Pentecost were accomplished, they were all together in one place:
 	and suddenly there came a sound from heaven, as of a mighty wind coming,
 	and it filled the whole house where they were sitting. 
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/veni-creator-spiritus}

{\itshape

	1. Come, Holy Spirit, Creator blest,
	and in our souls take up Thy rest;
	come with Thy grace and heavenly aid
	to fill the hearts which Thou hast made.
	
	2. O comforter, to Thee we cry,
	O heavenly gift of God Most High,
	O fount of life and fire of love,
	and sweet anointing from above.
	
	3. Thou in Thy sevenfold gifts are known;
	Thou, finger of God's hand we own;
	Thou, promise of the Father, Thou
	Who dost the tongue with power imbue. 
	
	4. Kindle our sense from above,
	and make our hearts o'erflow with love;
	with patience firm and virtue high
	the weakness of our flesh supply.
	
	5. Far from us drive the foe we dread,
	and grant us Thy peace instead;
	so shall we not, with Thee for guide,
	turn from the path of life aside.
	
	6. Oh, may Thy grace on us bestow
	the Father and the Son to know;
	and Thee, through endless times confessed,
	of both the eternal Spirit blest.
	
	7. Now to the Father and the Son,
	Who rose from death, be glory given,
	with Thou, O Holy Comforter,
	henceforth by all in earth and heaven.
	Amen. 
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Re(h)plé(h)ti(h) sunt(h) om(h)nes(h) Spí(h)ri(h)tu(h) San(hi)cto,(h_') (,)
al(h)le(fe)lú(f_h)ia.(hiH'Ghih.ghG'FE'fggf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) c<sp>oe</sp>(h)pé(h)runt(h) lo(hi)qui,(h_') (,) al(h)le(fe)lú(f_h)ia.(hiH'Ghih.ghG'FE'fggf.)(::)
}

\textit{\Vbar.~They were all filled with the Holy Spirit, alleluia.
\Rbar.~And began to speak, alleluia.}

\vfill\pagebreak

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} I will not leave you as orphans, alleluia; I go away and am coming to you, alleluia; and your heart shall rejoice, alleluia.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

{\color{red}After the Officiant intones up to the asterisk, all \textbf{sit} and join in singing.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-1}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Non vos relinquam\dots},
then \textbf{stand} for the prayer.}

\end{rubricbox}

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui hodiérna die corda fidélium Sancti Spíritus illustratióne docuísti:
	da nobis in eódem Spíritu recta sápere;
	et de ejus semper consolatióne gaudére.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen. 
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who on this day hast taught the hearts of the faithful by the light of the Holy Spirit,
	grant us by that same Spirit, to know what is right, and ever to rejoice in His consolation.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen. 
}

\textit{\color{red}The Officiant leads the following:}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solem}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}
\textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}