\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, August 13, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Tenth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 11:33)}

\begin{latinenglishsection}

 \latinenglish{
 	O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\ 
 	quam incomprehensibília sunt ju\textit{dícia }\textbf{e}jus, et investigábiles viæ ejus!
 	
 	\Rbar.~Deo grátias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!	
	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h) Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} All wisdom is of the Lord God, and was with Him from everlasting, yea before time was.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui omnipoténtiam tuam parcéndo máxime et miserándo maniféstas:\GreDagger\ multíplica super nos miseri\textit{córdiam} \textbf{tu}am; ut, ad tua promíssa curréntes, cæléstium bonórum fácias esse consórtes.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who declarest thine Almighty power most chiefly in showing mercy and pity, mercifully grant unto us such a measure of thy grace, that we, running the way of thy commandments, may obtain thy gracious promises, and be made partakers of thy heavenly treasure.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of Sts. Hippolytus and Cassian, Martyrs (August 13)}

\textit{\textnormal{Ant.} In heaven the souls of the Saints rejoice, who have followed in Christ's footsteps;
and since they shed their blood for him, they rejoice with Him forever.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Exsultábunt Sancti in glória.\\
	\Rbar.~Lætabúntur in cubílibus suis.
	
	Orémus.
	Da quæsumus, omnípotens Deus: ut beatórum Mártyrum tuórum Hippólyti et Cassiáni veneránda solémnitas, et devotiónem nobis áugeat, et salútem. 
	Per Dominum nostrum Jesum Christum Filium tuum, qui tecum vivit et regnat in unitate Spiritus Sancti, Deus, per omnia saecula saeculorum.
	
	\Rbar.~Amen.
}{
	The just exult in their glory.
	\Rbar.~They shout for joy on their beds.
	
	Let us pray.
	Grant, we beseech Thee, O almighty God, that the august solemnity of the blessed martyrs Hippolytus and Cassian may increase both our devotion and our salvation.
	Through Our Lord Jesus Christ Thy Son, who liveth and reigneth with Thee in the unity of the Holy Ghost, God, forever and ever.
	
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Eusebius, Confessor (August 14)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.} 

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ 
	\Rbar.~Stolam glóriæ induit eum.  
	
	Orémus.
	Deus, qui nos beáti Eusébii Confessóris tui ánnua solemnitáte lætíficas: concéde propítius; ut cujus natalícia cólimus, per ejus ad te exémpla gradiámur.
}{
	The Lord loved him and adorned him.
	\Rbar.~He clothed him with a robe of glory. 
	
	Let us pray.
	O God, Who, year by year, dost gladden us by the solemn feast-day of thy blessed confessor Eusebius, mercifully grant unto all who keep his birthday, grace to follow his steps toward thee.
}

\end{latinenglishsection}

\section*{Suffrage of the Saints}

\textit{\textnormal{Ant.} May the Blessed Mother of God, the Virgin Mary, and all the Saints, intercede for us before the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/suffrage-saints}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Mirificávit Dóminus Sanctos suos.\\
	\Rbar. Et exaudívit eos clamántes ad se.
	
	Orémus.
	A cunctis nos, qu\'{\ae}sumus Dómine, mentis et córporis defénde perículis:
	et intercedénte beáta et gloriósa semper Vírgine Dei Genitríce María,
	cum beáto Joseph, beátis Apóstolis tuis Petro et Paulo, atque beáto Thoma apóstolo et ómnibus Sanctis,
	salútem nobis tríbue benígnus et pacem;
	ut destrúctis adversitátibus et erróribus univérsis,
	Ecclésia tua secúra tibi sérviat libertáte.
	Per Dóminum.
}{
	The Lord hath made His holy ones wonderful.
	\Rbar.~And He hath heard them when they cried unto Him.
	
	Let us pray.
	From all perils of soul and body defend us, O Lord, we beseech Thee,
	and by the intercession of blessed and glorious ever Virgin Mary, Mother of God,
	of blessed Joseph, of thy blessed apostles Peter and Paul, and of blessed apostle Thomas and all the Saints,
	graciously grant us safety and peace that all adversities and errors being overcome,
	thy Church may serve Thee in security and freedom. Through our Lord. \Rbar.~Amen.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-double}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}