\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, 26 Feb 2022}

 \textit{Vespers today is based on the following Sunday, which is Quinqagesima Sunday.
 All is said as given in the main booklet until the Little Chapter.
  As noted in the main booklet, \textnormal{Alleluia} at the beginning of the Office is omitted and replaced with
 \textnormal{Laus tibi Domine, rex æternæ gloriæ}.}

 \section*{Little Chapter (1 Cor. 13:1)}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Si linguis hóminum loquar, et angelórum,~\GreDagger\
	 caritátem autem non hábeam,~*
	 factus sum velut æs sonans aut cýmbalum tínniens.
 	\Rbar.~Deo gratias.
 }{
 	Brethren: If I should speak with the tongues of men and of angels,
 	but do not have charity, I have become as sounding brass or a tinkling cymbal.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/o-lux-beata-trinitas}

{\itshape
	1. O Trinity of blessed Light,
	O Unity of sovereign might,
	as now the fiery sun departs,
	shed Thou Thy beams within our hearts.
	
	2. To Thee our morning song of praise,
	to Thee our evening prayer we raise;
	Thee may our glory evermore
	in lowly reverence adore.
	
	3. All laud to God the Father be;
	all praise, Eternal Son, to Thee;
	all glory, as is ever meet,
	to God the Holy Paraclete.
}

\vfill\pagebreak

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h) Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~May our evening prayer ascend to Thee, O Lord.
\Rbar.~And may Thy mercy descend upon us.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-6}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} The father of our faith, the great Abraham, offered a holocaust on the altar in place of his son.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Preces nostras, qu\'{\ae}sumus, Dómine, cleménter exáudi:
	atque, a peccatórum vínculis absolútos, ab omni nos adversitáte custódi.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, we beseech You, mercifully hear our prayers;
	loose us from the chains of our sins and keep us from all adversity.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Gabriel of Our Lady of Sorrows, Confessor (Feb. 27)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ induit eum.  %PT: allelúia
	
	Orémus.
	Deus, qui beátum Gabriélem dulcíssimæ Matris tuæ dolóres assídue recólere docuísti,
	ac per illam sanctitátis et miraculórum glória sublimásti:
	da nobis, ejus intercessióne et exémplo;
	ita Genitrícis tuæ consociári flétibus ut matérna ejúsdem protectióne salvémur.
	Qui vivis.
}{
	The Lord loved him and adorned him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	O God, Who didst teach blessed Gabriel diligently to ponder the sorrows of Thy most sweet Mother,
	and Who hast exalted him to the glory of sanctity and the working of miracles;
	grant us, through his intercession and example,
	so to mourn with Thy Mother, that we may be saved through her maternal care.
	Who livest.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}