\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of the Nativity of Our Lord Jesus Christ}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently.
Then, when the Officiant stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

%TODO: Add that the correct psalms, and verify that their tones, and their associated pointed text are correct

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} The King of Peace is magnified, whose face the whole earth desireth.
 \textnormal{Ps.} The Lord said to my Lord: * Sit thou at my right hand:}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-8} %%

}{
	\input{../psalms/english/ps109} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Rex pacíficus\dots}} %%

\end{rubricbox}

\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} The Peaceful King exceeded all the kings of the whole earth.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; * in the council of the just, and in the congregation.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps110-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps110-7} %%

}{
	\input{../psalms/english/ps110} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Magnificátus est\dots}} %%

\end{rubricbox}

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} The days of Mary * were accomplished that she should bring forth her first-born Son.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: * he shall delight exceedingly in his commandments.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps111-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-8} %%

}{
	\input{../psalms/english/ps111} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Compléti sunt\dots}} %%

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} Know ye that the kingdom * of God is at hand? Amen, I say unto you, it will not tarry.
 \textnormal{Ps.} Praise the Lord, ye children: * praise ye the name of the Lord.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-8} %%

}{
	\input{../psalms/english/ps112} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Scitóte\dots}} %%

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} Lift up your heads; * behold, your redemption draweth nigh.
 \textnormal{Ps.} Praise the Lord, all ye nations: * praise him, all ye people.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps116-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps116-8} %%

}{
	\input{../psalms/english/ps116} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Leváte cápita vestra: \dots}} %%

\end{rubricbox}

\vfill\pagebreak

\section*{Little Chapter (Titus 3:4-5)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Appáruit benígnitas, et humánitas Salvatóris nostri Dei: \GreDagger non ex opéri\textit{bus ju}\textbf{stí}tiæ, quæ fécimus nos,*~\\ sed secúndum suam misericórdiam salvos nos fecit.
	\Rbar.~Deo grátias.
}{
	But when the goodness and kindness of God our Saviour appeared: Not by the works of justice, which we have done, but according to his mercy, he saved us.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

% TODO: Verify that the hymn is correct for the feast (including the responsory after the hymn)

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jesu-redemptor-omnium}

{\itshape
	1. Jesus, the Ransomer of man,
	Who, ere created light began,
	Didst from the sovereign Father spring,
	His power and glory equalling.
	
	2. The Father’s Light and Splendor Thou,
	Their endless hope to Thee that bow;
	Accept the prayers and praise to-day
	That through the world Thy servants pay.
	
	3. Salvation’s Author, call to mind
	How, taking form of humankind,
	Born of a Virgin undefiled,
	Thou in man’s flesh becam’st a Child.
	
	4. Thus testifies the present day,
	Through every year in long array,
	That Thou, salvation’s source alone,
	Proceedest from the Father’s throne.
	
	5. The heavens above, the rolling main
	And all that earth’s wide realms contain,
	With joyous voice now loudly sing
	The glory of their new-born King.
	
	6. And we who, by Thy precious Blood
	From sin redeemed, are marked for God,
	On this the day that saw Thy birth,
	Sing the new song of ransomed earth.
	
	7. O Lord, the Virgin-born, to Thee
	Eternal praise and glory be,
	Whom with the Father we adore
	And Holy Ghost forevermore.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Crá(h)sti(h)na(h) di(h)e(h) de(h)lé(h)bi(h)tur(h) i(h)ní(h)qui(h)tas(h) ter(h)ræ.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) reg(h)ná(h)bit(h) su(h)per(h) nos(h) Sal(h)vá(h)tor(h) mun(h)di.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Glory and riches are in his house.
\Rbar.~And his justice endureth forever.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Behold a faithful and prudent servant, whom the Lord has set over His household.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.}

\begin{rubricbox}

{\color{red}The Cantor leads by intoning the antiphon and the first verse.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the Cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
3. Quia respéxit humilitátem ancíllæ \textbf{su}æ:~* 
ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.

4. Quia fecit mihi magna qui \textbf{pot}ens est:~* 
et sanctum \textit{no}\textit{men} \textbf{e}jus.

5. Et misericórdia ejus a progénie in pro\textbf{gé}nies~* 
timén\textit{ti}\textit{bus} \textbf{e}um.

6. Fecit poténtiam in bráchio \textbf{su}o:~* 
dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.

7. Depósuit poténtes de \textbf{se}de,~* 
et exal\textit{tá}\textit{vit} \textbf{hú}miles.

8.Esuriéntes implévit \textbf{bo}nis:~* 
et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.

9. Suscépit Israël púerum \textbf{su}um,~* 
recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.

10. Sicut locútus est ad patres \textbf{nos}tros,~* 
Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{\color{red}(bow)}Glória Patri, et \textbf{Fí}lio,~* 
et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{\color{red}(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~* 
et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Cum ortus fúerit \dots}, %%
then \textbf{stand} for the prayer.}

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Concéde, qu\'{\ae}sumus,\\ omnípotens Deus: \GreDagger\ ut nos Unigéniti tui nova per carnem Natí\textit{vitas} \textbf{lí}beret; quos sub peccáti jugo vetústa sérvitus tenet.
	Per eúndem Dóminum\\ nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Grant, we beseech thee, almighty God, that the new birth of thy Only-begotten Son as man may set us free, who are held by the old bondage under the yoke of sin.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red} The Suffrage of the Saints is omitted on this day.}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solem}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\section*{Marian Anthem}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/alma-redemptoris-monasticum}

{\itshape
	Holy mother of our Redeemer,
	thou gate leading to heaven and star of the sea; 
	help the falling people who seek to rise, 
	thou who, all nature wondering, didst give birth to thy holy Creator. 
	Virgin always, hearing the greeting from Gabriel’s lips, take pity on sinners.
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Post partum, Virgo, inviolata permansisti.\\
	\Rbar. Dei Genitrix, intercede pro nobis.
	
	Orémus.
	Deus, qui salutis aeternae, beatae Mariae virginitate fecunda, humano generi praemia praestitisti: \GreDagger tribue, quaesumus, ut ipsam pro nobis intercedere \textit{senti}\textbf{a}mus, per quam meruimus auctorem vitae\\ suscipere, ~* Dominum nostrum Iesum Christum, Filium tuum.\\
	\Rbar.~Amen.
}{
	Vouchsafe that I may praise thee, O sacred Virgin.
	\Rbar.~Give me stength against thine enemies.
	
	Let us pray.
	O God, Who by the fruitful virginity of blessed Mary hast offered unto the human race the rewards of eternal salvation, grant, we beseech thee, that we may know the effects of her intercession, through whom we have deserved to receive the author of life, our Lord Jesus Christ, Thy Son. 
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}