\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, October 15, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Nineteenth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Rom 11:33)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,\GreDagger\
	 quam incomprehensibília sunt ju\textit{dícia} \textbf{e}jus, et investigábiles viæ ejus!
 	\Rbar.~Deo gratias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

%% TODO: correct the hymn using the antiphonary
\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape

1. As fades the glowing orb of day,
To thee, great source of light, we pray;
Blest Three in One, to every heart
Thy beams of life and love impart.

2. At early dawn, at close of day,
To thee our vows we humbly pay;
May we, mid joys that never end,
With thy bright saints in homage bend.

3. To God the Father, and the Son,
And Holy Spirit, Three in One,
Be endless glory, as before
The world began, so evermore.
Amen.

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h)scén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

%% TODO: correct the magnificant antiphon and tone - add pointed text to the intonation if needed

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Moreover, all Israel made great lamentation for Judas, saying How art thou fallen, thou valiant man of war, that didst deliver the Lord's people.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens et miséricors Deus, univérsa nobis adversántia propitiátus exclúde: ut mente et córpore páriter expedíti, quæ tua sunt, líberis méntibus exsequámur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Almighty and most merciful God, of thy bountiful goodness keep us, we beseech thee, from all things that may hurt us; that we, being ready both in body and soul, may cheerfully accomplish those things that Thou wouldest have done.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of Pope St. Callixtus, Martyr (October 14)}

\textit{\textnormal{Ant.} Saints and just rejoice in the Lord, alleluia, the Lord had chosen you for his inheritance, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--sancti_et_justi--solesmes_1961}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Pretiósa in conspéctu Dómini, allelúja.\\
	\Rbar.~ Mors Sanctórum ejus, allelúja.
	
	Orémus.
	Deus, qui nos cónspicis ex nostra infirmitáte defícere: ad amórem tuum nos misericórditer per sanctórum tuórum exémpla restáura. 
}{
	\Vbar.~Precious in the sight of the Lord, alleluia.
	\Rbar.~Is the death of his saints, alleluia.
	
	Let us pray.
	O God, Who dost behold how we fail in our weakness, mercifully restore us unto Thy love through the examples of Thy saints.
}
\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of St. Teresa of Avila, Virgin (October 15)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Diffúsa est grátia in lábiis tuis.\\ 
	\Rbar.~Proptérea benedíxit te Deus in ætérnum. 
	
	Orémus.
	Exáudi nos, Deus, salutáris noster: ut, sicut de beátæ Terésiæ Vírginis tuæ festivitáte gaudémus; ita cæléstis ejus doctrínæ pábulo nutriámur, et piæ devotiónis erudiámur afféctu.
}{
	Grace is poured out upon thy lips. 
	\Rbar.~Therefore God hath blessed thee forever. 
	
	Let us pray.
	Graciously hear us, O God of our salvation, and grant that as we do rejoice for the Feast-Day of thy blessed hand-maiden Teresa, so we may feed to our ghostly health upon her heavenly teaching, and better ourselves by the example of her godly conversation.
}

\end{latinenglishsection}

\section*{Commemoration of St. Hedwig, Widow (October 16)}

\textit{\textnormal{Ant.} The kingdom of heaven is like a merchant seeking good pearls, who, when he had found one of great price,
gave all that he had and bought it.} 
\gresetinitiallines{1}
\gregorioscore{../commemorations/holy-woman-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua.\\
	\Rbar.~Inténde, próspere procéde, et regna. 
	
	Orémus.
	Deus, qui beátam Hedwígem a s\'{\ae}culi pompa ad húmilem tuæ crucis sequélam toto corde transíre docuísti: concéde; ut ejus méritis et exémplo discámus peritúras mundi calcáre delícias, et in ampléxu tuæ crucis ómnia nobis adversántia superáre:
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	In thy glory and thy splendor. 
	\Rbar.~Go forth, advance with victory and reign.
	
	Let us pray.
	O God, Who didst teach thy blessed hand-maid Hedwig to turn away from the glory of the world, and with all her heart to take up her Cross and follow thee, teach us, for her sake and after her example, to hold light the perishing pleasures of this present world, and cleaving ever unto thy Cross to rest in the end more than conquerors over all things that would hurt us.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}