\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, 5 Feb. 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fifth after Epiphany. All is said as in the main booklet until the Magnificat.}

\section*{Magnificat}

%% all the Sundays after Epiphany share the same antiphon
\gresetinitiallines{1}
\gregorioscore{../post-epiph-2/magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{../post-epiph-2/magnificat-antiphon}

\textit{\textnormal{Ant.} God hath received Israel His servant according as He spoke to Abraham and to his seed, and hath exalted the humble forever.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Famíliam tuam, qu\'{\ae}sumus Dómine, contínua pietáte custódi: ut quæ in sola spe grátiæ cæléstis innítitur,
	tua semper protectióne muniátur.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	In thine infinite goodness, we beseech thee, O Lord, to watch over Thy household,
	that even as it relies solely upon the hope of Thy heavenly grace, so it may ever be defended by Thy protection.
	Through our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Agatha, Virgin and Martyr (Feb. 5)}

\textit{\textnormal{Ant.} Blessed Agatha standing in prison with outstretched hands prayed unto the Lord: O Lord Jesus Christ, good Master, I give Thee thanks Who hast enabled me to overcome the tortures of the executioners; bid me, O Lord, come happily to Thine unfading glory.}

\gresetinitiallines{1}
\gregorioscore{comm-agatha}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Diffúsa est grátia in lábiis tuis.\\ %PT: allelúia
	\Rbar.~Proptérea benedíxit te Deus in ætérnum.  %PT: allelúia
	
	Orémus.
	Deus, qui inter cétera poténtiæ tuæ mirácula, étiam in sexu frágili victóriam martýrii contulísti:
	concéde propítius; ut qui beátæ Agathæ Vírginis et Mártyris tuæ natalítia cólimus,
	per ejus ad te exémpla gradiámur.
}{
	Grace is poured out upon thy lips. %PT: alleluia
	\Rbar.~Therefore God hath blessed thee forever. %PT: alleluia
	
	Let us pray.
	O God, who among the wonders of Thy power hast granted even to the weaker sex the triumph of martyrdom:
	mercifully grant that we who celebrate the heavenly birthday of blessed Agatha Thy virgin and martyr may,
	through her example, advance nearer to Thee.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of a St. Titus, Bishop and Confessor (Feb. 6)}

\textit{\textnormal{Ant.} O Priest and Bishop, thou worker of mighty works, good shepherd of the people, pray unto the Lord for us.}
%PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ induit eum.  %PT: allelúia
	
	Orémus.
	Deus, qui beátum Titum Confessórem tuum atque Pontíficem apostólicis virtútibus decorásti:
	ejus méritis et intercessióne concéde; ut juste et pie vivéntes in hoc s\'{\ae}culo,
	ad cæléstem pátriam perveníre mereámur.
}{
	The Lord loved him and adorned him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	O God, Who didst adorn blessed Titus, Thy confessor and bishop, with apostolic virtues: grant that through his merits and prayers,
	we may live justly and piously in this world, and thereby be made worthy to arrive at our heavenly country.
}

\end{latinenglishsection}

\section*{Commemoration of St. Dorothy, Virgin and Martyr (Feb. 6)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua.\\ %PT: allelúia
	\Rbar.~Inténde, próspere procéde, et regna.  %PT: allelúia
	
	Orémus.
	Indulgéntiam nobis, qu\'{\ae}sumus Dómine, beáta Doróthea Virgo et Martyr implóret:
	quæ tibi grata semper éxstitit et mérito castitátis,
	et tuæ professióne virtútis.
	Per Dóminum.
}{
	In thy glory and thy splendor. %PT: alleluia
	\Rbar.~Go forth, advance with victory and reign. %PT: alleluia
	
	Let us pray.
	May blessed Dorothy Thy virgin and martyr, we bseech Thee, O Lord,
	implore for us Thy forgiveness; for she was ever pleasing unto Thee,
	both by the merit of her chastity, and by her confession of Thy power.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}