\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, April 23}

 \textit{Vespers today is based on the following Sunday, which is the first after Easter, also called Low Sunday.
 All is said as in the Eastertide booklet until the Little Chapter, except that the antiphon is sung in full before the psalms, as follows:}
 
\gresetinitiallines{1}
\gregorioscore{pss-intonation}

\section*{Little Chapter (1 John 5:4)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 Caríssimi: Omne, quod natum est ex Deo, vincit mundum:~\GreDagger\
	 et hæc est victoria, quæ vincit mundum,~*
	 fides nostra.
 	\Rbar.~Deo gratias.
 }{
 	Beloved: Whatsoever is born of God, overcometh the world:
 	and this is the victory which overcometh the world, our faith. 
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\begin{rubricbox}

{\color{red}The hymn \textit{Ad cenam agni providi} and versicle \textit{\Rbar.~Mane nobiscum\dots} are said as given in the booklet for Eastertide.}

\end{rubricbox}

\section*{Magnificat}

\textit{\textnormal{Ant. Magn.} When it was late that same day, the first of the week,
though the doors where the disciples gathered had been closed, Jesus stood in the midst and said to them: Peace be to you, alleluia.}

\begin{rubricbox}

{\color{red}After the Cantor intones up to the asterisk, all \textbf{sit} and join in singing.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Cum esset sero\dots},
then \textbf{stand} for the prayer.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Præsta, qu\'{\ae}sumus, omnípotens Deus:
	ut, qui paschália festa perégimus,
	hæc, te largiénte, móribus et vita teneámus.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Grant, we beseech Thee, almighty God, that we who have celebrated the Paschal Feast,
	may, by Thy bounty, retain its fruits in our daily habits and behavior.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tp}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}