\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, March 26, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fourth of Lent (Laetare Sunday).
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Gal. 4:22-24)}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Scriptum est: Quóniam Abraham duos fílios habuit: unum de ancílla, et unum de líbera.~\GreDagger\
	 Sed qui de ancílla, secúndum carnem natus est: qui autem de líbera, per repromissiónem:~*
	 quæ sunt per allegóriam dicta.
 	\Rbar.~Deo gratias.
 }{ 	
 	Brethren: it is written that Abraham had two sons: the one by a bondwoman, and the other by a free woman.
 	But he who was of the bondwoman, was born according to the flesh: but he of the free woman, was by promise.
 	Which things are said by an allegory.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/audi-benigne-conditor}

\vfill
\pagebreak

{\itshape

1. O merciful Creator, hear!
To us in pity bow Thine ear:
accept the tearful prayer we raise
in this our fast of forty days.

2. Our hearts are open, Lord, to Thee:
Thou knowest our infirmity;
pour out on all who seek Thy face
abundance of Thy pardoning grace.

3. Our sins are many, this we know;
spare us, good Lord, Thy mercy show;
and for the honor of Thy name
our fainting souls to life reclaim.

4. Give us self-control that springs
from discipline of outward things,
that fasting inward secretly
the soul may purely dwell with Thee.

5. We pray Thee, Holy Trinity,
one God, unchanging Unity,
that we from this our abstinence
may reap the fruits of penitence. Amen.

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. An(h)ge(h)lis(h) su(h)is(h) De(h)us(h) man(h)dá(h)vit(h) de(h) te.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Ut(h) cus(h)tó(h)di(h)ant(h) te(h) in(h) óm(h)ni(h)bus(h) vi(h)is tu(h)is.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~	He hath given his angels charge over thee.
\Rbar.~To keep thee in all thy ways.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Hath no man condemned thee, woman? No man, Lord. Neither do I condemn thee; henceforth sin no more.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Concéde, qu\'{\ae}sumus, omnípotens Deus:
	ut, qui ex merito nostræ actiónis afflígimur,
	tuæ grátiæ consolatióne respirémus.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Grant, we beseech Thee, almighty God, that we who justly suffer for our sins may find relief in the help of Thy grace.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. John of Damascus, Confessor and Doctor of The Church (March 27)}

\textit{\textnormal{Ant.} O right excellent Teacher, light of the Holy Church, St. John, Lover of God's law, pray for us to the Son of God.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{an--o_doctor_optime_john_damascene--solesmes_1961}
% PT: \gregorioscore{../commemorations/common-bvm-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ índuit eum.  %PT: allelúia
	
	Orémus.
	Omnípotens sempitérne Deus, qui ad cultum sacrárum imáginum asseréndum, beátum Joánnem cælésti doctrína et admirábili spíritus fortitúdine imbuísti: concéde nobis ejus intercessióne et exémplo; ut, quorum cólimus imágines, virtútes imitémur et patrocínia sentiámus. Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	The Lord loved him and beautified him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	O Almighty and everlasting God, Who didst fill Thy blessed servant John with heavenly teaching, and wondrous strength of spirit to maintain the honouring of holy images. Grant unto us at his prayers and after his example to take pattern by their holy lives whose images we honour, and ever to feel the power of their help.
Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}


\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}