\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, April 2, 2022}

 \textit{Vespers today is based on the following Sunday, which is Passion Sunday.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Heb. 9:11-12)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Christus assístens Pontifex futurórum bonórum,
	 per ámplius et perféctius tabernáculum non manufáctum, id est, non hujus creatiónis:~\GreDagger\
	 neque per sánguinem hircórum aut vitulórum, sed per próprium sánguinem introívit semel in Sancta,~*
	 ætérna redemptióne invénta.
 	\Rbar.~Deo gratias.
 }{ 	
 	Brethren: Christ, appearing as high priest of the good things to come,
 	by a greater and more perfect tabernacle, not made with hands, that is, not of this creation:
 	neither by the blood of goats, or of calves, but by his own blood, entered once into the holies,
 	having obtained eternal redemption.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\begin{rubricbox}

{\color{red}All \textbf{kneel} during the verse of this hymn beginning \textit{O Crux Ave}.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../hymns/vexilla-regis-passiontide}

{\itshape

1. Abroad the regal banners fly,
now shines the Cross's mystery:
upon it Life did death endure,
and yet by death did life procure.

2. Who, wounded with a direful spear,
did purposely to wash us clear
from stain of sin, pour out a flood
of precious water mixed with blood. 

3. That which the prophet-king of old
hath in mysterious verse foretold,
is now accomplished, whilst we see
God ruling the nations from a Tree. 

4. O lovely and refulgent Tree,
adorned with purpled majesty;
culled from a worthy stock, to bear
those limbs which sanctified were.

5. Blest Tree, whose happy branches bore
the wealth that did the world restore;
the beam that did that Body weigh
which raised up Hell's expected prey.

6. Hail Cross, of hopes the most sublime!
Now, in the mournful Passion time;
grant to the just increase of grace,
and every sinner's crimes efface. 

7. Blest Trinity, salvation's spring
may every soul Thy praises sing;
to those Thou grantest conquest by
the Holy Cross, rewards supply. Amen. 

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. E(h)ri(h)pe(h) me(h) Dó(h)mi(h)ne(h) ab(h) hó(h)mi(h)ne(h) ma(h)lo.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h) vi(h)ro(h) i(h)ní(h)quo(h) é(h)ri(h)pe(h) me.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Deliver me, O Lord, from the evil man.
\Rbar.~Rescue me from the unjust man.}

\vfill\pagebreak

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} I am one that give testimony of Myself, and the Father that sent Me giveth testimony of Me.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Qu\'{\ae}sumus, omnípotens Deus, familiam tuam propítius réspice:
	ut, te largiénte, regátur in córpore; et, te servánte, custodiátur in mente.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Look graciously upon Thy household, almighty God, we beseech Thee,
	that by Thy grace we may be governed in body, and by Thy protection safeguarded in mind.
	Through Jesus Christ, Thy Son our Lord, Who liveth and reigneth with Thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of a Saint Francis of Paola, Confessor (April 2)}

\textit{\textnormal{Ant.} This man, despising the world, and triumphing over earthly things, hath laid up treasures in heaven by word and deed.} 

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ 
	\Rbar.~Et osténdit illi regnum Dei.  
	
	Orémus.
	Deus, humílium celsitúdo, qui beátum Francíscum Confessórem Sanctórum tuórum glória sublimásti: tríbue, qu\'{\ae}sumus; ut, ejus méritis et imitatióne, promíssa humílibus pr\'{\ae}mia felíciter\\ consequámur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	The Lord led the just in right paths.
	\Rbar.~And showed him the kingdom of God. 
	
	Let us pray.
	O God, Who exaltest the meek, and hast raised up thy Blessed Confessor Francis even unto the glory of thy Saints, grant unto us, we beseech thee, for his sake, so to walk after him in lowliness of heart, that in the end we may attain, as he hath, to that great reward which Thou hast promised unto all such as be so minded.
	Through Jesus Christ, Thy Son our Lord, Who liveth and reigneth with Thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}