\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Resurrection of the Alleluia; Easter Sunday, April 17, 2022}

 \textit{In accordance with traditional custom, the Alleluia is resurrected (unburied), in honor of Our Lord and of the joyous returning of this word of exaltation in the traditional liturgy.}
 
 \begin{rubricbox}
 
 {\color{red} The Cantor intones the Alleluia and the beginning of Psalm 150, which is then sung by the choir. All process to the burial location while the Psalm is sung.}
 
 \end{rubricbox}
 
\noindent Ant. \textit{Alleluia}. Ps. \textit{Praise ye the Lord in His holy places: praise ye Him in the firmament of His power.}
 
 \gresetinitiallines{0}
 \gregorioscore{alleluia_psalm_150_1_9}
 
\begin{latinenglishsection}

\latinenglish{
	\input{psalm_150_9}
}{
	2. Praise ye Him for His mighty acts: praise ye Him according to the multitude of His greatness.\\
	3. Praise Him with sound of trumpets: praise Him with psaltery and harp.\\
	4. Praise Him with timbrel and choir: praise Him with strings and organs.\\
	5. Praise Him on high sounding cymbals: praise Him on cymbals of joy. Let every spirit praise the Lord.\\
	6. Glory be to the Father, and to the Son, and to the Holy Spirit.\\
	7. As it was in the beginning, is now, and ever shall be, world without end. Amen.
}
\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{alleluia}

\vfill\pagebreak

 \begin{rubricbox}
 
 {\color{red} The Hymn `Ad Coenam' is sung by the choir.}
 
 \end{rubricbox}
 
 \gresetinitiallines{1}
 \gregorioscore{ad_coenam}
 
 \textit{1. The Lamb's high banquet we await in snow-white robes of royal state: and now, the Red Sea's channel past, to Christ our Prince we sing at last.}\\
\textit{2. Upon the Altar of the Cross His Body hath redeemed our loss: and tasting of his roseate Blood, our life is hid with Him in God.}\\
\textit{3. That Paschal Eve God's arm was bared, the devastating Angel spared: by strength of hand our hosts went free from Pharaoh's ruthless tyranny.}\\
\textit{4. Now Christ, our Paschal Lamb, is slain, the Lamb of God that knows no stain, the true Oblation offered here, our own unleavened Bread sincere.}\\
\textit{5. O Thou, from whom hell's monarch flies, O great, O very Sacrifice, Thy captive people are set free, and endless life restored in Thee.}\\
\textit{6. For Christ, arising from the dead, from conquered hell victorious sped, and thrust the tyrant down to chains, and Paradise for man regains.}\\
\textit{7. We pray Thee, King with glory decked, in this our Paschal joy, protect from all that death would fain effect Thy ransomed flock, Thine own elect.}\\
\textit{8. To Thee who, dead, again dost live, all glory Lord, Thy people give; all glory, as is ever meet, to Father and to Paraclete. Amen.}

\begin{rubricbox}

{\color{red} The Alleluia is dug out of the ground. After it has been obtained, the Officiant prays the following prayer:}

\end{rubricbox}

Oremus. Effundam de aliis super nos, Domine, spiritum amoris Tui: {\color{red}\GreDagger} ut, quorum fames Tu habes contentus Sacra\textit{menta} \textbf{Pas}cha, ut in Tuam misericordiam unum esse in corde. {\color{red}*} Per Dominum nostrum Iesum Christum, Fílium tuum, Qui vivit et regnat Tecum in unitate Spiritus Sancti, Deus, per omnia in saecula saeculorum.\\
{\color{red}\Rbar.} Amen.

\begin{rubricbox}

{\color{red} The Officiant may now sprinkle the `Alleluia' with Holy Water: He then leads the following:}

\end{rubricbox}

{\color{red}\Vbar.} Dominus vobiscum.\\
{\color{red}\Rbar.} Et cum spiritu tuo.

{\color{red}\Vbar.} Pax tibi. Alleluia. Alleluia.\\
{\color{red}\Rbar.} Deo gratias. Alleluia. Alleluia

\textit{Let us pray: Pour forth upon us, O Lord, the spirit of Thy love: that those whose hunger Thou hast satisfied with the Sacraments of Easter may in Thy kindness be one in heart.  Through our Lord Jesus Christ, Thy Son, Who lives and reigns with Thee in the unity of the Holy Spirit, God, forever and ever. \Rbar. Amen. \Vbar. Peace be with you. Alleluia. Alleluia. \Rbar. Thanks be to God. Alleluia. Alleluia.}

\begin{rubricbox}

{\color{red} The Cantor leads the Marian anthem \textit{Regina Caeli} (no. 221 in the \textit{Traditional Roman Hymnal}).}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{regina_caeli}

\textit{Queen of Heaven, rejoice, alleluia. For He whom you did merit to bear, alleluia.}\\
\textit{Has risen, as he said, alleluia. Pray for us to God, alleluia.}\\
\textit{Rejoice and be glad, O Virgin Mary, alleluia. For the Lord has truly risen, alleluia.}

{\color{red} \textit{The Officiant leads the concluding collect of the Marian anthem:}}

\begin{latinenglishsection}

\latinenglish{
	Oremus. Deus, qui per resurrectionem Filii tui, Domini nostri Iesu Christi, mundum laetificare dignatus es: {\color{red}\GreDagger}\  praesta, quaesumus;  ut per eius Genetricem Vir\textit{ginem} \textbf{Ma}riam, perpetuae capiamus gaudia vitae. {\color{red}*} Per eundem Christum Dominum nostrum. Amen.
}{
	Let us pray. 
	O God, who gave joy to the world through the resurrection of Thy Son, our Lord Jesus Christ, grant we beseech Thee, that through the intercession of the Virgin Mary, His Mother, we may obtain the joys of everlasting life. Through the same Christ our Lord. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red} The ceremony is now ended. All may recess to the church.}

\end{rubricbox}

\end{document}