\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, May 21, 2022}

 \textit{Vespers today is based on the following Sunday, which is the fifth after Easter.
 All is said as in the Eastertide booklet until the Little Chapter.}

\section*{Little Chapter (James 1:22:24)}

\begin{latinenglishsection}

 \latinenglish{
	 Caríssimi: Estóte factóres verbi, et non auditóres tantum, falléntes vosmetípsos.
	 Quia si quis audítor est verbi et non factor: hic comparábitur viro consideránti vultum nativitátis suæ in spéculo:
	 considerávit enim se et ábiit, et statim oblítus est, qualis fúerit.
 	\Rbar.~Deo gratias.
 }{
 	Beloved: Be ye doers of the word, and not hearers only, deceiving your own selves.
 	For if a man be a hearer of the word, and not a doer, he shall be compared to a man beholding his own countenance in a glass.
 	For he beheld himself, and went his way, and presently forgot what manner of man he was.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\begin{rubricbox}

The hymn \textit{Ad cenam agni providi} and versicle \textit{\Rbar.~Mane nobiscum\dots} are said as given in the booklet for Eastertide.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../hymns/ad-coenam-agni}

{\itshape
	1. The Lamb's high banquet we await
	in snow-white robes of royal state:
	and now, the Red Sea's channel past,
	to Christ our Prince we sing at last. 
	
	2. Upon the Altar of the Cross
	His Body hath redeemed our loss:
	and tasting of his roseate Blood,
	our life is hid with Him in God. 
	
	3. That Paschal Eve God's arm was bared,
	the devastating Angel spared:
	by strength of hand our hosts went free
	from Pharaoh's ruthless tyranny. 
	
	4. Now Christ, our Paschal Lamb, is slain,
	the Lamb of God that knows no stain,
	the true Oblation offered here,
	our own unleavened Bread sincere. 
	
	5. O Thou, from whom hell's monarch flies,
	O great, O very Sacrifice,
	Thy captive people are set free,
	and endless life restored in Thee.
	
	6. For Christ, arising from the dead,
	from conquered hell victorious sped,
	and thrust the tyrant down to chains,
	and Paradise for man regains. 
	
	7. We pray Thee, King with glory decked,
	in this our Paschal joy, protect
	from all that death would fain effect
	Thy ransomed flock, Thine own elect.
	
	8. To Thee who, dead, again dost live,
	all glory Lord, Thy people give;
	all glory, as is ever meet,
	to Father and to Paraclete. Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ma(h)ne(h) no(h)bís(h)cum(h) Dó(h)mi(h)ne,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Quó(h)ni(h)am(h) ad(h)ves(h)pe(h)rá(h)scit,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Stay with us, Lord, alleluia.
\Rbar.~For it is getting towards evening, alleluia.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-2}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Hitherto you have not asked anything in My Name. Ask, and you shall receive, alleluia.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, a quo bona cuncta procédunt, largíre supplícibus tuis:
	ut cogitémus, te inspiránte, quæ recta sunt;
	et, te gubernánte, éadem faciámus.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, from Whom all good things do come,
	grant to us Thy humble servants that by Thy holy inspiration we may think those things that be good,
	and by Thy merciful guiding may perform the same. 
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\pagebreak

\section*{Suffrage of the Saints in Paschal Time}

\textit{\textnormal{Ant.} He that was crucified is risen from the dead, and hath redeemed us. Alleluia, Alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/suffrage-saints-pascha}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Dícite in natiónibus, allelúja.\\
	\Rbar. Quia Dóminus regnávit a ligno, allelúja.
	
	Orémus.
	Deus qui pro nobis Fílium tuum Crucis patíbulum subíre voluísti, ut inimíci a nobis expélleres potestátem: concéde nobis fámulis tuis; ut resurrectiónis grátiam consequámur.
	Per eúmdem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	\Vbar. ~Say among the heathen, Alleluia
	\Rbar.~That the Lord reigneth from the tree, Alleluia
	
	Let us pray.
O God, who didst send thy Son to suffer death for us upon the Cross, that thou mightest deliver us from the power of the enemy grant unto us thy servants to be made partakers of his Resurrection.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain with us always. \Rbar.~Amen.
}

\end{latinenglishsection}

\end{document}