\documentclass{../vespers-sheet}

\begin{document}

%% TODO: correct date
\chapter*{Vespers Insert for Saturday, 12 Feb 2022}

 \textit{Vespers today is based on the following Sunday, which is Septuagesima Sunday.
 All is said as in the main booklet until the Little Chapter;
this liturgy is the last time the word \textnormal{Alleluia} is said until Easter, as it is entirely omitted from the liturgy
during the seasons of Septuagesima and Lent.}

 \section*{Little Chapter (1 Cor. 9:24)}

\begin{latinenglishsection}

%% IMPORTANT NOTE: this chapter has to be chanted using the tone for a question (consult the Antiphonal or Liber)
 \latinenglish{
 	Fratres: nescítis quod ii qui in stádio current,~\GreDagger\
 	omnes quidem current, sed unus áccipit bravíum?~*
 	Sic cúrrite ut comprehendátis.
 	\Rbar.~Deo gratias.
 }{
 	Brethren: Know you not that they that run in the race, all run indeed, but one receiveth the prize? So run that you may obtain. 
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/o-lux-beata-trinitas}

{\itshape
	1. O Trinity of blessed Light,
	O Unity of sovereign might,
	as now the fiery sun departs,
	shed Thou Thy beams within our hearts.
	
	2. To Thee our morning song of praise,
	to Thee our evening prayer we raise;
	Thee may our glory evermore
	in lowly reverence adore.
	
	3. All laud to God the Father be;
	all praise, Eternal Son, to Thee;
	all glory, as is ever meet,
	to God the Holy Paraclete.
}

\vfill\pagebreak

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h) Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~May our evening prayer ascend to Thee, O Lord.
\Rbar.~And may Thy mercy descend upon us.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} The Lord said to Adam: of the tree which is in the middle of paradise, thou shalt not eat;
in the hour thou eatest of it, thou shalt surely die.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Preces pópuli tui, qu\'{\ae}sumus, Dómine, cleménter exáudi:
	ut, qui juste pro peccátis nostris afflígimur,
	pro tui nóminis glória misericórditer liberémur.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, we beseech Thee, graciously hear the prayers of Thy people,
	that we who are justly punished for our sins may be mercifully delivered for the glory of Thy name.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of the Seven Founders of the Servite Order, Confessors (Feb. 12)}

\textit{\textnormal{Ant.} Their name endureth for ever, enduring unto their sons, the glory of holy men.}

\gresetinitiallines{1}
\gregorioscore{comm-servite-founders-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Hi viri misericórdiæ sunt, quorum pietátes non defuérunt.\\
	\Rbar.~Semen eórum et glória eórum non derelinquétur.
	
	Orémus.
	Dómine Jesu Christe, qui ad recoléndam memóriam dolórum sanctíssimæ Genitrícis tuæ,
	per septem beátos Patres nova Servórum ejus família Ecclésiam tuam fecundásti:
	concéde propítius; ita nos eórum consociári flétibus,
	ut perfruámur et gáudiis.
	Qui vivis.
}{
	These were merciful men whose righteousness hath not been forgotten.
	\Rbar.~Their seed and their glory shall not be blotted out.
	
	Let us pray.
	Lord Jesus Christ, who, that Thou mightest recall to mind the woes of Thy most holy Mother,
	didst through the Seven blessed Fathers make Thy Church herself the mother of a new household of her servants,
	grant unto us in mercy that we may so share their tears as to share their blessedness also.
	Who livest.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord, alleluia, alleluia. \Rbar.~Thanks be to God, alleluia, alleluia.
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{../common/benedicamus-alleluia}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}