\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers of the Saturday in the Octave of the Ascension}

\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} Men of Galilee. \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}

\begin{rubricbox}

All \textbf{stand} until the asterisk halfway through the first verse, then \textbf{sit}.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps109-7}
	
}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\pagebreak

\gresetinitiallines{0}
\gregorioscore{ps109-antiphon}
\textit{\textnormal{Ant.} Men of Galilee, why do you stand looking up to heaven? This Jesus Who has been taken up from you in heaven, will come in the same way, alleluia.}

\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} While. \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}

\begin{rubricbox}

All \textbf{stand} until the asterisk halfway through the first verse, then \textbf{sit}.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps110-8}
	
}{
	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps110-antiphon}
\textit{\textnormal{Ant.} While they were beholding Him going up to heaven, they said: alleluia.}

\pagebreak

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} Lifting up His hands. \textnormal{Ps.} Blessed is the man that feareth the Lord: he shall delight exceedingly in his commandments.}

\begin{rubricbox}

All \textbf{stand} until the asterisk halfway through the first verse, then \textbf{sit}.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps111-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps111-4-E-A-d}
	
}{
	\input{../psalms/english/ps111}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps111-antiphon}
\textit{\textnormal{Ant.} Lifting up His hands, He blessed them and was carried up to heaven, alleluia.}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} Praise ye. \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}

\begin{rubricbox}

All \textbf{stand} until the asterisk halfway through the first verse, then \textbf{sit}.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps112-8}
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps112-antiphon}
\textit{\textnormal{Ant.} Praise ye the King of Kings, and sing a hymn to God, alleluia.}

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} While they were looking on. \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\begin{rubricbox}

All \textbf{stand} until the asterisk halfway through the first verse, then \textbf{sit}.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps116-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps116-8}
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps116-antiphon}
\textit{\textnormal{Ant.} While they were looking on, He was raised up and a cloud received Him into heaven, alleluia.}

\section*{Little Chapter (1 Pet. 4:7-8)}

\begin{rubricbox}

All \textbf{stand} as the leader chants:

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	 Caríssimi: Estóte prudéntes et vigiláte in oratiónibus.~\GreDagger\
	 Ante ómnia autem mútuam in vobismetípsis caritátem contínuam habéntes:~*
	 quia cáritas óperit multitúdinem peccatórum.
 	\Rbar.~Deo gratias.
 }{
 	Beloved: Be prudent, and watch in prayers.
 	But before all things have a constant mutual charity among yourselves:
 	for charity covereth a multitude of sins.
 	\Rbar.~Thanks be to God.
 }

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/jesu-nostra-redemptio}

{\itshape

 1. Hail, Thou who man's Redeemer art,
Jesu, the joy of every heart;
Great Maker of the world's wide frame,
And purest love’s delight and flame: 

2. What nameless mercy Thee o'ercame,
To bear out load of sin and shame?
For guiltless, Thou Thy life didst give,
That sinful erring man might live. 

3. The realms of woe are forced by Thee,
Its captives from their chains set free;
And Thou, amid Thy ransomed train,
At God’s right hand dost victor reign.

4. Let mercy sweet with Thee prevail,
To cure the wounds we now bewail;
Oh, bless us with Thy holy sight,
And fill us with eternal light.

5. Our guide, our way to heavenly rest,
Be Thou the aim of every breast;
Be Thou the soother of our tears,
Our sweet reward above the spheres.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Dó(h)mi(h)nus(h) in(h) cæ(h)lo,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Pa(h)rá(h)vit(h) se(h)dem(h) su(h)am,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~The Lord in heaven, alleluia.
\Rbar.~Hath established his throne, alleluia.}

\section*{Magnificat}

\textit{\textnormal{Ant. Magn.} When he has come.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/magnificat-8}
	
}{
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} When the Paraclete has come, Whom I will send you from the Father, the Spirit of Truth Who proceeds from the Father, He will bear witness concerning Me, alleluia.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}
\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens sempitérne Deus:
	fac nos tibi semper et devótam gérere voluntátem;
	et majestáti tuæ sincéro corde servíre.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and eternal God, give us a will ever dedicated to You, and a true heart to serve Your majesty.
	Through our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Augustine of Canterbury, Bishop and Confessor (May 28)}

\textit{\textnormal{Ant.} The Lord loved him, and adorned him; He clothed him with a robe of glory, and crowned him at the gates of Paradise, alleluia}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-pont-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas, allelúia.\\
	\Rbar.~Et osténdit illi regnum Dei, allelúia.
	
	Orémus.
	Deus, qui Anglórum gentes, prædicatióne et miráculis beáti Augustíni Confessóris tui atque Pontíficis, veræ fídei luce illustráre dignátus es: concéde; ut, ipso interveniénte, errántium corda ad veritátis tuæ rédeant unitátem, et nos in tua simus voluntáte concórdes.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	The Lord led the just in right paths, alleluia.
	\Rbar.~And showed him the kingdom of God, alleluia.
	
	Let us pray.
	O God, Who graciously enlightened the English peoples with the light of the true faith by the preaching and miracles of blessed Augustine, thy Confessor and Bishop, grant, through his intercession, that the hearts of those who have strayed may return to the unity of the true faith and that we may be in harmony with thy will.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\section*{Commemoration of St. Mary Magdalene de' Pazzi, Virgin (May 29)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-1v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua, allelúia.\\
	\Rbar.~Inténde, próspere procéde, et regna, allelúia.
	
	Orémus.
	Deus, virginitátis amátor, qui beátam Maríam Magdalénam Vírginem, tuo amóre succénsam, cæléstibus donis decorásti: da; ut, quam festíva celebritáte venerámur, puritáte et caritáte imitémur.
}{
	In thy glory and thy splendor, alleluia.
	\Rbar.~Go forth, advance with victory and reign, alleluia.
	
	Let us pray.
	O God, lover of chastity, Who endowed with heavenly gifts blessed Mary Magdalen, a virgin on fire with love for You, grant that we who keep this feast-day in her honor may imitate her by purity and love.
}

\end{latinenglishsection}

\section*{Commemoration of the Ascension}

\textit{\textnormal{Ant.} King of Glory, Lord of hosts, Who this day hast triumphantly ascended above all the heavens, 
leave us not orphans, but send upon us the promise of the Father, the Spirit of Truth, alleluia.}

\gresetinitiallines{1}
\gregorioscore{comm-oct}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Ascéndit Deus in jubilatióne, allelúia.\\
	\Rbar. Et Dóminus in voce tubæ, allelúia.
	
	Orémus.
	Concéde, qu\'{\ae}sumus, oninípotens Deus:
	ut, qui hodiérna die Unigénitum tuum, Redemptórem nostrum, ad c{\oe}los ascendísse crédimus;~*
	ipsi quoque mente in c{\oe}léstibus habitémus. 
	Per eúmdem Dóminum.
}{
	God ascends among shouts of triumph, alleluia.
	\Rbar.~And the Lord with the sound of trumpets, alleluia.
	
	Let us pray.
	Grant, we beseech thee, Almighty God,
	that just as we believe Thine Only-Begotten Son our Saviour to have this day ascended into the heavens,
	so we may also in heart and mind thither ascend, and with Him continually dwell.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\gresetinitiallines{0}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\rubric{If no other service follows:}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Cæli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain with us always. \Rbar.~Amen.
}

\end{latinenglishsection}

\end{document}