\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

% TODO: Update the title for the specific feast
\chapter*{Second Vespers of the Most Precious Blood of Our Lord Jesus Christ}

%\section*{Beginning of the Office}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary). Then all rise with the Officiant make the sign of the cross with him as he intones:}

\end{rubricbox}

% TODO: Make sure that the tone of the deus adjutorium matches the season primarily and the solemnity of the feast secondarily
 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
{\color{red}\Vbar.}~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

%TODO: Add that the correct psalms, and verify that their tones, and their associated pointed text are correct

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} Who is this that * cometh from Edom, with dyed garments from Bozrah? this, that is glorious in His apparel?
 \textnormal{Ps.} The Lord said to my Lord: * Sit thou at my right hand.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-7} %%

}{
	\input{../psalms/english/ps109} %%
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

%\begin{rubricbox}

%{\color{red}The antiphon is repeated: \textit{Quis est iste\dots}} %%

%\end{rubricbox}

%TODO: Verify that the little chapter is fitting for the feast

\section*{Little Chapter (Hebrews 9:11-12)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Fratres: Christus assistens póntifex futurórum bonórum, per ámplius et perféctius tabernáculum non\\ manufáctum, id est, non hujus creatiónis: {\color{red}\GreDagger} neque per sánguinem hircórum, aut vitulórum, sed per próprium sánguinem introívit semel in Sancta, ætérna redemptióne\\ invénta. 
	{\color{red}\Rbar.}~Deo grátias.
}{
	Brethren, Christ being come, a high priest of the good things to come, by a greater and more perfect tabernacle, not made with hands, that is, not of this creation, neither by the blood of goats or of calves, but by His own blood, entered once into the Holies, having obtained eternal redemption. 
	 {\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

% TODO: Verify that the hymn is correct for the feast (including the responsory after the hymn)

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/festivis_resonent}
% Translation from: https://tosingistopraytwice.wordpress.com/2017/03/29/forth-let-the-long-procession-stream/
{\itshape
	1.  Forth let the long procession stream,
	And through the streets in order wend;
	Let the bright waving line of torches gleam,
	The solemn chant ascend.
	
	2. While we, with tears and sighs profound,
	That memorable Blood record,
	Which, stretch’d on his hard Cross, from many a wound
	The dying Jesus pour’d.
	
	3. By the first Adam’s fatal sin
	Came death upon the human race;
	In this new Adam doth new life begin,
	And everlasting grace.
	
	4. For scarce the Father heard from Heaven
	The cry of his expiring Son,
	When in that cry our sins were all forgiven,
	And boundless pardon won.
	
	5. Henceforth, whoso in that dear Blood
	Washeth, shall lose his every stain;
	And in immortal roseate beauty rob’d,
	An angel’s likeness gain.
	
	6. Only, run thou with courage on
	Straight to the goal set in the skies;
	He, who assists thy course, will give thee soon
	The everlasting prize.
	
	7. Father supreme! vouchsafe that we,
	For whom thine only Son was slain,
	And whom thy Holy Ghost doth sanctify,
	May heavenly joys attain.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>V/</sp>.</c> Te(h) er(h)go(h), quáe(h)su(h)mus(h), tu(h)is(h) fá(h)mu(h)lis(h) súb(h)ve(h)ni.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>R/</sp>.</c> Quos(h) pre(h)ti(h)ó(h)so(h) sán(h)gui(h)ne(h) re(h)de(h)mís(h)ti.(g'_/hvGF'E/fgf.) (::)
}

\textit{{\color{red}\Vbar.}~We therefore pray Thee, help Thy servants.
{\color{red}\Rbar.}~O Thou hast redeemed us with Thy precious blood.}

%TODO: Verify that the magnificat antiphon is correct and match the mangificat intonation and pointed text with the tone

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Behold a faithful and prudent servant, whom the Lord has set over His household.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.}

\begin{rubricbox}

{\color{red}The Cantor leads by intoning the antiphon and the first verse.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All \textbf{remain standing} and make the sign of the cross with the Cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
3. Quia respéxit humilitátem \textit{an}\textit{cíl}\textit{læ} \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.

4. Quia fecit mihi \textit{ma}\textit{gna} \textit{qui} \textbf{pot}\textbf{ens} est:~* et sanctum \textit{no}\textit{men} \textbf{e}jus.

5. Et misericórdia ejus a progéni\textit{e} \textit{in} \textit{pro}\textbf{gé}\textbf{ni}es~* timén\textit{ti}\textit{bus} \textbf{e}um.

6. Fecit poténtiam in \textit{brá}\textit{chi}\textit{o} \textbf{su}o:~* dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.

7. Depósuit pot\textit{én}\textit{tes} \textit{de} \textbf{se}de,~* et exal\textit{tá}\textit{vit} \textbf{hú}\textbf{mi}les.

8. Esuriéntes \textit{im}\textit{plé}\textit{vit} \textbf{bo}nis:~* et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.

9. Suscépit Israël \textit{pú}\textit{e}\textit{rum} \textbf{su}um,~* recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.

10. Sicut locútus est \textit{ad} \textit{pa}\textit{tres} \textbf{nos}tros,~* Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}\textbf{cu}la.
}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{\color{red}(bow)}Glória Patri, et \textbf{Fí}lio,~* 
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{\color{red}(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All  \textbf{remain standing}.}

\end{rubricbox}

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}
% Concluding collect from the Litany of the Most Precious Blood: https://www.ewtn.com/catholicism/library/litany-of-the-most-precious-blood-of-our-lord-jesus-christ-11888
\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
	
	Orémus.
	Omnipotens sempiterne Deus, qui unigenitum Filium tuum mundi Redemptorem constituisti, ac ejus sanguine placari voluisti: {\color{red}\GreDagger} concede, quaesumus, salutis nostrae pretium solemni cultu ita venerari, atque a praesentis vitae malis ejus virtute defendi in terris; * ut fructu perpetuo laetemur in caelis. Per eumdem Dominum nostrum Iesum Christum Filium Tuum, qui vivit et regnat in unitate Spiritu Sancti, Deus, per omnia saecula saeculurom.
	{\color{red}\Rbar.}~Amen.
}{
	{\color{red}\Vbar.}Lord, hear my prayer.
	{\color{red}\Rbar.}~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and eternal God, Thou hast appointed Thine only-begotten Son the Redeemer of the world and willed to be appeased by his blood. Grant, we beg of Thee, that we may worthily adore this price of our salvation and through its power be safeguarded from the evils of the present life so that we may rejoice in its fruits forever in heaven. Through the very same Jesus Christ Thy Son, who liveth and reigneth with Thee in the unity of the Holy Spirit, God, world without end.
	{\color{red}\Rbar.}~Amen.
}
\end{latinenglishsection}

%TODO: Add commemorations for the date

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of the Visitation of the Blessed Virgin Mary}

\textit{\textnormal{Ant.} Blessed art thou, Mary, who didst believe the Lord; what the Lord said to thee shall be fulfilled in thee. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--beata_es_maria--solesmes}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Benedicta tu in mulieribus.\\
	{\color{red}\Rbar.}~Et benedictus fructus ventris tui.
	
	Orémus.
	Famulis tuis, quaesumus Domine, caelestis gratiae munus impertire: {\color{red}\GreDagger} ut quibus beatae Virginis partus exstitit salutis ex6rdium, • Visitati6nis ejus votiva solemnitas pacis tribuat incre- mentum.
	Per Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	{\color{red}\Vbar.}~Blessed art thou amongst women.
	{\color{red}\Rbar.}~And blessed is the fruit of thy womb.
	
	Let us pray.
	Grant, O Lord, we beseech thee, unto all thy servants, the gift of thine heavenly grace, that even as the Blessed Virgin being made a Mother hath been unto them the first step unto their salvation, so the godly and solemn memorial of her Visitation may be the bringer of an increase of peace.
	Through Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Dómine exáudi oratiónem meam.\\
	{\color{red}\Vbar.} Et clamor meus ad te véniat.
}{
	{\color{red}\Vbar.} Lord, hear my prayer. {\color{red}\Rbar.} And let my cry come unto Thee.
	{\color{red}\Vbar.} Let us bless the Lord. {\color{red}\Rbar.} Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-2v-solem}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Fidélium ánimæ, per\\ misericórdiam Dei, requiéscant in pace. \\
	{\color{red}\Rbar.} Amen.
}{
	{\color{red}\Vbar.} May the souls of the faithful departed, through the mercy of God, rest in peace. {\color{red}\Rbar.}~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	{\color{red}\Vbar.} Dóminus det nobis suam pacem. \\
	{\color{red}\Rbar.} Et vitam ætérnam. Amen.
}{
	{\color{red}\Vbar.} May the Lord grant us his peace. 
	{\color{red}\Rbar.} And life eternal. Amen.
}

\end{latinenglishsection}

%TODO: Add the Marian Anthem for the season and verify that the oration afterwards is correct

\section*{Marian Anthem}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/salve-regina-solemn-tone}

{\itshape
	Hail, holy Queen, Mother of mercy, our life, our sweetness and our hope.
	To thee do we cry, poor banished children of Eve. 
	To thee do we send up our sighs, mourning and weeping in this valley of tears. 
	Turn, then, most gracious advocate, thine eyes of mercy toward us, 
	and after this, our exile, show unto us the blessed fruit of thy womb, Jesus. 
	O clement, O loving, O sweet Virgin Mary.
}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Ora pro nobis, sancta Dei Genitrix.\\
	{\color{red}\Rbar.} Ut digni efficamur promissionibus Christi.
	
	Oremus. 
	Omnipotens sempiterne Deus, qui gloriosae Virginis Matris Mariae corpus et animam, ut dignum Filii tui habitaculum effici mereretur, Spiritu Sancto cooperante, praeparasti, da, ut cuius commemoratione laetamur; eius pia intercessione, ab instantibus malis et a morte perpetua liberemur. Per eundem Christum Dominum nostrum.
	\Rbar. Amen.
}{
	{\color{red}\Vbar.} Pray for us, O holy Mother of God.
	{\color{red}\Rbar.} That we may be made worthy of the promises of Christ.
	
	Let us pray. 
	Almighty and everlasting God, Who by the working of the Holy Spirit didst prepare both body and soul of the glorious Virgin Mother, Mary, that she might deserve to be made a worthy dwelling for Thy Son, grant that we who rejoice in her memory, may, by her loving intercession, be delivered from present evils and from lasting death, through the same Christ our Lord. 
	{\color{red}\Rbar.} Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Divínum auxílium máneat semper nobíscum.\\
	{\color{red}\Rbar.} Amen.
}{
	May the divine assistance remain always with us.
	{\color{red}\Rbar.}~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

{\color{red} After the Office, all \textbf{kneel} and pray in silence for a time.}

\end{rubricbox}

\end{document}