\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, June 10, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fourteenth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Rom 11:33)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\
	 quam incomprehensibília sunt ju\textit{dícia} \textbf{e}jus, et investigábiles viæ ejus!
 	\Rbar.~Deo gratias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape

1. As fades the glowing orb of day,
To thee, great source of light, we pray;
Blest Three in One, to every heart
Thy beams of life and love impart.

2. At early dawn, at close of day,
To thee our vows we humbly pay;
May we, mid joys that never end,
With thy bright saints in homage bend.

3. To God the Father, and the Son,
And Holy Spirit, Three in One,
Be endless glory, as before
The world began, so evermore.
Amen.

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h)scén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\vfill\pagebreak

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Remember not, Lord mine offences, nor the offences of my forefathers, neither take Thou vengeance of my sins.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Ecclésiam tuam, Dómine, miserátio continuáta mundet et múniat: et quia sine te non potest salva consístere; tuo semper múnere gubernétur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, we beseech thee, let thy continual pity cleanse and defend thy Church, and because it cannot continue in safety without thy succour, preserve it evermore by thy help and goodness.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of the Impression of the Stigmata of St. Francis (September 17)}

\textit{\textnormal{Ant.} This man, despising the world, and triumphing over earthly things, hath laid up treasures in heaven by word and deed.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-2v}
% PT: \gregorioscore{../commemorations/conf-non-pont-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ %PT: allelúia
	\Rbar.~Et osténdit illi regnum Dei.  %PT: allelúia
	
	Orémus.
	Dómine Jesu Christe, qui, frigescénte mundo, ad inflammándum corda nostra tui amóris igne, in carne beatíssimi Francísci passiónis tuæ sacra Stígmata renovásti: concéde propítius; ut ejus méritis et précibus crucem iúgiter ferámus, et dignos fructus pæniténtiæ faciámus:
	Qui vivis et regnas cum Deo Patre, in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	The Lord led the just in right paths. %PT: alleluia
	\Rbar.~And showed him the kingdom of God. %PT: alleluia
	
	Let us pray.
	O Lord Jesus Christ, Who, when the love of many was waxing cold, didst manifest once more the holy marks of thine own Suffering in the flesh of thy most blessed servant Francis, to the end that our hearts might kindle again with the fire of the love of thyself, be Thou entreated for thy servant's sake, and grant to his and our prayers that we may effectually carry thy Cross and bring forth fruits meet for repentance.
	Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Joseph Cupertino (September 18)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}
% PT: \gregorioscore{../commemorations/conf-non-pont-1v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ induit eum.  %PT: allelúia
	
	Orémus.
	Deus, qui ad unigénitum Fílium tuum exaltátum a terra ómnia tráhere disposuísti: pérfice propítius; ut, méritis et exémplo seráphici Confessóris tui Joséphi, supra terrénas omnes cupiditátes eleváti, ad eum perveníre mereámur:
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	The Lord loved him and adorned him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	O God, Who art pleased that thine Only-begotten Son being lifted up from the earth should draw all things unto Him, be entreated for the sake of thy servant Joseph, whom Thou didst make like unto one of the Seraphim, and so effectually work in us, that even as he, we also may be drawn up above all earthly lusts, and worthily attain unto Him:
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}