\documentclass{../vespers-sheet}

\begin{document}

%% TODO: correct date
\chapter*{Vespers Insert for Saturday, September 24, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Sixteenth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

%% TODO: correct the little chapter
 \section*{Little Chapter (Rom 11:33)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,\GreDagger\
	 quam incomprehensibília sunt ju\textit{dícia} \textbf{e}jus, et investigábiles viæ ejus!
 	\Rbar.~Deo gratias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

%% TODO: correct the hymn using the antiphonary
\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h)scén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

%% TODO: correct the magnificant antiphon and tone - add pointed text to the intonation

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} O Adonai, O Lord God, Thou art great and glorious, Who hast given salvation into the hand of a woman, graciously hear the prayers of thy servants.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

%% TODO: proper collect for the date

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Tua nos, qu\'{\ae}sumus, Dómine, grátia semper et prævéniat et sequátur: \GreDagger ac bonis opéribus júgiter præstet es\textit{se in}\textbf{tén}tos.
Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Lord, we pray thee, that thy grace may always prevent and follow us, and make us continually to be given to all good works.
Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

%% TODO: commemorations based on the date
\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of Our Lady of Ransom (September 24)}

\textit{\textnormal{Ant.}  All generations shall call me blessed, for God hath regarded the lowliness of His hand-maiden.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--beatam_me_dicent--solesmes_1934}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Dignáre me laudáre te, Virgo sacráta.\\
	\Rbar.~Da mihi virtútem contra hostes tuos.
	
	Orémus.
	Deus, qui per gloriosíssimam Fílii tui Matrem, ad liberándos Christi fidéles a potestáte paganórum nova Ecclésiam tuam prole amplificáre dignátus es: præsta, qu\'{\ae}sumus; ut, quam pie venerámur tanti operis institutrícem, ejus páriter méritis et intercessióne, a peccátis ómnibus et captivitáte d\'{\ae}monis liberémur.
	Per eúndem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	\Vbar.~Let the evening prayer ascend unto thee, O Lord.
	\Rbar.~And let there descend upon us thy mercy.
	
	O God, Who didst use the glorious Mother of thy Son as a mean to ransom Christ's faithful people out of the hands of the unbelievers, by enriching thy Church with yet another family, grant, we beseech thee, that we who reverently honour her as the Foundress of that great work, may for her sake and by her prayers, be redeemed from all sin and all bondage unto the evil one.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

%% TODO: benedicamus in line with the season

\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

%% TODO: proper Marian anthem for the seasons

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}