\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, July 30}

 \textit{Vespers today is based on the following Sunday, which is the Eighth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 11:33)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
 	O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\
 	quam incomprehensibília sunt judícia ejus, et investigábiles viæ ejus!
	 
 	\Rbar.~Deo grátias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h), Dó(h)mi(f)ne(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a(f.). (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Wisdom hath builded her house, she hath hewn out her seven pillars, she hath broken the nations under her in the greatness of her strength, she hath trodden under her feet the necks of such as be haughty and proud.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Largíre nobis, qu\'{\ae}sumus, Dómine, semper spíritum cogitándi quæ recta sunt, propítius et agéndi: ut, qui sine te esse non póssumus, secúndum te vívere valeámus.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Grant to us, Lord, we beseech thee, the spirit to think and do always such things as be rightful, that we, who cannot do anything that is good without thee, may by thee be enabled to live according to thy will.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of Sts. Abdon and Sennen, Marytrs (July 30)}

\textit{\textnormal{Ant.} Theirs is the kingdom of heaven, who despising a worldy life have attained the rewards of the kingdom,
and have washed their robes in the Blood of the Lamb.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Lætámini in Dómino, et exsultáte justi.\\
	\Rbar.~Et gloriámini omnes recti corde.
	
	Orémus.
	Deus, qui sanctis tuis Adbdon et Sennen ad hanc gloriam veniendi copiosum munus gratiae contulisti: \GreDagger da famulis tuis suorum veniam \textit{pecca}\textbf{to}rum; * ut Sanctorum tuorum intercedentibus meritis, ab omnibus mereanter adversitatibus liberari.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	Be glad in the Lord and rejoice, ye just.
	\Rbar.~And exult, all ye upright of heart.
	
	Let us pray.
	O God, your holy saints Abdon and Sennen were given the rich gift of your grace: grant to us by their service the forgivenss of our sins; by the merits of the Saints who intercede for us, that we may be delivered from all adversity.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\section*{Commemoration of St. Ignatius of Loyola, Confessor (July 31)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ 
	\Rbar.~Stolam glóriæ induit eum.
	
	Orémus.
	Deus, qui ad majórem tui nóminis glóriam propagándam, novo per beátum Ignátium subsídio militántem Ecclésiam roborásti: concéde; ut, ejus auxílio et imitatióne certántes in terris, coronári cum ipso mereámur in cælis.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	The Lord loved him and adorned him. 
	\Rbar.~He clothed him with a robe of glory. 
	
	Let us pray.
	O God, Who, to spread abroad the greater glory of Thy name through St. Ignatius, strengthened the Church militant with new power; grant that we, who are struggling on earth, may, by his help and after his example, be found worthy to be crowned with him in heaven.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}