\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Burial of the Alleluia}

\begin{rubricbox}

{\color{red}A traditional paraliturgical custom in the West involves the burying of a sign with `Alleluia' written on it, immediately after Vespers preceding Septuagesima Sunday, for this is the last time the sweet word `Alleluia' shall be heard until Easter. The faithful, led by vested clerics and servers, process with a banner with the word \textit{Alleluia} on it, to the location where it should be buried. As the procession proceeds, the cantors lead the singing of the following psalm and hymn.}

\end{rubricbox}

\section*{Psalm 125}

Ant. \textit{Alleluia.} Ps. \textit{When the Lord brought back the captivity of Sion, we became like men comforted.}

\gresetinitiallines{0}
\gregorioscore{antiphon_alleluia_intonation_ps_125}

\begin{latinenglishsection}
\latinenglish{
	2. Tunc replétum est gáudio \textit{os} \textbf{nos}trum:~* et lingua nostra exsul\textit{ta}\textit{ti}\textbf{ó}ne.

	3. Tunc dicent in\textit{ter} \textbf{Gen}tes:~* Magnificávit Dóminus fáce\textit{re} \textit{cum} \textbf{e}is.

	4. Magnificávit Dóminus fácere \textit{no}\textbf{bís}cum:~* facti su\textit{mus} \textit{læ}\textbf{tán}tes.
	
	5. Convérte, Dómine, captivitá\textit{tem} \textbf{nos}tram,~* sicut tor\textit{rens} \textit{in} \textbf{aus}tro.
	
	6. Qui séminant \textit{in} \textbf{lá}crimis,~* in exsultati\textit{ó}\textit{ne} \textbf{me}tent.
	
	7. Eúntes ibant \textit{et} \textbf{fle}bant,~* mitténtes sé\textit{mi}\textit{na} \textbf{su}a.
	
	8. Veniéntes autem vénient cum exsulta\textit{ti}\textbf{ó}ne,~* portántes maní\textit{pu}\textit{los} \textbf{su}os.
	
	9. {\color{red}\textit{(bow)}} Glória Patri, \textit{et} \textbf{Fí}lio,~* et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.
	
	10. {\color{red}\textit{(rise)}} Sicut erat in princípio, et nunc, \textit{et} \textbf{sem}per,~* et in sǽcula sæcu\textit{ló}\textit{rum}. \textbf{A}men.
}{
	2. Then was our mouth filled with gladness; and our tongue with joy.
	
	3. Then shall they say among the Gentiles: The Lord hath done great things for them.
	
	4. The Lord hath done great things for us: we are become joyful.
	
	5. Turn again our captivity, O Lord, as a stream in the south.
	
	6. They that sow in tears shall reap in joy.
	
	7. Going they went and wept, casting their seeds.
	
	8. But coming they shall come with joyfulness, carrying their sheaves.
	
	9. Glory be to the Father, and to the Son, and to the Holy Spirit, as it was in the beginning, is now, and ever shall be, world without end. Amen.
}
\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{alleluia}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{alleluia_dulce_carmen}

\textit{1. Alleluia, song of sweetness, Voice of joy, eternal lay; Alleluia is the anthem of the choirs in Heavenly day, which the angles sing, abiding in the House of God always.}\\
\textit{2. Alleluia thou resoundest, Jerusalem, Mother ever blessed, Alleluias without ending fit beyond place of gladsom rest; exiles wem by Babel's water sit in bondage and distressed.}\\
\textit{3. Alleluia we deserve not here to chant forevermore: Alleluia our transgressions make us for a while give over; for the holy time is coming bidding us our sins deplore.}\\
\textit{4. Trinity of endless glory, hear thy people as they cry; grant us all to keep Thy Easter in our home beyond the sky, there to Thee our Alleluia singing everlastingly. Amen.}

\begin{rubricbox}

{\color{red} The \textit{Alleluia} is now buried (by the servers or any able-bodied men present to aid in such).}

\end{rubricbox}

\vfill\pagebreak

\section*{Blessings}

\begin{rubricbox}

{\color{red}After the \textit{Alleluia} has been buried, the Officiant leads the following blessings (in a \textit{recto tono}):}

\end{rubricbox}

\begin{latinenglishsection}
\latinenglish{
	Allel\'{u}ia, nomen pium, atque joc\'{u}ndum, * dilat\'{e}tur ad laudem Dei in ora \'{o}mnium popul\'{o}rum. {\color{red}\Rbar.} Amen.
	
	Sit in vocibus cred\'{e}ntium clara, * qu\ae in angel\'{o}rum ost\'{e}nditur conc\'{e}ntibus glori\'{o}sa. {\color{red}\Rbar.} Amen.
	
	Et, qu\ae\ in \ae t\'{e}rnis c\'{i}vibus sine son\'{o}rum str\'{e}pitus en\'{i}tet, * in vestris c\'{o}rdibus aff\'{e}ctu plani\'{o}re fruct\'{i}ficet. {\color{red}\Rbar.} Amen.
	
	Angelus D\'{o}mini bonus comit\'{e}tur tecu, Allel\'{u}ia: et \'{o}mnia bona pr\ae p\'{a}ret it\'{i}neri tuo. * Et \'{i}terum cum g\'{a}udio revert\'{a}ris ad nos, Allel\'{u}ia. {\color{red}\Rbar.} Amen.
}{
	May Alleluia, that sacred and joyful word, resound to God’s praise from the lips of all people. {\color{red}\Rbar.} Amen.
	
	May this word, which expresses glory as chanted by the choirs of angels, be sweet as sung by the voices of the believers. {\color{red}\Rbar.} Amen.
	
	And may that which noiselessly gleams in the citizens of heaven, yield fruit in your hearts by ever-growing love. {\color{red}\Rbar.} Amen.
	
	May the Lord’s good angel go with thee, Alleluia; and prepare all good things for thy journey. And again come back to us with joy, Alleluia. {\color{red}\Rbar.} Amen.
}
\end{latinenglishsection}

\section*{Marian Anthem}

\begin{rubricbox}

{\color{red}The Cantor then leads the following:}

\end{rubricbox}


\gresetinitiallines{1}
\gregorioscore{../marian-anthems/ave-regina-caelorum-simple}

\begin{rubricbox}

{\color{red}The Officiant leads the versicle and collect for the Marian Anthem:}

\end{rubricbox}

\begin{latinenglishsection}
\latinenglish{
	{\color{red}\Vbar.} Dign\'{a}re me laud\'{a}re te, Virgo sacr\'{a}ta.
	
	{\color{red}\Rbar.} Da mihi virt\'{u}tem contra hostes tuos.
	
	Or\'{e}mus. Conc\'{e}de, mis\'{e}ricors Deus, fragilit\'{a}ti nostr\ae\ pr\ae s\'{i}dium:  {\color{red}\GreDagger} ut sanct\ae\ Dei Genitr\'{i}cis mem\textbf{\'{o}}\textit{riam} \textbf{\'{a}}gimus; intercessi\'{o}nis ejus aux\'{i}lio, a nostris iniquit\'{a}tibus\\ resurg\'{a}mus. {\color{red}*} Per e\'{u}mdem Christum D\'{o}minum nostrum.
	
	{\color{red}\Rbar.} Amen.
}{
	{\color{red}\Vbar.} Vouchsafe that I may praise thee, O Sacred Virgin.
	
	{\color{red}\Rbar.} Give me strength against thine enemies.
	
	Let us pray. Grant, O merciful God, to our weak natures Thy protection, that we who commemorate the holy Mother of God, may, by the help of her intercession, arise from our iniquities. Through the same Christ our Lord.
	
	{\color{red}\Rbar.} Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The Officiant leads the following in a low recto tono, after which all may depart:}

\end{rubricbox}

\begin{latinenglishsection}
\latinenglish{
	{\color{red}\Vbar.} Div\'{i}num aux\'{i}lium m\'{a}neat semper nob\'{i}scum.
	
	{\color{red}\Rbar.} Amen.
}{
	{\color{red}\Vbar.} May the divine assistance remain always with us.
	
	{\color{red}\Rbar.} Amen.
}
\end{latinenglishsection}

\end{document}