\documentclass{../vespers-sheet}

\begin{document}

\chapter*{The Burial of the Alleluia}

%{\color{red}\textit{A traditional paraliturgical custom in the West involves the burial of a banner with the word ``Alleluia'' on it, usually immediately after Vespers on the evening directly preceding the beginning of Septuagesima, for this is last time that sweet word will be heard, until Easter.}}

\section*{Hymn}

\begin{rubricbox}
{\color{red}All process from the church to the burial site, singing the following hymn is sung:}
\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{alleluia-dulce-carmen}

{\itshape

1. Alleluia, song of sweetness,
    Voice of joy, eternal lay;
    Alleluia is the anthem
    Of the Choirs in Heav’nly day,
    Which the Angels sing, abiding
    In the House of God alway.

2. Alleluia thou resoundest,
    Salem, Mother ever blest;
    Alleluias without ending
    Fit yon place of gladsome rest;
    Exiles we, by Babel’s waters
    Sit in bondage and distress’d.

3. Alleluia we deserve not
    Here to chant forevermore:
    Alleluia our transgressions
    Make us for a while give o’er,
    For the holy time is coming
    Bidding us our sins deplore.

4. Trinity of endless glory,
    Hear Thy people as they cry;
    Grant us all to keep Thy Easter
    In our Home beyond the sky,
    There to Thee our Alleluia
    Singing everlastingly. Amen.
}

\section*{Blessings}

\begin{rubricbox}
{\color{red}At the burial site, the leader chants the following, in a simple \textit{recto tono}.}
\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	Allelúia, nomen pium, atque jucúndum,~* dilatétur ad laudem Dei in ora ómnium populórum. {\color{red}\Rbar.}~Amen.
	
	Sit in vocibus credéntium clara,~* quæ in angelórum osténditur concéntibus gloriósa. {\color{red}\Rbar.}~Amen. %% TODO is concentibus right?
	
	Et, quæ in ætérnis cívibus sine sonórum strépitus enítet,~* in vestris córdibus afféctu planióre fructíficet. {\color{red}\Rbar.}~Amen.
	
	Angelus Dómini bonus comitétur tecum, Allelúia: et ómnia bona præpáret itíneri tuo.~*
	Et íterum cum gáudio revertáris ad nos, Allelúia. {\color{red}\Rbar.}~Amen.
}{
	May Alleluia, that sacred and joyful word, resound to God's praise from the lips of all people. {\color{red}\Rbar.}~Amen.
	
	May this word, which expresses glory as chanted by the choirs of angels, be sweet as sung by the voices of the believers. {\color{red}\Rbar.}~Amen.
	
	And may that which noiselessly gleams in the citizens of heaven, yield fruit in your hearts by ever-growing love. {\color{red}\Rbar.}~Amen.
	
	May the Lord's good angel go with thee, Alleluia; and prepare all good things for thy journey.
	And again come back to us with joy, Alleluia. {\color{red}\Rbar.}~Amen.
}

\end{latinenglishsection}

\section*{Responsory}

\begin{rubricbox}
{\color{red}The \textit{Alleluia} is now buried, while the following responsory is sung:}
\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{resp-media-vita-short}

{\itshape
In the midst of life, we are in death; whom shall we seek to help us, but Thee, o Lord, who for our sins art justly wroth?~* Holy God, holy mighty one, holy and merciful Savior, hand us not over to bitter death.
{\color{red}\Vbar.} In Thee our fathers hoped: they hoped, and Thou didst deliver them.~* Holy God\dots.
}

\begin{rubricbox}
{\color{red}All return to the church in silence.}
\end{rubricbox}

\end{document}