\documentclass{../vespers-sheet}

\begin{document}

%% TODO: correct date
\chapter*{Vespers Insert for Saturday, September 24, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Sixteenth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

%% TODO: correct the little chapter
 \section*{Little Chapter (Rom 11:33)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei, \GreDagger
	 quam incomprehensibília sunt ju\textit{dícia} \textbf{e}jus, et investigábiles viæ ejus!
 	\Rbar.~Deo gratias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

%% TODO: correct the hymn using the antiphonary
\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h)scén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

%% TODO: correct the magnificant antiphon and tone - add pointed text to the intonation

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Then thou shalt call, and the Lord shall answer: thou shalt cry, and He shall say: Here I am.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

%% TODO: proper collect for the date

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Da, qu\'{\ae}sumus, Dómine, pópulo tuo diabólica vitáre contágia: \GreDagger et te solum Deum pura men\textit{te sec}\textbf{tá}ri.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, we beseech thee, grant thy people grace to withstand the temptations of the devil, and with pure hearts to follow thee the only God.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

%% TODO: commemorations based on the date
\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of St. Remigius, Bishop and Confessor (October 1)}

\textit{\textnormal{Ant.} O Priest and Bishop, thou worker of mighty works, good shepherd of the people, pray unto the Lord for us.}
%PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-pont-1v}
% PT: \gregorioscore{../commemorations/conf-pont-1v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ induit eum.  %PT: allelúia
	
	Orémus.
	Da, qu\'{\ae}sumus, omnípotens Deus: ut beáti Remígii Confessóris tui atque Pontíficis veneránda solémnitas, et devotiónem nobis áugeat, et salútem.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	The Lord loved him and adorned him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	Grant, we beseech Thee, O almighty God, that the august solemnity of Thy blessed confessor and bishop, Remigius, may increase our devotion and promote our salvation. 
	Through Jesus Christ our Lord, who lives and reigns with you and the Holy Spirit, God, for ever and ever. 
	\Rbar.~Amen.
}

\end{latinenglishsection}

\pagebreak

\section*{Commemoration of the Holy Guardian Angels (October 2)}

\textit{\textnormal{Ant.}  Are they not all ministering spirits, sent forth to minister for them who shall be heirs of salvation?}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--omnes_sunt_administratorii--solesmes_1934}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~In conspéctu Angelórum psallam tibi, Deus meus.\\
	\Rbar.~Adorábo ad templum sanctum tuum, et confitébor nómini tuo.
	
	Orémus.
	Deus, qui ineffábili providéntia sanctos Angelos tuos ad nostram custódiam míttere dignáris: largíre supplícibus tuis; et eórum semper protectióne deféndi, et ætérna societáte gaudére.
	Per eúndem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	\Vbar.~Before the Angels will I sing praise unto thee, O my God.
	\Rbar.~I will worship toward thy holy Temple, and praise thy Name.
	
	O God, Who in thine unspeakable Providence hast been pleased to give thine holy Angels charge over us, to keep us, mercifully grant unto our prayers, that we be both ever fenced by their wardship here, and everlastingly blessed by their fellowship hereafter.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

%% TODO: benedicamus in line with the season

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

%% TODO: proper Marian anthem for the seasons

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}