\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, May 14, 2022}

 \textit{Vespers today is based on the following Sunday, which is the fourth after Easter.
 All is said as in the Eastertide booklet until the Little Chapter.}

\section*{Little Chapter (James 1:17)}

\begin{latinenglishsection}

 \latinenglish{
	 Caríssimi: Omne datum óptimum, et omne donum perféctum desúrsum est, descéndens a Patre lúminum,~\GreDagger\
	 apud quem non est transmutátio,~*
	 nec vicissitúdinis obumbrátio.
 	\Rbar.~Deo gratias.
 }{
 	Beloved: Every best gift, and every perfect gift, is from above, coming down from the Father of lights,
 	with whom there is no change, nor shadow of alteration.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\begin{rubricbox}

The hymn \textit{Ad cenam agni providi} and versicle \textit{\Rbar.~Mane nobiscum\dots} are said as given in the booklet for Eastertide.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../hymns/ad-coenam-agni}

{\itshape
	1. The Lamb's high banquet we await
	in snow-white robes of royal state:
	and now, the Red Sea's channel past,
	to Christ our Prince we sing at last. 
	
	2. Upon the Altar of the Cross
	His Body hath redeemed our loss:
	and tasting of his roseate Blood,
	our life is hid with Him in God. 
	
	3. That Paschal Eve God's arm was bared,
	the devastating Angel spared:
	by strength of hand our hosts went free
	from Pharaoh's ruthless tyranny. 
	
	4. Now Christ, our Paschal Lamb, is slain,
	the Lamb of God that knows no stain,
	the true Oblation offered here,
	our own unleavened Bread sincere. 
	
	5. O Thou, from whom hell's monarch flies,
	O great, O very Sacrifice,
	Thy captive people are set free,
	and endless life restored in Thee.
	
	6. For Christ, arising from the dead,
	from conquered hell victorious sped,
	and thrust the tyrant down to chains,
	and Paradise for man regains. 
	
	7. We pray Thee, King with glory decked,
	in this our Paschal joy, protect
	from all that death would fain effect
	Thy ransomed flock, Thine own elect.
	
	8. To Thee who, dead, again dost live,
	all glory Lord, Thy people give;
	all glory, as is ever meet,
	to Father and to Paraclete. Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ma(h)ne(h) no(h)bís(h)cum(h) Dó(h)mi(h)ne,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Quó(h)ni(h)am(h) ad(h)ves(h)pe(h)rá(h)scit,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Stay with us, Lord, alleluia.
\Rbar.~For it is getting towards evening, alleluia.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} I am going to Him Who sent Me, and none of you ask Me: Where art Thou going? alleluia, alleluia.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui fidélium mentes uníus éfficis voluntátis:
	da pópulis tuis id amáre quod prǽcipis, id desideráre quod promíttis;
	ut inter mundánas varietátes ibi nostra fixa sint corda, ubi vera sunt gáudia.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who makest the minds of the faithful to be of one will, grant to Thy people to love that which Thou commandest
	and desire that which Thou dost promise; that so, among the changing things of this world,
	our hearts may be set where true joys are to be found.
	Through our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Jean-Baptiste de La Salle (May 15)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum, allelúia.\\ 
	\Rbar.~Stolam glóriæ induit eum, allelúia.
	
	Orémus.
Deus, qui, ad christiánam páuperum eruditiónem et ad juvéntam in via veritátis firmándam, sanctum Joánnem Baptístam Confessórem excitásti, et novam per eum in Ecclésia famíliam collegísti: concéde propítius; ut ejus intercessióne et exémplo, stúdio glóriæ tuæ in animárum salúte fervéntes, ejus in cælis corónæ partícipes fíeri valeámus.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	The Lord loved him and adorned him, alleluia
	\Rbar.~He clothed him with a robe of glory, alleluia
	
	Let us pray.
O God, who didst raise up the Confessor, St. John Baptist, to promote the Christian education of the poor and to strengthen the young in the way of truth, and, through him, didst gather together a new family within thy Church, mercifully grant through his intercession and example, that we may burn with zeal for thy glory, through the salvation of souls, and may share his crown in heaven.
Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain with us always. \Rbar.~Amen.
}

\end{latinenglishsection}

\end{document}