\documentclass{../../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Friday, September 2, 2022}

 \textit{Vespers today is that of a Ferial Friday, of semidouble rank.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Sirach 44:16-17)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
 	Ecce sacérdos magnus, {\color{red}\GreDagger}\ qui in diébus suis plá\textit{cuit} \textbf{De}o, et invéntus est justus: * et in témpore iracúndiæ factus est reconciliátio.
 	{\color{red}\Rbar.}~Deo gratias.
 }{
	Behold an high priest, who in his days pleased God, and was found righteous, and in the time of wrath he was made a propitiation.
 	{\color{red}\Rbar.}~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../../hymns/iste_confessor}

%\vfill
%\pagebreak

{\itshape

	1. This the confessor of the Lord, whose triumph
	Now all the faithful celebrate, with gladness
	On this feast day year by year receiveth
	Merited honors.
	
	2. Saintly and prudent, modest in behaviour,
	Peaceful and sober, chaste was he, and lowly.
	While that life’s vigour, coursing through his members,
	Quickened his being.
	
	3. Sick ones of old time, to his tomb resorting,
	Sorely by ailments manifold afflicted,
	Oft-times have welcomed health and strength returning,
	At his petition.
	
	4. Whence we in chorus gladly do him honour,
	Chanting his praises with devout affection,
	That in his merits we may have a portion,
	Now and forever.
	
	5. His be the glory, power and salvation,
	Who over all things reigneth in the highest,
	Earth’s mighty fabric ruling and directing,
	Onely and Trinal.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>V/</sp>.</c>  A(h)má(h)vit(h) e(h)um(h) Dó(h)min(h)us(h), et(h) or(h)ná(h)vit(h) e(h)um.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>R/</sp>.</c> Sto(h)lam(h) gló(h)ri(h)æ(h) ín(h)du(h)it(h) e(h)um.(g'_/hvGF'E/fgf.) (::)
}

\textit{{\color{red}\Vbar.}~The Lord loved him and beautified him.
{\color{red}\Rbar.}~He clothed him with a robe of glory.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} O thou Priest and Bishop, * thou worker of mighty works, thou good shepherd over God’s people, pray for us unto the Lord.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui ad tuéndam cathólicam fidem, et univérsa in Christo instauránda sanctum Pium, Summum Pontíficem, cælésti sapiéntia et apostólica fortitúdine replevísti: {\color{red}\GreDagger} concéde propítius; ut, ejus institúta et exém\textit{pla sec}\textbf{tán}tes, pr\'{\ae}mia consequámur ætérna. *
	Per eúmdem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	{\color{red}\Rbar.}~Amen.
}{
	Lord, hear my prayer.
	{\color{red}\Rbar.}~And let my cry come unto Thee.
	
	Let us pray.
	O God, who to safeguard Catholic faith and to restore all things in Christ, didst fill the Supreme Pontiff, Saint Pius, with heavenly wisdom and apostolic fortitude: grant in thy mercy: that by striving to fulfill his ordinances and to follow his example, we may reap eternal rewards.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	{\color{red}\Rbar.}~Amen.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of St. Stephen, King of Hungary, Confessor (September 2)}

\textit{\textnormal{Ant.} This servant of God, despising the world, and triumphing over earthly things, hath laid up treasures in heaven by word and deed.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../../commemorations/conf-non-pont-2v}
% PT: \gregorioscore{../commemorations/conf-non-pont-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Justum dedúxit Dóminus per vias rectas.\\ %PT: allelúia
	{\color{red}\Rbar.}~Et osténdit illi regnum Dei.  %PT: allelúia
	
	Orémus.
	Concéde, qu\'{\ae}sumus, Ecclésiæ tuæ, omnípotens Deus: {\color{red}\GreDagger}\ ut beátum Stéphanum Confessórem tuum, quem regnántem in terris propagató\textit{rem há}\textbf{bu}it,\\ propugnatórem habére mereátur gloriósum in cælis. *
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	{\color{red}\Vbar.}~The Lord led the just in right paths. %PT: alleluia
	{\color{red}\Rbar.}~And showed him the kingdom of God. %PT: alleluia
	
	Let us pray.
	Grant unto thy Church, we beseech thee, O Almighty God, that even as thy blessed Confessor Stephen, while he was a King upon earth, was her forwarder, so, now that he is a glorious Saint in heaven, he may be her defender.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
}{
	{\color{red}\Vbar.}~Lord, hear my prayer. {\color{red}\Rbar.}~And let my cry come unto Thee.
	{\color{red}\Vbar.}~Let us bless the Lord. {\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	{\color{red}\Rbar.} Amen.
}{
	{\color{red}\Vbar.} May the souls of the faithful departed, through the mercy of God, rest in peace. {\color{red}\Rbar.}~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	{\color{red}\Vbar.} Dóminus det nobis suam pacem. \\
	{\color{red}\Rbar.} Et vitam ætérnam. Amen.
}{
	{\color{red}\Vbar.} May the Lord grant us his peace. {\color{red}\Rbar.}~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Divínum auxílium máneat semper nobíscum.\\
	{\color{red}\Rbar.}~Amen.
}{
	May the divine assistance remain always with us.
	{\color{red}\Rbar.}~Amen.
}
\end{latinenglishsection}

\end{document}