\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, March 5, 2022}

 \textit{Vespers today is based on the following Sunday, which is the First Sunday of Lent.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (2 Cor. 6:1-2)}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Exhortámur vos, ne in vácuum grátiam Dei recipiátis.~\GreDagger\
	 Ait enim: Témpore accépto exaudívi te,~*
	 et in die salútis adjúvi te.
 	\Rbar.~Deo gratias.
 }{
	Brethren: we exhort you, that you receive not the grace of God in vain.
 	For he saith: In an acceptable time have I heard thee; and in the day of salvation have I helped thee.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/audi-benigne-conditor}

\vfill
\pagebreak

{\itshape

1. O merciful Creator, hear!
To us in pity bow Thine ear:
accept the tearful prayer we raise
in this our fast of forty days.

2. Our hearts are open, Lord, to Thee:
Thou knowest our infirmity;
pour out on all who seek Thy face
abundance of Thy pardoning grace.

3. Our sins are many, this we know;
spare us, good Lord, Thy mercy show;
and for the honor of Thy name
our fainting souls to life reclaim.

4. Give us self-control that springs
from discipline of outward things,
that fasting inward secretly
the soul may purely dwell with Thee.

5. We pray Thee, Holy Trinity,
one God, unchanging Unity,
that we from this our abstinence
may reap the fruits of penitence. Amen.

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. An(h)ge(h)lis(h) su(h)is(h) De(h)us(h) man(h)dá(h)vit(h) de(h) te.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Ut(h) cus(h)tó(h)di(h)ant(h) te(h) in(h) óm(h)ni(h)bus(h) vi(h)is tu(h)is.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~He hath given his angels charge over thee.
\Rbar.~To keep thee in all thy ways.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Then thou shalt call, and the Lord shall answer: thou shalt cry, and He shall say: Here I am.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui Ecclésiam tuam ánnua quadragesimáli observatióne puríficas:
	præsta famíliæ tuæ; ut, quod a te obtinére abstinéndo nítitur, hoc bonis opéribus exsequátur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who purifiest Thy Church by the yearly Lenten observance,
	grant to Thy household that what they strive to obtain from Thee by abstinence, they may achieve by good works.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of Saints Perpetua and Felicity, Martyrs (March 6)}

\textit{\textnormal{Ant.} Theirs is the kingdom of heaven, who despising a worldy life have attained the rewards of the kingdom,
and have washed their robes in the Blood of the Lamb.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Lætámini in Dómino, et exsultáte justi.\\
	\Rbar.~Et gloriámini omnes recti corde.
	
	Orémus.
	Da nobis, quǽsumus, Dómine, Deus noster, sanctárum Mártyrum tuárum Perpétuæ et Felicitátis palmas incessábili devotióne venerári: ut, quas digna mente non póssumus celebráre, humílibus saltem frequentémus obséquiis.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	\Vbar.~Be glad in the Lord and rejoice, ye just.
	\Rbar.~And exult, all ye upright of heart.
	
	Let us pray.
	Grant us we pray, O Lord our God, to reverence with unceasing devotion the triumph of thy holy martyrs Perpetua and Felicitas; and although we cannot pay the honor what is their due, let us at least present them our humble service.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}