\documentclass{../../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Friday, November 4, 2022}

 \textit{Vespers today is that of a Ferial Friday, of semidouble rank.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Sirach 44:16-17)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
 	Ecce sacérdos magnus, qui in diébus suis plácuit Deo, {\color{red}\GreDagger}\ et invén\textit{tus est} \textbf{ju}stus: et in témpore iracúndiæ factus est reconciliátio.
 	{\color{red}\Rbar.}~Deo gratias.
 }{
	Behold an high priest, who in his days pleased God, and was found righteous, and in the time of wrath he was made a propitiation.
 	{\color{red}\Rbar.}~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../../hymns/iste_confessor}

%\vfill
%\pagebreak

{\itshape

	1. This the confessor of the Lord, whose triumph
	Now all the faithful celebrate, with gladness
	On this feast day year by year receiveth
	Merited honors.
	
	2. Saintly and prudent, modest in behaviour,
	Peaceful and sober, chaste was he, and lowly.
	While that life’s vigour, coursing through his members,
	Quickened his being.
	
	3. Sick ones of old time, to his tomb resorting,
	Sorely by ailments manifold afflicted,
	Oft-times have welcomed health and strength returning,
	At his petition.
	
	4. Whence we in chorus gladly do him honour,
	Chanting his praises with devout affection,
	That in his merits we may have a portion,
	Now and forever.
	
	5. His be the glory, power and salvation,
	Who over all things reigneth in the highest,
	Earth’s mighty fabric ruling and directing,
	Onely and Trinal.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>V/</sp>.</c>  Jus(h)tum(h) de(h)dú(h)xit(h) Dó(h)mi(h)nus(h) per(h) vi(h)as(h) rec(h)tas.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>R/</sp>.</c> Et(h) os(h)tén(h)dit(h) il(h)li(h) reg(h)num(h) De(h)i.(g'_/hvGF'E/fgf.) (::)
}

\textit{{\color{red}\Vbar.}~The Lord guided the just in right paths.
{\color{red}\Rbar.}~And showed him the kingdom of God.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} The Lord loved him * and beautified him. He clothed him with a robe of glory, and crowned him at the gates of Paradise.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
	
	Orémus.
	Ecclésiam tuam, Dómine, sancti Cároli Confessóris tui atque Pontíficis contínua protectióne custódi: ut, sicut illum pastorális sollicitúdo gloriósum réddidit; ita nos ejus intercéssio in tuo semper fáciat amóre fervéntes.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	{\color{red}\Rbar.}~Amen.
}{
	Lord, hear my prayer.
	{\color{red}\Rbar.}~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, give unto thy Church for an unsleeping warder thine holy Confessor Bishop Charles; upon earth his carefulness did make him glorious as a shepherd, there where he is may his prayerfulness make him effectual as a bedesman, pleading with thee to make us to love thee more.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	{\color{red}\Rbar.}~Amen.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of the Fifth Day of the Octave of All Saints}

\textit{\textnormal{Ant.} O ye Angels, ye Archangels, ye Thrones and Dominions, ye Principalities and Powers, ye mighty ones of the heavens, ye Cherubim and Seraphim, O ye Patriarchs and Prophets, ye holy Teachers of the Law, O ye Apostles, O all ye Martyrs of Christ, ye holy Confessors, ye Virgins of the Lord, ye Hermits, O all ye holy children of God, make intercession for us.}

\gresetinitiallines{1}
\gregorioscore{../../commemorations/an--angeli_archangeli_caelorum--solesmes_1934}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}.~Lætámini in Dómino, et exsultáte justi.\\
	{\color{red}\Rbar.}.~Et gloriámini omnes recti corde.
	
	Orémus.
	Omnípotens sempitérne Deus, qui nos ómnium Sanctórum tuórum mérita sub una tribuísti celebritáte venerári: qu\'{\ae}sumus; ut desiderátam nobis tuæ propitiatiónis abundántiam, multiplicátis\\ intercessóribus, largiáris.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	{\color{red}\Vbar.}.~Be glad in the Lord, and rejoice, ye righteous.
	{\color{red}\Rbar.}.~And shout for joy, all ye that are upright in heart.
	
	Let us pray.
	Almighty, eternal God, Who granted us to honor the merits of all Thy Saints in a single solemn festival, bestow on us, we beseech Thee, through their manifold intercession, that abundance of Thy mercy for which we yearn.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
}{
	{\color{red}\Vbar.}~Lord, hear my prayer. {\color{red}\Rbar.}~And let my cry come unto Thee.
	{\color{red}\Vbar.}~Let us bless the Lord. {\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	{\color{red}\Rbar.} Amen.
}{
	{\color{red}\Vbar.} May the souls of the faithful departed, through the mercy of God, rest in peace. {\color{red}\Rbar.}~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	{\color{red}\Vbar.} Dóminus det nobis suam pacem. \\
	{\color{red}\Rbar.} Et vitam ætérnam. Amen.
}{
	{\color{red}\Vbar.} May the Lord grant us his peace. {\color{red}\Rbar.}~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Divínum auxílium máneat semper nobíscum.\\
	{\color{red}\Rbar.}~Amen.
}{
	May the divine assistance remain always with us.
	{\color{red}\Rbar.}~Amen.
}
\end{latinenglishsection}

\end{document}