\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of the Holy Family}

\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} And Jacob begot Joseph the husband of Mary of whom was born Jesus, who is called the Christ.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. Donec ponam ini\textbf{mí}cos \textbf{tu}os,~*
	scabéllum pe\textit{dum} \textit{tu}\textbf{ó}rum.

3. Virgam virtútis tuæ emíttet Dómi\textbf{nus} ex \textbf{Si}on:~*
	domináre in médio inimicó\textit{rum} \textit{tu}\textbf{ó}rum.

4. Tecum princípium in die virtútis tuæ in splendóri\textbf{bus} sanc\textbf{tó}rum:~*
	ex útero ante lucíferum \textit{gé}\textit{nu}\textbf{i} te.

5. Jurávit Dóminus, et non p{\oe}ni\textbf{té}\-bit \textbf{e}um:~*
	Tu es sacérdos in ætérnum secúndum órdi\textit{nem} \textit{Mel}\textbf{chí}se\-dech.

6. Dóminus a \textbf{dex}tris \textbf{tu}is,~*
	confrégit in die iræ \textit{su}\textit{æ} \textbf{re}ges.

7. Judicábit in natiónibus, im\textbf{plé}bit ru\textbf{í}nas:~*
	conquassábit cápita in ter\textit{ra} \textit{mul}\textbf{tó}rum.

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}


8. De torrénte in \textbf{vi}a \textbf{bi}bet:~*
	proptérea exal\textit{tá}\textit{bit} \textbf{ca}put.

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~*
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Jacob autem\dots}, then all \textbf{stand}.

\end{rubricbox}


\section*{Psalm 112}

\textit{\textnormal{Ant. 2.} The angel of the Lord appeared in a dream to Joseph, saying: Joseph, son of David, fear not to take unto thee Mary thy wife; for that which is born of her is of the Holy Ghost.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}


\gresetinitiallines{1}
\gregorioscore{ps112-antiphon-only}

\vfill\pagebreak

\gresetinitiallines{0}
\gregorioscore{ps112-psalm-intonation}

 \begin{latinenglishsection}

\latinenglish{
	
 2. \textit{(bow)} Sit nomen Dómini \textbf{be}ne\-\textbf{díc}tum,~*
	ex hoc nunc, et \textbf{us}que in \textbf{s\'{\ae}}culum.

 3. A solis ortu usque \textbf{ad} oc\textbf{cá}sum,~*
	laudábile \textbf{no}men \textbf{Dó}mini.

 4. Excélsus super omnes \textbf{gen}tes \textbf{Dó}\-minus,~*
	et super cælos \textbf{gló}ria \textbf{e}jus.

 5. Quis sicut Dóminus, Deus noster, qui in \textbf{al}tis \textbf{há}bitat,~*
	et humília réspicit in cælo \textbf{et} in \textbf{ter}ra?

 6. Súscitans a \textbf{ter}ra \textbf{ín}opem,~*
	et de stércore \textbf{é}rigens \textbf{páu}perem:

 7. Ut cóllocet eum \textbf{cum} prin\textbf{cí}pi\-bus,~*
	cum princípibus \textbf{pó}puli \textbf{su}i.

 8. Qui habitáre facit stéri\textbf{lem} in \textbf{do}\-mo,~*
	matrem fili\textbf{ó}rum læ\textbf{tán}tem.
	
\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~*
	et Spi\textbf{rí}tui \textbf{Sanc}to.

}{
 	\input{../psalms/english/ps112}
}

\end{latinenglishsection}


\textit{(rise)}  Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textbf{ló}rum. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Angelus Domini\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 121}

\textit{\textnormal{Ant. 3.} The shepherds came with haste, and found Mary and Joseph, and the Child laid in a manger.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps121-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. Stantes erant \textbf{pe}des \textbf{nos}tri,~* in átriis \textbf{tu}is, Je\textbf{rú}salem.

3. Jerúsalem, quæ ædifi\textbf{cá}tur ut \textbf{cí}vitas:~* cujus participátio ejus \textbf{in} id\textbf{íp}sum.

4. Illuc enim ascendérunt tribus, \textbf{tri}bus \textbf{Dó}mini:~* testimónium Israël ad confiténdum \textbf{nó}mini \textbf{Dó}mini.

5. Quia illic sedérunt sedes \textbf{in} ju\textbf{dí}cio,~* sedes super \textbf{do}mum \textbf{Da}vid.

6. Rogáte quæ ad pacem \textbf{sunt} Je\textbf{rú}salem:~* et abundántia dili\textbf{gén}ti\textbf{bus} te:

7. Fiat pax in vir\textbf{tú}te \textbf{tu}a:~* et abundántia in \textbf{túr}ribus \textbf{tu}is.

8. Propter fratres meos, et \textbf{pró}ximos \textbf{me}os,~* loquébar \textbf{pa}cem \textbf{de} te:

9. Propter domum Dómini, \textbf{De}i \textbf{nos}tri,~* quæsívi \textbf{bo}na \textbf{ti}bi.
	
}{
	\input{../psalms/english/ps121}
}

\end{latinenglishsection}

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~* et Spi\textbf{rí}tui \textbf{Sanc}to.

\textit{(rise)}  Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~* et in s\'{\ae}cula sæcu\textbf{ló}rum. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Pastores\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 126}

\textit{\textnormal{Ant. 4.} The Magi entering the house found the Child with Mary His Mother.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps126-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. Nisi Dóminus custodíerit civi\textbf{tá}tem,~* frustra vígilat qui custó\textit{dit} \textbf{e}am.

3. Vanum est vobis ante lucem \textbf{súr}gere:~* súrgite postquam sedéritis, qui manducátis panem \textit{do}\textbf{ló}ris.

4. Cum déderit diléctis suis \textbf{som}\-num:~* ecce heréditas Dómini fílii: merces, fruc\textit{tus} \textbf{ven}tris.

5. Sicut sagíttæ in manu pot\textbf{én}tis:~* ita fílii ex\textit{cus}\textbf{só}rum.

6. Beátus vir qui implévit desidérium suum ex \textbf{ip}sis:~* non confundétur cum loquétur inimícis suis \textit{in} \textbf{por}ta.

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~* et Spirítu\textit{i} \textbf{Sanc}to.
	
}{
	\input{../psalms/english/ps126}
}

\end{latinenglishsection}

\textit{(rise)}  Sicut erat in princípio, et nunc, et \textbf{sem}per,~* et in s\'{\ae}cula sæculó\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Magi\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 147}

\textit{\textnormal{Ant. 5.} And His father and his mother marvelled at what was said about Him.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps147-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. Quóniam confortávit seras por\textbf{tá}\-rum tu\textbf{á}rum:~*
	benedíxit fíliis \textit{tu}\textit{is} \textbf{in} te.

3. Qui pósuit fines \textbf{tu}os \textbf{pa}cem:~*
	et ádipe fruménti \textit{sá}\textit{ti}\textbf{at} te.

4. Qui emíttit elóquium \textbf{su}um \textbf{ter}ræ:~*
	velóciter currit \textit{ser}\textit{mo} \textbf{e}jus.

5. Qui dat nivem \textbf{sic}ut \textbf{la}nam:~*
	nébulam sicut cí\textit{ne}\textit{rem} \textbf{spar}git.

6. Mittit crystállum suam \textbf{sic}ut buc\textbf{cél}las:~*
	ante fáciem frígoris ejus quis \textit{sus}\textit{ti}\textbf{né}bit?

7. Emíttet verbum suum, et lique\textbf{fá}ciet \textbf{e}a:~*
	flabit spíritus ejus, et \textit{flu}\textit{ent} \textbf{a}quæ.

8. Qui annúntiat verbum \textbf{su}um \textbf{Ja}cob:~*
	justítias, et judícia \textit{su}\textit{a} \textbf{Is}raël.
	
}{
2. Because he hath strengthened the bolts of thy gates, he hath blessed thy children within thee.

3. Who hath placed peace in thy borders: and filleth thee with the fat of corn.

4. Who sendeth forth his speech to the earth: his word runneth swiftly.

5. Who giveth snow like wool: scattereth mists like ashes.

6. He sendeth his crystal like morsels: who shall stand before the face of his cold?

7. He shall send out his word, and shall melt them: his wind shall blow, and the waters shall run.

8. Who declareth his word to Jacob: his justices and his judgments to Israel.

9. He hath not done in like manner to every nation: and his judgments he hath not made manifest to them.

%Glory be.
}

\end{latinenglishsection}

\vfill\pagebreak

9. Non fecit táliter omni \textbf{na}ti\textbf{ó}ni:~* et judícia sua non manifes\textit{tá}\textit{vit} \textbf{e}is.
	
%\gresetinitiallines{0}
%\gabcsnippet{
%(c4) 9. Non(h) fe(h)cit(h) tá(h)li(h)ter(h) om(h)ni(h) <b>na</b>(ixi)ti(h)<b>ó</b>(g)ni:(h.) *(:) et(h) ju(h)dí(h)ci(h)a(h) su(h)a(h) non(h) ma(h)ni(h)fes(h)<i>tá</i>(g)<i>vit</i>(f) <b>e</b>(gh)is.(gf..) (::)
%}

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~*
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Erat Pater ejus\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Luke 2:51)}

\begin{latinenglishsection}

\latinenglish{
	
	Descéndit Jesus cum María et Joseph, et venit Názareth,~*
	et erat súbditus illis.
	\Rbar.~Deo grátias.
}{
	 And Jesus went down with them, and came to Nazareth, and was subject to them.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/o-lux-beata-caelitum}

{\itshape

1. O highest Hope of mortals,
blest Light of Saints above,
O Jesu, on whose boyhood
home smiled with kindly love; 

2. And thou whose bosom nursed Him,
O Mary, highly graced,
whose breast gave milk to Jesus,
whose arms thy God embraced. 

3. And thou of all men chosen
to guard the Virgin's fame,
to whom God's Son refused not
a Father's gracious name. 

4. Born for nation's healing,
of Jesse's lineage high,
behold the supplicants kneeling,
o hear the sinner's cry! 

5. The sun returned to evening,
dusks all the twilight air:
we, lingering here before you,
pour out our heartfelt prayer. 

6. Your home was as a garden
made glad with fairest flowers;
may life thus blossom sweetly
in every home of ours. 

7. All praise to Thee, O Jesus,
who parents dost obey;
praise to the sovereign Father
and Paraclete for aye. 
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Be(h)á(h)ti(h) qui(h) há(h)bi(h)tant(h) in(h) do(h)mo(h) tu(h)a,(h) Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. In(h) s<sp>'ae</sp>(h)cu(h)la(h) sæ(h)cu(h)ló(h)rum(h) lau(h)dá(h)bunt(h) te.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Blessed are they who dwell in Thy house, O Lord.
\Rbar.~They will praise Thee for ever and ever.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} The Word was made flesh, and dwelt among us, full of grace and truth; of Whose fullness we have all received, and grace for grace, alleluia.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
3. Quia respéxit humilitátem ancíllæ \textbf{su}æ:~*
	ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.

4. Quia fecit mihi magna qui \textbf{pot}ens est:~* 
	et sanctum \textit{no}\textit{men} \textbf{e}jus.

5. Et misericórdia ejus a progénie in pro\textbf{gé}nies~*
	timén\textit{ti}\textit{bus} \textbf{e}um.

6. Fecit poténtiam in bráchio \textbf{su}o:~*
	dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.

7. Depósuit poténtes de \textbf{se}de,~*
	et exal\textit{tá}\textit{vit} \textbf{hú}miles.

8. Esuriéntes implévit \textbf{bo}nis:~*
	et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.

9. Suscépit Israël púerum \textbf{su}um,~*
	recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.

10. Sicut locútus est ad patres \textbf{nos}tros,~*
	Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~* 
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon: \textit{Verbum caro factum est\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\vfill

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Dómine Jesu Christe, qui Maríæ et Joseph súbditus, domésticam vitam ineffabílibus virtútibus consecrásti:
	fac nos, utriúsque auxílio, Famíliæ sanctæ tuæ exémplis ínstrui;
	et consórtium cónsequi sempitérnum.
	Qui vivis.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord Jesus Christ, Who wast subject to Mary and Joseph, and didst consecrate family life with Thine unspeakable virtues;
	grant us, with the aid of them both, that we may be taught by the examples of Thy Holy family,
	and may obtain eternal fellowship therewith.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\section*{Commemoration of the Epiphany}

%% TODO: this is different every year
\textit{\textnormal{Ant.} O Christ, Thou Light of light hast appeared, unto Whom the Magi offer gifts, alleluia, alleluia, alleluia.}

\gresetinitiallines{1}
\gregorioscore{comm-oct-jan-8}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Reges Tharsis et ínsulæ múnera ófferent.\\
	\Rbar.~Reges Arabum et Saba dona addúcent.\\ 
	
	Orémus.
	Deus, qui hodiérna die Unigénitum tuum géntibus stella duce revelásti:
	concéde propítius; ut qui jam te ex fide cognóvimus,
	usque ad contemplándam spéciem tuæ celsitúdinis perducámur.
}{
	The kings of Tharsis and the isles shall offer presents.
	\Rbar.~The kings of Arabia and Saba shall bring gifts.
	
	Let us pray.
	O God, Who on this day through the guidance of a star didst manifest Thine only-begotten Son to the Gentiles;
	mercifully grant that we who know Thee now by faith, may one day be brought to the contemplation of the beauty of Thy majesty.
}
\end{latinenglishsection}

\section*{Commemoration of the Sunday}

\textit{\textnormal{Ant.} The Child Jesus remained in Jerusalem; and His parents knew it not, thinking Him to be among the caravan. And they sought Him among their relatives and acquaintances.}

\gresetinitiallines{1}
\gregorioscore{comm-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Omnes de Saba vénient, allelúia.\\
	\Rbar.~Aurum et thus deferéntes, allelúia.\\
	
	Orémus.
	Vota, qu\'{\ae}sumus Dómine, supplicántis pópuli cælésti pietáte proséquere:
	ut et quæ agénda sunt, vídeant, et ad implénda quae víderint, convaléscant.
	Per Dóminum.
}{
	All they from Saba shall come, alleluia.
	\Rbar.~Bringing gold and frankincense, alleluia.
	
	Let us pray.
	We beseech Thee, O Lord, mercifully hear the prayers of Thy people who call upon Thee;
	grant that they may recognize their duites, and may receive the grace to fulfill them faithfully.
	Through Our Lord.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-double}

\vfill\pagebreak

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

%\begin{rubricbox}
%
%The Marian anthem \textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).
%
%\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/alma-redemptoris-mater-simple}

{\itshape
	Mother of Christ, hear thou thy people's cry,
	Star of the deep and Portal of the sky!
	Mother of Him who thee from nothing made.
	Sinking we strive and call to thee for aid:
	oh, by what joy which Gabriel brought to thee,
	thou Virgin first and last, let us thy mercy see. 
}

\vfill\pagebreak

\begin{latinenglishsection}
\latinenglish{	
	\Vbar. Post partum, Virgo, invioláta permansísti.\\
	\Rbar. Dei Génitrix, intercéde pro nobis.
	
	Orémus.
	Deus qui salútis ætérnæ, beátæ Maríæ virginitáte fecúnda, humáno géneri pr\'{\ae}mia præstitísti:~\GreDagger\
	tríbue, qu\'{\ae}sumus; ut ipsam pro nobis intercédere sentiámus,~*
	per quam merúimus auctórem vitæ suscípere, Dóminum nostrum Jesum Christum Fílium tuum.
	\Rbar.~Amen.
}{	
	After child-birth thou didst remain a pure virgin.
	\Rbar.~Intercede for us, O Mother of God.
	
	Let us pray.
	O God, who, by the fruitful virginity of the blessed Mary, hast given to mankind the rewards of eternal  salvation;
	grant, we beseech thee, that we may experience her intercession for us, through whom we have received the author of life,
	Thy Son Jesus Christ, our Lord.
	\Rbar.~Amen.
}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}