\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, 29 Jan 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fourth after Epiphany. All is said as in the main booklet until the Magnificat.}

\section*{Magnificat}

%% all the Sundays after Epiphany share the same antiphon
\gresetinitiallines{1}
\gregorioscore{../post-epiph-2/magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{../post-epiph-2/magnificat-antiphon}

\textit{\textnormal{Ant.} God hath received Israel His servant according as He spoke to Abraham and to his seed, and hath exalted the humble forever.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui nos in tantis perículis constitútos, pro humána scis fragilitáte non posse subsístere:
	da nobis salútem mentis et córporis; ut ea quæ pro peccátis nostris pátimur, te adjuvánte vincámus.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, who knowest us to be set in the midst of dangers so great that, by reason of the frailty of our nature we cannot always withstand;
	grant to us health of mind and body, that being helped by Thee, we may overcome the things which we suffer for our sins.
	Through our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Francis de Sales, Bishop, Confessor, and Doctor (Jan 29)}

\textit{\textnormal{Ant.} O blest teacher, light of holy Church, blessed Francis, thou lover of God's law, plead with the Son of God for us.}

\gresetinitiallines{1}
\gregorioscore{comm-francis-de-sales}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ %PT: allelúia
	\Rbar.~Et osténdit illi regnum Dei.  %PT: allelúia
	
	Orémus.
	Deus, qui ad animárum salútem beátum Francíscum Confessórem tuum atque Pontíficem ómnibus ómnia factum esse voluísti:
	concéde propítius; ut caritátis tuæ dulcédine perfúsi,
	ejus dirigéntibus mónitis ac suffragántibus méritis,
	ætérna gáudia consequámur.
}{
	The Lord led the just in right paths. %PT: alleluia
	\Rbar.~And showed him the kingdom of God. %PT: alleluia
	
	Let us pray.
	O God, by whose gracious will blessed Francis, Thy confessor and bishop, became all things to all men, for the salvation of their souls;
	grant, we beseech Thee that being filled with the sweetness of Thy leove,
	we may, by the guidance of his counsels and by the aid of his merits,
	attain unto the joys of everlasting life.
}

\end{latinenglishsection}

\section*{Commemoration of St. Martina, Virgin and Martyr (Jan 30)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua.\\ %PT: allelúia
	\Rbar.~Inténde, próspere procéde, et regna.  %PT: allelúia
	
	Orémus.
	Deus, qui inter cétera poténtiæ tuæ mirácula, étiam in sexu frágili victóriam martýrii contulísti:
	concéde propítius; ut qui beátæ Martínæ Vírginis et Mártyris tuæ natalítia cólimus,
	per ejus ad te exémpla gradiámur.
	Per Dóminum.
}{
	In thy glory and thy splendor. %PT: alleluia
	\Rbar.~Go forth, advance with victory and reign. %PT: alleluia
	
	Let us pray.
	O God, who among the wonders of Thy power hast granted even to the weaker sex the triumph of martyrdom:
	mercifully grant that we who celebrate the heavenly birthday of blessed Martina Thy virgin and martyr may,
	through her example, advance nearer to Thee.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}


\end{document}