\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, 19 Feb 2022.}

 \textit{Vespers today is based on the following Sunday, which is Sexagesima Sunday.
 All is said as in the main booklet until the Little Chapter.
 As noted in the main booklet, \textnormal{Alleluia} at the beginning of the Office is omitted and replaced with
 \textnormal{Laus tibi Domine, rex æternæ gloriæ}.}

 \section*{Little Chapter (2 Cor. 11:19-20)}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Libénter suffértis insipiéntens, cum sitis ipsi sapiéntes:~\GreDagger\
	 sustinétis enim, si quis vos in servitútem rédigit, si quis dévorat, si quis áccipit, si quis extóllitur,~*
	 si quis in fáciem vos cædit.
 	\Rbar.~Deo gratias.
 }{
 	Brethren: You gladly put up with fools, because you are wise yourselves!
 	For you suffer it if a man enslaves you, if a man devours you, if a man takes from you, if a man is arrogant, if a man slaps your face!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/o-lux-beata-trinitas}

{\itshape
	1. O Trinity of blessed Light,
	O Unity of sovereign might,
	as now the fiery sun departs,
	shed Thou Thy beams within our hearts.
	
	2. To Thee our morning song of praise,
	to Thee our evening prayer we raise;
	Thee may our glory evermore
	in lowly reverence adore.
	
	3. All laud to God the Father be;
	all praise, Eternal Son, to Thee;
	all glory, as is ever meet,
	to God the Holy Paraclete.
}

\vfill\pagebreak

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h) Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~May our evening prayer ascend to Thee, O Lord.
\Rbar.~And may Thy mercy descend upon us.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} The Lord said to Noah: the end of all flesh is come before me:
make thee an ark of timber planks, that the seed of all flesh may be saved therein.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui cónspicis, quia ex nulla nostra actióne confídimus:
	concéde propítius; ut, contra advérsa ómnia, Doctóris géntium protectióne muniámur.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who seest that we put no trust in anything we do,
	mercifully grant that by the protection of the Doctor of the Gentiles we may be defended against all adversity.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Suffrage of the Saints}

\textit{\textnormal{Ant.} May the Blessed Mother of God, the Virgin Mary, and all the Saints, intercede for us before the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/suffrage-saints}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Mirificávit Dóminus Sanctos suos.\\
	\Rbar. Et exaudívit eos clamántes ad se.
	
	Orémus.
	A cunctis nos, qu\'{\ae}sumus Dómine, mentis et córporis defénde perículis:
	et intercedénte beáta et gloriósa semper Vírgine Dei Genitríce María,
	cum beáto Joseph, beátis Apóstolis tuis Petro et Paulo, atque beáto Thoma apóstolo et ómnibus Sanctis,
	salútem nobis tríbue benígnus et pacem;
	ut destrúctis adversitátibus et erróribus univérsis,
	Ecclésia tua secúra tibi sérviat libertáte.
	Per Dóminum.
}{
	The Lord hath made His holy ones wonderful.
	\Rbar.~And He hath heard them when they cried unto Him.
	
	Let us pray.
	From all perils of soul and body defend us, O Lord, we beseech Thee,
	and by the intercession of blessed and glorious ever Virgin Mary, Mother of God,
	of blessed Joseph, of thy blessed apostles Peter and Paul, and of blessed apostle Thomas and all the Saints,
	graciously grant us safety and peace that all adversities and errors being overcome,
	thy Church may serve Thee in security and freedom. Through our Lord. \Rbar.~Amen.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

%% TODO: make sure this is correct
\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}