\documentclass[12pt]{../vespers-sheet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers on Saturdays in Eastertide}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia. \textnormal{Or:} Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

\section*{Psalm 143}

\textit{\textnormal{Ant.} Alleluia. \textnormal{Ps.} Blessed be the Lord my God, who teacheth my hands to fight, and my fingers to war.}

 \begin{rubricbox}

After the psalm is intoned, all \textbf{sit} at the asterisk midway through the first verse.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps143-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps143-6}	
}{
	\input{../psalms/english/ps143}
}

\end{latinenglishsection}

\section*{Psalm 144}

\textit{\textnormal{Ps.} I will extol thee, O God my king: and I will bless thy name for ever; yea, for ever and ever.}

 \begin{rubricbox}

Everyone else will \textbf{remain seated} as the cantor stands to intone the psalm.
The same is done at the beginning of each of the following psalms.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{ps144-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps144-6}
}{
	\input{../psalms/english/ps144}
}

\end{latinenglishsection}

\section*{Psalm 145}

\textit{\textnormal{Ps.} Praise the Lord, O my soul, in my life I will praise the Lord: I will sing to my God as long as I shall be.}

\gresetinitiallines{0}
\gregorioscore{ps145-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps145-6}
}{
	\input{../psalms/english/ps145}
}

\end{latinenglishsection}

\pagebreak

\section*{Psalm 146}

\textit{\textnormal{Ps.} Praise ye the Lord, because psalm is good: to our God be joyful and comely praise.}

\gresetinitiallines{0}
\gregorioscore{ps146-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps146-6}
}{
	\input{../psalms/english/ps146}
}

\end{latinenglishsection}

\pagebreak

\section*{Psalm 147}

\textit{\textnormal{Ps.} Praise the Lord, O Jerusalem: praise thy God, O Sion.}

\gresetinitiallines{0}
\gregorioscore{ps147-intonation}

 \begin{latinenglishsection}

\latinenglish{
	\input{../psalms/ps147-6}
}{
	\input{../psalms/english/ps147}
}

 \end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{pss-antiphon}
%\textit{\textnormal{Ant.} Alleluia, alleluia, alleluia.}

\section*{Little Chapter}

\begin{rubricbox}

This section of the Office changes each week and is found in today's insert.

\end{rubricbox}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/ad-coenam-agni}

{\itshape
	1. The Lamb's high banquet we await
	in snow-white robes of royal state:
	and now, the Red Sea's channel past,
	to Christ our Prince we sing at last. 
	
	2. Upon the Altar of the Cross
	His Body hath redeemed our loss:
	and tasting of his roseate Blood,
	our life is hid with Him in God. 
	
	3. That Paschal Eve God's arm was bared,
	the devastating Angel spared:
	by strength of hand our hosts went free
	from Pharaoh's ruthless tyranny. 
	
	4. Now Christ, our Paschal Lamb, is slain,
	the Lamb of God that knows no stain,
	the true Oblation offered here,
	our own unleavened Bread sincere. 
	
	5. O Thou, from whom hell's monarch flies,
	O great, O very Sacrifice,
	Thy captive people are set free,
	and endless life restored in Thee.
	
	6. For Christ, arising from the dead,
	from conquered hell victorious sped,
	and thrust the tyrant down to chains,
	and Paradise for man regains. 
	
	7. We pray Thee, King with glory decked,
	in this our Paschal joy, protect
	from all that death would fain effect
	Thy ransomed flock, Thine own elect.
	
	8. To Thee who, dead, again dost live,
	all glory Lord, Thy people give;
	all glory, as is ever meet,
	to Father and to Paraclete. Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ma(h)ne(h) no(h)bís(h)cum(h) Dó(h)mi(h)ne,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Quó(h)ni(h)am(h) ad(h)ves(h)pe(h)rá(h)scit,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Stay with us, Lord, alleluia.
\Rbar.~For it is getting towards evening, alleluia.}

\section*{Magnificat, Collect, and Commemorations}

\begin{rubricbox}

This section of the Office changes each week and is found in today's insert.

\end{rubricbox}

\section*{Conclusion}

\textit{\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.}

\gresetinitiallines{0}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\rubric{If no other service follows:}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Cæli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain with us always. \Rbar.~Amen.
}

\end{latinenglishsection}

\end{document}