\documentclass{../../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Friday, February 3, 2023}

 \textit{Vespers today is that of St. Andrew Corsini, of double rank.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Sirach 44:16-17)}

\begin{latinenglishsection}

 \latinenglish{
 	Ecce sacérdos magnus,~\GreDagger\ qui in diébus suis plácuit Deo, et invéntus est justus:~*
 	et in témpore iracúndiæ factus est reconciliátio.
 	\Rbar.~Deo gratias.
 }{
	Behold a great priest, who in his days pleased God, and was found just,
	and in the time of wrath was made a means of reconciliation.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../../hymns/iste_confessor_mtv}

%\vfill
%\pagebreak

{\itshape

	1. This the confessor of the Lord, whose triumph
	Now all the faithful celebrate, with gladness
	On this feast day year by year receiveth
	Merited honors.
	
	2. Saintly and prudent, modest in behaviour,
	Peaceful and sober, chaste was he, and lowly.
	While that life’s vigour, coursing through his members,
	Quickened his being.
	
	3. Sick ones of old time, to his tomb resorting,
	Sorely by ailments manifold afflicted,
	Oft-times have welcomed health and strength returning,
	At his petition.
	
	4. Whence we in chorus gladly do him honour,
	Chanting his praises with devout affection,
	That in his merits we may have a portion,
	Now and forever.
	
	5. His be the glory, power and salvation,
	Who over all things reigneth in the highest,
	Earth’s mighty fabric ruling and directing,
	Onely and Trinal.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>.  A(h)má(h)vit(h) e(h)um(h) Dó(h)mi(h)nus(h), et(h) or(h)ná(h)vit(h) e(h)um.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Sto(h)lam(h) gló(h)ri(h)æ(h) ín(h)du(h)it(h) e(h)um.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~The Lord loved him and beautified him.
\Rbar.~He clothed him with a robe of glory.}

\section*{Magnificat}

\begin{rubricbox}

The cantor intones the antiphon and leads the first verse:

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../../psalms/magnificat-1}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.}  O Priest and Bishop, thou worker of mighty works, good shepherd of the people, pray unto the Lord for us.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the Officiant.

\end{rubricbox}

\section*{Collect and conclusion}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui in Ecclesia tua nova semper instauras exempla virtutum: da pópulo tuo beáti Andréæ Confessóris tui
atque Pontíficis ita sequi vestígia, ut assequá\textit{tur et} pr\'{\ae}mia.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.\\
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, who in every age dost manifest in Thy Church new examples of virtue;
	grant that Thy people may so follow in the footsteps of blessed Andrew, Thy confessor and bishop,
	that they may obtain the reward now enjoyed by him.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../../common/benedicamus-1v-double}

\vfill\pagebreak

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	\Vbar. May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	\Vbar. May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

%\vfill\pagebreak

\section*{Marian Anthem}

\gresetinitiallines{1}
\gregorioscore{../../marian-anthems/ave-regina-caelorum-simple}

{\itshape
	Hail, O Queen of Heaven enthroned.
	Hail, by angels mistress owned.
	Root of Jesse, Gate of Morn
	whence the world's true light was born.
	
	Glorious Virgin, Joy to thee,
	loveliest whom in heaven they see;
	fairest thou, where all are fair,
	plead with Christ our souls to spare.
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Dignáre me laudáre te, Virgo sacráta.\\
	\Rbar. Da mihi virtútem contra hostes tuos.
	
	Orémus.
	Concéde, miséricors Deus, fragilitáti nostræ præsídium:
	ut qui sanctæ Dei Genitrícis memóriam ágimus;
	intercessiónis ejus auxílio, a nostris iniquitátibus resurgámus.
	Per eúmdem Christum Dóminum nostrum.
	\Rbar.~Amen.
}{
	Vouchsafe that I may praise thee, O sacred Virgin.
	\Rbar.~Give me stength against thine enemies.
	
	Let us pray.
	Grant, O merciful God, to our weak natures Thy protection,
	that we who commemorate the holy Mother of God may,
	by the help of her intercession, arise from our iniquities.
	Through the same Christ our Lord.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Vbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}
\end{document}