\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, July 16}

 \textit{Vespers today is based on the following Sunday, which is the Sixth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 11:33)}

\begin{latinenglishsection}

 \latinenglish{
 	O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\
 	quam incomprehensibília sunt judícia ejus, et investigábiles viæ ejus!
	 
 	\Rbar.~Deo grátias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h), Dó(h)mi(f)ne(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a(f.). (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} I beseech thee, O Lord, do away the iniquity of thy servant, for I have done very foolishly.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus virtútum, cujus est totum quod est óptimum: ínsere pectóribus nostris amórem tui nóminis, et præsta in nobis religiónis augméntum; ut, quæ sunt bona, nútrias, ac pietátis stúdio, quæ sunt nutríta, custódias.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
Lord of all power and might, Who art the Author and Giver of all good things, graft in our hearts the love of thy Name, increase in us true religion, nourish us with all goodness, and, of thy great mercy, keep us in the same.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of St. Alexius of Rome, Confessor (July 17)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\
	\Rbar.~Stolam glóriæ induit eum.
	
	Orémus.
	Deus, qui nos beáti Aléxii Confessóris tui ánnua solemnitáte lætíficas: concéde propítius; ut cujus natalícia cólimus, étiam actiónes imitémur.
	Per Dóminum.
}{
	The Lord loved him and adorned him. 
	\Rbar.~He clothed him with a robe of glory.
	
	Let us pray.
	O God, Who, year by year, dost gladden us by the solemn feast-day of thy blessed confessor Alexis, mercifully grant unto all who keep his birthday, grace to follow after the pattern of his godly conversation.
}

\end{latinenglishsection}

\section*{Commemoration of Our Lady of Mt. Carmel}

\textit{\textnormal{Ant.} The glory of Lebanon shall be given unto it, the excellency of Carmel and Sharon, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--gloria_libani--solesmes}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Dignáre me laudáre te, Virgo sacráta.\\
	\Rbar.~Da mihi virtútem contra hostes tuos.
	
	Orémus.
	Deus, qui beatíssimæ semper Vírginis et Genetrícis tuæ Maríæ singulári título Carméli órdinem decorásti: concéde propítius; ut, cujus hódie Commemoratiónem solémni celebrámus offício, ejus muníti præsídiis, ad gáudia sempitérna perveníre mereámur:
	Qui vivis et regnas cum Deo Patre, in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar. Amen.
}{
	\Vbar.~The Lord loved him and beautified him.
	\Rbar.~He clothed him with a robe of glory.
	
	Let us pray.
	O Lord, Who hast given this excellency unto the Order of Carmel that the same should be especially styled the Order of the Most Blessed Mary, always a Virgin, thine Own Mother, mercifully grant that we who on this day do renew her memory in solemn worship, may worthily be shielded by her protection, and finally attain unto thine everlasting joy.
	Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.
	\Rbar. Amen.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}