\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, 22 January 2022}

 \textit{Vespers today is based on the following Sunday, which is the Third after Epiphany. All is said as in the main booklet until the Magnificat.}

\section*{Magnificat}

%% all the Sundays after Epiphany share the same antiphon
\gresetinitiallines{1}
\gregorioscore{../post-epiph-2/magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{../post-epiph-2/magnificat-antiphon}

\textit{\textnormal{Ant.} God hath received Israel His servant according as He spoke to Abraham and to his seed, and hath exalted the humble forever.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens sempitérne Deus, infirmitátem nostram propítius réspice:
	atque ad protegéndum nos dexteram tuæ majestátis exténde.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and eternal God, mercifully look upon our infirmity,
	and stretch forth the right hand of Thy majesty to help and defend us.
	Through our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of Sts Vincent and Anastasius, Martyrs (Jan. 22)}

\textit{\textnormal{Ant.} In heaven the souls of the Saints rejoice, who have followed in Christ's footsteps;
and since they shed their blood for him, they rejoice with Him forever.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Exsultábunt Sancti in glória.\\
	\Rbar.~Lætabúntur in cubílibus suis.
	
	Orémus.
	Adésto Dómine supplicatiónibus nostris:
	ut qui ex iniquitáte nostra reos nos esse cognóscimus,
	beatórum Mártyrum tuórum Vincéntii et Anastásii intercessióne liberémur.
}{
	The just exult in their glory.
	\Rbar.~They shout for joy on their beds.
	
	Let us pray.
	Graciously hear, O Lord, the humble prayers of those who acknowledge the guilt of their sins;
	that by the intercession of Thy blessed martyrs, Vincent and Anastasius, they may be delivered.
}

\end{latinenglishsection}

\section*{Commemoration of a St. Raymond of Peñafort, Confessor (Jan. 23)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ induit eum.  %PT: allelúia
	
	Orémus.
	Deus qui beátum Raymúndum pæniténtiæ sacraménti insígnem minístrum elegísti,
	et per maris undas mirabíliter traduxísti:
	concéde; ut ejus intercessióne dignos pæniténtiæ fructus fácere,
	et ad ætérnæ salútis portum perveníre valeámus.
}{
	The Lord loved him and adorned him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	O God, Who didst elect blessed Raymund to be an eminent minister of the sacrament of penance,
	and didst miraculously guide him through the waves of the sea,
	grant that by his intercession, we may bring forth worthy fruits of penance,
	and reach the haven of eternal life.
}

\end{latinenglishsection}

\section*{Commemoration St. Emerentiana, Virgin and Martyr (Jan. 23)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua.\\ %PT: allelúia
	\Rbar.~Inténde, próspere procéde, et regna.  %PT: allelúia
	
	Orémus.
	Indulgéntiam nobis, qu\'{\ae}sumus Dómine, beáta Emerentiána Virgo et Martyr implóret:
	quæ tibi grata semper éxstitit et mérito castitátis,
	et tuæ professióne virtútis.
}{
	In thy glory and thy splendor. %PT: alleluia
	\Rbar.~Go forth, advance with victory and reign. %PT: alleluia
	
	Let us pray.
	May blessed Emerentiana Thy virgin and martyr, we bseech Thee, O Lord,
	implore for us Thy forgiveness; for she was ever pleasing unto Thee,
	both by the merit of her chastity, and by her confession of Thy power.
}

\end{latinenglishsection}

\section*{Suffrage of the Saints}

\textit{\textnormal{Ant.} May the Blessed Mother of God, the Virgin Mary, and all the Saints, intercede for us before the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/suffrage-saints}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Mirificávit Dóminus Sanctos suos.\\
	\Rbar. Et exaudívit eos clamántes ad se.
	
	Orémus.
	A cunctis nos, qu\'{\ae}sumus Dómine, mentis et córporis defénde perículis:
	et intercedénte beáta et gloriósa semper Vírgine Dei Genitríce María,
	cum beáto Joseph, beátis Apóstolis tuis Petro et Paulo, atque beáto Thoma apóstolo et ómnibus Sanctis,
	salútem nobis tríbue benígnus et pacem;
	ut destrúctis adversitátibus et erróribus univérsis,
	Ecclésia tua secúra tibi sérviat libertáte.
	Per Dóminum.
}{
	The Lord hath made His holy ones wonderful.
	\Rbar.~And He hath heard them when they cried unto Him.
	
	Let us pray.
	Let us pray. From all perils of soul and body defend us, O Lord, we beseech Thee,
	and by the intercession of blessed and glorious ever Virgin Mary, Mother of God,
	of blessed Joseph, of thy blessed apostles Peter and Paul, and of blessed apostle Thomas and all the Saints,
	graciously grant us safety and peace that all adversities and errors being overcome,
	thy Church may serve Thee in security and freedom. Through our Lord. \Rbar.~Amen.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}