\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers of the Visitation of the Blessed Virgin Mary}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently.
Then, when the Officiant stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} Mary arose * and went into the hill country with haste, into a city of Judah.
 \textnormal{Ps.} The Lord said to my Lord: * Sit thou at my right hand:}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-8}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Exsúrgens María\dots}}

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 2.} Mary entered * into the house of Zacharias, and saluted Elizabeth.
 \textnormal{Ps.} Praise the Lord, ye children: * praise ye the name of the Lord.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the Psalm is intoned by the Cantor, all \textbf{sit} at the asterisk. The same postures are observed for the remaining Psalms.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-2}

}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Intrávit María\dots}}

\end{rubricbox}

\section*{Psalm 126}

\textit{\textnormal{Ant. 4.} Blessed art thou among women, * and blessed is the Fruit of thy womb.
 \textnormal{Ps.} Unless the Lord build the house, * they labour in vain that build it.}

\gresetinitiallines{1}
\gregorioscore{ps126-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps126-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps126-7a}
	
}{
	\input{../psalms/english/ps126}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Benedícta\dots}}

\end{rubricbox}

\pagebreak

\section*{Psalm 147}

\textit{\textnormal{Ant. 5.} As soon as * the voice of thy salutation sounded in mine ears, the babe leaped in my womb for joy. Alleluia.
 \textnormal{Ps.} Praise the Lord, O Jerusalem: * praise thy God, O Sion.}

\gresetinitiallines{1}
\gregorioscore{ps147-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps147-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps147-7a}

}{
	\input{../psalms/english/ps147}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Ex quo facta est\dots}}

\end{rubricbox}

\section*{Little Chapter (Sirach 24:14)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Ab inítio et ante s\'{\ae}cula creáta sum\greheightstar\,
	et usque ad futúrum s\'{\ae}culum \textit{non dési}nam, et in habitatióne sancta coram ipso ministrávi.
	\Rbar.~Deo grátias.
}{
	From the beginning, and before the world, was I created, and unto the world to come I shall not cease to be, and in the holy dwelling place I have ministered before him.
	\Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/ave-maris-stella}

{\itshape
	1. Hail, star of ocean,
	Child divine who barest,
	Mother, ever-Virgin,
	Heaven's portal fairest.
	
	2. Taking that sweet Ave
	Erst by Gabriel spoken,
	Eva's name reversing,
	Be of peace the token.
	
	3. Break the sinners' fetters,
	Light to blind restoring,
	All our ills dispelling,
	Every boon imploring.
	
	4. Show thyself a mother
	In thy supplication;
	He will hear who chose thee
	At his incarnation.
	
	5. Maid all maids excelling,
	Passing meek and lowly,
	Win for sinners pardon,
	Make us chaste and holy.
	
	6. As we onward journey
	Aid our weak endeavour,
	Till we gaze on Jesus
	And rejoice forever.
	
	7. Father, Son, and Spirit,
	Three in One confessing,
	Give we equal glory
	Equal praise and blessing.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{(c3) <sp>V/</sp>. Be(h)ne(h)dí(h)cta(h) tu(h) in(h) mu(h)li(h)é(h)ri(f)bus(f.). (::)}

\gresetinitiallines{0}
\gabcsnippet{(c3) <sp>R/</sp>. Et(h) be(h)ne(h)díc(h)tus(h) fru(h)ctus(h) ven(h)tris(h) tu(h)i(f.). (::)}

\textit{\Vbar.~Blessed art thou among women.
\Rbar.~And blessed is the Fruit of thy womb.}

\pagebreak

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Behold a faithful and prudent servant, whom the Lord has set over His household.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

{\color{red}After the Cantor intones up to the asterisk, all \textbf{sit} and join in singing.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	3. \textit{Quia} respéxit humilitátem ancíllæ \textbf{su}æ:~* ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.
	
	4. \textit{Quia} fecit mihi magna qui \textbf{pot}ens est:~* et sanctum \textit{no}\textit{men} \textbf{e}jus.
	
	5. \textit{Et\ mi}sericórdia ejus a progénie in pro\textbf{gé}nies~* timén\textit{ti}\textit{bus} \textbf{e}um.
	
	6. \textit{Fecit} poténtiam in bráchio \textbf{su}o:~* dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.
	
	7. \textit{Depó}suit poténtes de \textbf{se}de,~* et exal\textit{tá}\textit{vit} \textbf{hú}miles.
	
	8. \textit{Esu}riéntes implévit \textbf{bo}nis:~* et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.
	
	9. \textit{Suscé}pit Israël púerum \textbf{su}um,~* recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.
	
	10. \textit{Sicut} locútus est ad patres \textbf{nos}tros,~* Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.
	
	11. \textit{Glóri}a Patri, et \textbf{Fí}lio,~* et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.
	
	12. \textit{Sicut} erat in princípio, et nunc, et \textbf{sem}per,~* et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.
}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~* 
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Beátam me dicent\dots},
then \textbf{stand} for the prayer.}

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Fámulis tuis, qu\'{\ae}sumus, Dómine, cæléstis grátiæ munus impertíre: ut, quibus beátæ Vírginis partus éxstitit salútis exórdium; Visitatiónis ejus votíva solémnitas, pacis tríbuat increméntum.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Grant, O Lord, we beseech thee, unto all thy servants, the gift of thine heavenly grace, that even as the Blessed Virgin being made a Mother hath been unto them the first step unto their salvation, so the godly and solemn memorial of her Visitation may be the bringer of an increase of peace.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of Pope St. Leo, Confessor (June 3)}

\textit{\textnormal{Ant.} While he was supreme Pontiff, he feared not earthly powers, but gloriously went his way to the heavenly kingdom.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an-dum_esset_summus_pontifex}
%PT: an-dum_esset_summus_pontifex_pt

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ índuit eum. %PT: allelúia
	
	Orémus.
	Gregem tuum, Pastor ætérne, placátus inténde: et per beátum Leónem Summum Pontíficem perpétua protectióne custódi; quem totíus Ecclésiæ præstitísti esse pastórem.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar. Amen.
}{
	\Vbar.~The Lord loved him and beautified him. %PT: allelúia
	\Rbar.~He clothed him with a robe of glory. %PT: allelúia
	
	Let us pray.
	Look forgivingly on thy flock, Eternal Shepherd, and keep it in thy constant protection, by the intercession of blessed Leo thy Sovereign Pontiff, whom thou didst constitute Shepherd of the whole Church.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar. Amen.
}
\end{latinenglishsection}

\section*{Commemoration of the Fourth Sunday after Pentecost}

\textit{\textnormal{Ant.} So David prevailed over the Philistine with a sling, and with a stone, in the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/praevaluit_david}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Vespertína orátio ascéndat ad te, Dómine.\\
	\Rbar.~Et descéndat super nos misericórdia tua.
	
	Orémus.
	Da nobis, qu\'{\ae}sumus, Dómine: ut et mundi cursus pacífice nobis tuo órdine dirigátur; et Ecclésia tua tranquílla devotióne lætétur.
}{
	\Vbar.~Let the evening prayer ascend unto thee, O Lord.
	\Rbar.~And let there descend upon us thy mercy.
	
	Let us pray.
	Grant, O Lord, we beseech thee, that the course of this world may be so peaceably ordered by thy governance, that thy Church may joyfully serve thee in all godly quietness.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-2v-double}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}

\section*{Marian Anthem}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/salve-regina-simple}

{\itshape
	Hail, Holy Queen, Mother of Mercy, our life, our sweetness and our hope! To thee do we cry, poor banished children of Eve; to thee do we send up our sighs, mourning and weeping in this valley of tears. Turn then, O most gracious advocate, thine eyes of mercy toward us, and after this our exile, show unto us the blessed fruit of thy womb, Jesus. O clement, O loving, O sweet Virgin Mary!
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Ora pro nóbis, sáncta Déi Génitrix.\\
	\Rbar. Ut dígni efficiámur promissiónibus Chrísti.
	
	Orémus.
	Omnípotens sempitérne Deus, qui gloriósæ Vírginis Matris Maríæ corpus et ánimam, ut dignum Fílii tui habitáculum éffici mererétur, Spíritu Sancto cooperánte præparásti: \greheightstar da, ut cujus commemoratióne \textit{læ}\textbf{tá}mur, * ejus pia intercessióne ab instántibus malis et a morte perpétua liberémur. Per eúmdem Christum Dóminum nostrum.
	\Rbar.~Amen.
}{
	\Vbar.~Pray for us, o holy Mother of God.
	\Rbar.~That we may be worthy of the promises of Christ.
	
	Let us pray. 
	Almighty and everlasting God, Who by the working of the Holy Spirit didst prepare both body and soul of the glorious Virgin Mother, Mary, that she might deserve to be made a worthy dwelling for Thy Son, grant that we who rejoice in her memory, may, by her loving intercession, be delivered from present evils and from lasting death, through the same Christ our Lord. 		
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}