1. Thou hast done well with thy servant, O Lord, * according to thy word.

2. Teach me goodness and discipline and knowledge; * for I have believed thy commandments.

3. Before I was humbled I offended; * therefore have I kept thy word.

4. Thou art good; * and in thy goodness teach me thy justifications.

5. The iniquity of the proud hath been multiplied over me: * but I will seek thy commandments with my whole heart.

6. Their heart is curdled like milk: * but I have meditated on thy law.

7. It is good for me that thou hast humbled me, * that I may learn thy justifications.

8. The law of thy mouth is good to me, * above thousands of gold and silver.

9. Thy hands have made me and formed me: * give me understanding, and I will learn thy commandments.

10. They that fear thee shall see me, and shall be glad: * because I have greatly hoped in thy words.

11. I know, O Lord, that thy judgments are equity: * and in thy truth thou hast humbled me.

12. O! let thy mercy be for my comfort, * according to thy word unto thy servant.

13. Let thy tender mercies come unto me, and I shall live: * for thy law is my meditation.

14. Let the proud be ashamed, because they have done unjustly towards me: * but I will be employed in thy commandments.

15. Let them that fear thee turn to me: * and they that know thy testimonies.

16. Let my heart be undefiled in thy justifications, * that I may not be confounded.

17. Glory be to the Father, and to the Son, * and to the Holy Ghost.

18. As it was in the beginning, is now, * and ever shall be, world without end. Amen.