2. Thou hast understood my thoughts afar off: my path and my line thou hast searched out
 	
3. And thou hast foreseen all my ways: for there is no speech in my tongue
 	
4. Behold, O Lord, thou hast known all things, the last and those of old: thou hast formed me, and hast laid thy hand upon me.
 	
5. Thy knowledge is become wonderful to me: it is high, and I cannot reach to it.
 	
6. Whither shall I go from thy spirit? or whither shall I flee from thy face?
 	
7. If I ascend into heaven, thou art there: if I descend into hell, thou art present.
 	
8. If I take my wings early in the morning, and dwell in the uttermost parts of the sea:
 	
9. Even there also shall thy hand lead me: and thy right hand shall hold me.
 	
10. And I said: Perhaps darkness shall cover me: and night shall be my light in my pleasures.
 	
11. But darkness shall not be dark to thee, and night shall be light as day: the darkness thereof, and the light thereof are alike to thee.
 	
12. For thou hast possessed my reins: thou hast protected me from my mother's womb.
 	
13. I will praise thee, for thou art fearfully magnified: wonderful are thy works, and my soul knoweth right well.
 	
14. My bone is not hidden from thee, which thou hast made in secret: and my substance in the lower parts of the earth.
 	
15. Thy eyes did see my imperfect being, and in thy book all shall be written: days shall be formed, and no one in them.
 	
16.  But to me thy friends, O God, are made exceedingly honourable: their principality is exceedingly strengthened.
 	
17.  I will number them, and they shall be multiplied above the sand: I rose up and am still with thee.
 	
18. If thou wilt kill the wicked, O God: ye men of blood, depart from me:
 	
19. Because you say in thought: They shall receive thy cities in vain.
 	
20. Have I not hated them, O Lord, that hated thee: and pine away because of thy enemies?

21. I have hated them with a perfect hatred: and they are become enemies to me.
 	
22. Prove me, O God, and know my heart: examine me, and know my paths.
 	
23. And see if there be in me the way of iniquity: and lead me in the eternal way.
 	
24. Glory be to the Father, and to the Son, and to the Holy Spirit.
 	
25. As it was in the beginning, is now, and ever shall be, world without end. Amen.
