\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers of the Saturday within the Octave of the Sacred Heart of Jesus}

\section*{Beginning of the Office}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently.
Then, when the Officiant stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.}  Elizabeth, the wife of Zacharias, * was the mother of a mighty man, even of John the Baptist, the Fore-runner of the Lord.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gabcsnippet{(c4)1. Di(g)xit(hj) Dó(j)mi(j)nus(j) <b>Dó</b>(jk)mi(j)no(j) <b>me</b>(ih)o:(hj..) <v>\greheightstar</v>(:) Se(j)de(j) a(j) <i>dex</i>(ji)<i>tris</i>(hi) <b>me</b>(h)is:(gh..) (::) E(j) u(j) o(j) u(h) a(j) e.(i.) (::)}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-1}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Elisabeth Zacharíæ\dots}}

\end{rubricbox}

\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} They made signs to his father * how he would have him called; and he asked for a writing-table, and wrote, saying: His name is John.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

\gresetinitiallines{0}
\gabcsnippet{(c4)1. Con(h)fi(gh)té(h)bor(h) ti(h)bi,(h) Dó(h)mi(h)ne,(h) in(h) to(h)<i>to</i>(hg) <i>cor</i>(gi)<i>de</i>(i) <b>me</b>(hi)o:(h.) <v>\greheightstar</v>(:) in(h) con(h)sí(h)li(h)o(h) jus(h)tó(h)rum,(h) et(h) con(h)gre(h)ga(h)ti(h)<b>ó</b>(h)ne.(g.) (::)
<i>Flex:</i>()  su(h)ó(h)rum, †(g. h h h  ::) E(h) u(h) o(h) u(h) a(h) e.(g.) (::)}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-4g}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Innuébant\dots}}

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} Thou shalt call his name John; * and many shall rejoice at his birth.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: * he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

\gresetinitiallines{0}
\gabcsnippet{(c4)1. Be(f)á(gh)tus(h) vir,(h) <i>qui</i>(hg) <i>ti</i>(ixgi)<i>met</i>(h) <b>Dó</b>(h)<b>mi</b>(hg)num:(gh..) <v>\greheightstar</v>(:) in(h) man(h)dá(h)tis(h) e(h)jus(h) <i>vo</i>(g)<i>let</i>(f) <b>ni</b>(gh)mis.(gf..) (::)
<i>Flex:</i>()  cóm(h)mo(g)dat, †(g. h h h  ::) E(h) u(h) o(g) u(f) a(gh) e.(gf..) (::)}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps111-1f}

}{
 	\input{../psalms/english/ps111}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Joánnes vocábitur\dots}}

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} Among them that are born of women * there hath not risen a greater than John the Baptist.
 \textnormal{Ps.} Praise the Lord, ye children: * praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

\gresetinitiallines{0}
\gabcsnippet{(c4)1. Lau(g)dá(hj)te,(j) <b>pú</b>(jk)e(j)ri,(j) <b>Dó</b>(j)<b>mi</b>(ih)num:(hj..) <v>\greheightstar</v>(:) lau(j)dá(j)te(j) <i>no</i>(ji)<i>men</i>(hi) <b>Dó</b>(h)mi(g)ni.(gh..) (::) E(j) u(j) o(ji) u(hi) a(h) e.(gh..) (::)}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps112-3a2}

}{
 	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Inner natos\dots}}

\end{rubricbox}

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} Thou, child, shalt be called the Prophet of the Highest; * for thou shalt go before the face of the Lord, to prepare His ways.
 \textnormal{Ps.} Praise the Lord, all ye nations: * praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

\gresetinitiallines{0}
\gabcsnippet{(c4)1. Lau(g)dá(hj)te(j) Dó(j)mi(j)num,(j) <b>om</b>(jk)nes(j) <b>Gen</b>(ih)tes:(hj..) <v>\greheightstar</v>(:) lau(j)dá(j)te(j) e(j)um,(j) <i>om</i>(ji)<i>nes</i>(hi) <b>pó</b>(h)pu(g)li:(gh..) (::) E(j) u(j) o(ji) u(hi) a(h) e.(gh..) (::)}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps116-3a2}

}{
 	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Tu, puer,\dots}}

\end{rubricbox}

\section*{Little Chapter (Isaias 49:1)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Audíte, ínsulæ, et atténdite, pópuli de longe: \GreDagger\ 
	Dóminus ab úte\textit{ro vocávit} me, de ventre matris meæ recordátus est nóminis mei.
	\Rbar.~Deo grátias.
}{
	Give ear, ye islands, and hearken, ye people from afar. The Lord hath called me from the womb, from the bowels of my mother he hath been mindful of my name.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/ut_queant_laxis--solesmes}

{\itshape
	1. O for thy spirit, holy John, to chasten
	Lips sin-polluted, fettered tongues to loosen;
	So by thy children might thy deeds of wonder
	Meetly be chanted.

	2. Lo! a swift herald, from the skies descending,
	Bears to thy father promise of thy greatness;
	How he shall name thee, what thy future story,
	Duly revealing.
	
	3. Scarcely believing message so transcendent,
	Him for season power of speech forsaketh,
	Till, at thy wondrous birth, again returneth
	Voice to the voiceless.
	
	4. Thou, in thy mother's womb all darkly cradled,
	Knewest thy Monarch, biding in his chamber,
	Whence the two parents, through their children's merits,
	Mysteries uttered.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Is(h)te(h) pu(h)er(h) mag(h)nus(h) co(h)ram(h) Dó(h)mi(f)no(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Nam(h) et(h) ma(h)nus(h) ej(h)us(h) cum(h) ip(h)so(f) est(f.). (::)
}

\textit{\Vbar.~The child shall be great in the sight of the Lord.
\Rbar.~For His hand is with him.}

\vfill\pagebreak

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} The child, that unto us is born, * is much more than a Prophet, for this is he of whom the Saviour saith: Among them that are born of women there hath not risen a greater than John the Baptist.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

{\color{red}After the Cantor intones up to the asterisk, all \textbf{sit} and join in singing.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-7}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Puer\dots},
then \textbf{stand} for the prayer.}

\end{rubricbox}

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Sancti nóminis tui, Dómine, timórem páriter et amórem fac nos habére perpétuum: quia nunquam tua gubernatióne destítuis, quos in soliditáte tuæ dilectiónis instítuis.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, Who never failest to help and govern them whom Thou dost bring up in thy steadfast fear and love keep us, we beseech thee, under the protection of thy good providence, and make us to have a perpetual fear and love of thy Holy Name.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of St. John and Paul, Martyrs (June 26)}

\textit{\textnormal{Ant.} Theirs is the kingdom of heaven, who despising a worldy life have attained the rewards of the kingdom,
and have washed their robes in the Blood of the Lamb.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Lætámini in Dómino, et exsultáte justi.\\
	\Rbar.~Et gloriámini omnes recti corde.
	
	Orémus.
	Qu\'{\ae}sumus, omnípotens Deus: ut nos gemináta lætítia hodiérnæ festivitátis excípiat, quæ de beatórum Joánnis et Pauli glorificatióne procédit; quos éadem fides et pássio vere fecit esse germános.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	Be glad in the Lord and rejoice, ye just.
	\Rbar.~And exult, all ye upright of heart.
	
	Let us pray.
	Almighty God, fill us, we beseech thee, with the twofold gladness which doth flow down upon this bright day from the glory of thy blessed servants John and Paul, whom one faith and one suffering made to be brothers indeed.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of the Saturday within the Octave of the Sacred Heart of Jesus}

\textit{\textnormal{Ant.}  And they all knew, from Dan even unto Beersheba, that Samuel was established to be a Prophet of the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--cognoverunt_omnes--solesmes}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Memóriam fecit mirabílium suórum miserátor Dóminus.\\
	\Rbar.~Escam dedit timéntibus se.
	
	Orémus.
	Protéctor in te sperántium, Deus, sine quo nihil est válidum, nihil sanctum; multíplica super nos misericórdiam tuam; ut, te rectóre, te duce, sic transeámus per bona temporália, ut non amittámus ætérna.
}{
	\Vbar.~The merciful Lord hath instituted a memorial of his wondrous deeds.
	\Rbar.~He hath given Meat unto them that fear him.
	
	Let us pray.
	O God, the protector of all them that trust in thee, without whom nothing is strong, nothing is holy; increase and multiply upon us thy mercy, that thou being our ruler and guide, we may so pass through things temporal, that we finally lose not the things eternal.
}
\end{latinenglishsection}

\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solemn}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}
\textit{Salve Regina} follows (no. 224 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}