\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers of the Saturday within the Octave of Corpus Christi}

\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.}  Christ the Lord, being made an High Priest for ever after the order of Melchisedec, hath offered bread and wine.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-1}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Sacerdos\dots}, then all \textbf{stand}.

\end{rubricbox}


\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} He hath made His wonderful works to be remembered; the Lord is gracious and full of compassion. He hath given meat unto them that fear Him.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-2}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Miserator Dominus\dots}, then all \textbf{stand}.

\end{rubricbox}

\pagebreak

\section*{Psalm 115}

\textit{\textnormal{Ant. 3.}  I will take the cup of salvation, and offer the sacrifice of thanksgiving.
 \textnormal{Ps.} I have believed, therefore have I spoken; * but I have been humbled exceedingly.}

\gresetinitiallines{1}
\gregorioscore{ps115-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps115-3}
	
}{
	\input{../psalms/english/ps115}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{Cálicem salutáris\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 127}

\textit{\textnormal{Ant. 3.}  Let the children of the Church be like olive-plants round about the table of the Lord.
 \textnormal{Ps.} Blessed are all they that fear the Lord: * that walk in his ways.}

\gresetinitiallines{1}
\gregorioscore{ps127-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps127-4E}
	
}{
	\input{../psalms/english/ps127}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{Sicut novéllæ olivárum\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 147}

\textit{\textnormal{Ant. 3.}  The Lord, That maketh peace in the borders of the Church, filleth her with the finest of the wheat.
 \textnormal{Ps.} Praise the Lord, O Jerusalem: * praise thy God, O Sion.}

\gresetinitiallines{1}
\gregorioscore{ps147-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps147-5}
	
}{
	\input{../psalms/english/ps147}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{Qui pacem\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (1 John 3:13-14)}

\begin{latinenglishsection}

\latinenglish{
	Caríssimi: Nolíte mirári, si odit vos mundus. \GreDagger\
	Nos scimus quóniam transláti sumus de mor\textit{te ad vitam}, quóniam dilígimus fratres.
	\Rbar.~Deo grátias.
}{
	Wonder not, brethren, if the world hate you. We know that we have passed from death to life, because we love the brethren. He that loveth not, abideth in death.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/pange-lingua-gloriosi-corporis}

{\itshape
	1. Sing, my tongue, the Saviour's glory,
	Of his flesh the mystery sing;
	Of the blood, all price exceeding,
	Shed by our immortal King,
	Destined, for the world's redemption,
	From a noble womb to spring.

	2. Of a pure and spotless Virgin
	Born for us on earth below,
	He, as man, with man conversing,
	Stayed, the seeds of truth to sow;
	Then he closed in solemn order
	Wondrously his life of woe.
	
	3. On the night of that last supper
	Seated with his chosen band,
	He, the paschal victim eating,
	First fulfills the law's command:
	Then as food to all his brethren
	Gives himself with his own hand.
	
	4. Word made flesh, the bread of nature
	By his word to flesh he turns;
	Wine into his blood he changes:
	What though sense no change discerns?
	Only be the heart in earnest,
	Faith her lesson quickly learns.
	
	[If the Office is celebrated during Exposition, the following verses are said while kneeling]
	
	5. Down in adoration falling,
	Lo! the sacred Host we hail;
	Lo! o'er ancient forms departing,
	Newer rites of grace prevail;
	Faith for all defects supplying,
	Where the feeble senses fail.
	
	6. To the everlasting Father,
	And the Son who reigns on high,
	With the Holy Ghost proceeding
	Forth from each eternally,
	Be salvation, honour, blessing,
	Might, and endless majesty.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ci(h)bá(h)vit(h) il(h)los(h) ex(h) á(h)di(h)pe(h) fru(h)mé(h)nti(h), al(h)le(h)lú(h)ja(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h) pe(h)tra(h), me(h)lle(h) sa(h)tu(h)rá(h)vit(h) e(h)os(h), al(h)le(h)lú(h)ja(f.). (::)
}

\textit{\Vbar.~He fed them with the finest of the wheat. Alleluia.
\Rbar.~And with honey out of the rock did He satisfy them. Alleluia.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} The child Samuel ministered unto God before Eli, and the word of the Lord was precious in his sight.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-1}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon: \textit{Puer Sámuel\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\pagebreak

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Sancti nóminis tui, Dómine, timórem páriter et amórem fac nos habére perpétuum: quia nunquam tua gubernatióne destítuis, quos in soliditáte tuæ dilectiónis instítuis.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, Who never failest to help and govern them whom Thou dost bring up in thy steadfast fear and love keep us, we beseech thee, under the protection of thy good providence, and make us to have a perpetual fear and love of thy Holy Name.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\section*{Commemoration of St. Juliana Falconieri, Virgin (June 19)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-1v}
% PT: \gregorioscore{../commemorations/virgin-1v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua.\\ %PT: allelúia
	\Rbar.~Inténde, próspere procéde, et regna.  %PT: allelúia
	
	Orémus.
	Deus, qui beátam Juliánam Vírginem tuam extrémo morbo laborántem, pretióso Fílii tui Córpore mirabíliter recreáre dignátus es: concéde, quǽsumus; ut, ejus intercedéntibus méritis, nos quoque eódem in mortis agóne refécti ac roboráti, ad cæléstem pátriam perducámur.
}{
	In thy glory and thy splendor. %PT: alleluia
	\Rbar.~Go forth, advance with victory and reign. %PT: alleluia
	
	Let us pray.
	O God, Who, when thy blessed hand-maiden Juliana was lying sick unto death wast pleased in wondrous wise to comfort her with the Precious Body of thy Son, be Thou entreated for the same thy servant's sake, and grant unto us also the same Comfort in our last agony, that we may go in the strength of that Meat unto our very Fatherland, which is in heaven.

}

\end{latinenglishsection}

\section*{Commemoration of St. Ephrem the Syrian, Confessor and Doctor of the Church (June 18)}

\textit{\textnormal{Ant.} O right excellent Teacher, light of the Holy Church, St. N, Lover of God's law, pray for us to the Son of God.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--o_doctor_optime--solesmes_1961}
% PT: \gregorioscore{../commemorations/common-bvm-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ %PT: allelúia
	\Rbar.~Et osténdit illi regnum Dei . %PT: allelúia
	
	Orémus.
	Deus, qui Ecclésiam tuam beáti Ephræm Confessóris tui et Doctóris mira eruditióne et præcláris vitæ méritis illustráre voluísti: te súpplices exorámus; ut, ipso intercedénte, eam advérsus erróris et pravitátis insídias perénni tua virtúte deféndas.
}{
	\Vbar.~The Lord guided the just in right paths. %PT: alleluia
	\Rbar.~And showed him the kingdom of God. %PT: alleluia
	
	Let us pray.
	O God, who didst will to illumine thy Church by the wondrous learning and glorious merits of blessed Ephrem, thy Confessor and Doctor, do thou, we humbly entreat thee, by his intercession and thy continual power, defend her against the wiles of heresy and wickedness.
}

\end{latinenglishsection}

\pagebreak

\section*{Commemoration of Sts. Gervasius and Protasius (June 19)}

\textit{\textnormal{Ant.} Theirs is the kingdom of heaven, who despising a worldy life have attained the rewards of the kingdom,
and have washed their robes in the Blood of the Lamb.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Lætámini in Dómino, et exsultáte justi.\\
	\Rbar.~Et gloriámini omnes recti corde.
	
	Orémus.
	Deus, qui nos ánnua sanctórum Mártyrum tuórum Gervásii et Protásii solemnitáte lætíficas: concéde propítius; ut, quorum gaudémus méritis, accendámur exémplis.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	Be glad in the Lord and rejoice, ye just.
	\Rbar.~And exult, all ye upright of heart.
	
	Let us pray.
	O God, who dost every year by the solemn memorial of your blessed martyrs, Gervase and Protase, grant us grace, we beseech thee, not only to rejoice because of their worthy deeds, but also to tread in their footsteps.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\pagebreak

\section*{Commemoration of the Octave of Corpus Christi}

\textit{\textnormal{Ant.}  Holy exceedingly is the Supper of the Lord, * wherein we do feed on Christ, do show His death till He come, do get grace abundantly to our souls, and do take pledge of the glory which shall hereafter be revealed in us. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--o_sacrum_convivium--solesmes}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Panem de cælo præstitísti eis, allelúja..\\
	\Rbar.~Omne delectaméntum in se habéntem, allelúja.
	
	Orémus.
	Deus, qui nobis sub Sacraménto mirábili passiónis tuæ memóriam reliquísti: tríbue, quǽsumus, ita nos córporis, et sánguinis tui sacra mystéria venerári; ut redemptiónis tuæ fructum in nobis júgiter sentiámus:
}{
	\Vbar.~Thou didst send them from heaven Alleluia.
	\Rbar.~Bread able to content every man's delight Alleluia
	
	Let us pray.
	O God, Who under a wonderful Sacrament hast left unto us whereby to show forth thy Suffering Death, grant unto us, we beseech thee, so reverently to handle the Sacred Mysteries of thy Body and thy Blood that we may alway feel within ourselves the fruit of thy Redeeming Work.
}
\end{latinenglishsection}

\begin{latinenglishsection}
\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen.  \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solem}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 224 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}