\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, November 26, 2021}

 \textit{Vespers today is based on the following Sunday, which is the First Sunday of Advent.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 13:11)}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Hora est jam nos de somno súrgere:~\GreDagger\
	nunc enim própior est nostra salus,~*
	quam cum credídimus.
	\Rbar.~Deo grátias.
}{
	 Brethren: It is now the hour for us to rise from sleep:
	 for now our salvation is nearer than when we believed.
	 \Rbar.~Thanks be to God.
}
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

\vfill
\pagebreak

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

%% TODO: correct the magnificat antiphon text.

\textit{\textnormal{Ant.} Behold, the Name of the Lord comes from afar, and His glory fills all the earth.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

%% TODO: proper collect for the date

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Excita, qu\'{\ae}sumus, Dómine, poténtiam tuam, et veni:~\GreDagger\
	ut ab imminéntibus peccatórum nostrórum perículis, te mereámur protegénte éripi, te liberánte salvári.
	Qui vivis et regnas cum Deo Patre, in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Show forth Thy power, O Lord, we beseech Thee, and come, that with Thee as our protector we may be rescued from the impending danger of our sins; and with Thee as our deliverer, we may obtain our salvation. 
	Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of a St. Catherine of Alexandria Virgin Martyr (November 25)}

\textit{\textnormal{Ant.} He that will come after Me, let him deny himself, and take up his cross, and follow Me.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/one-martyr-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justus ut palma florébit.\\
	\Rbar.~Sicut cedrus Líbani sic multiplicábitur.
	
	Orémus.
	Deus, qui dedísti legem Móysi in summitáte montis Sínai, et in eódem loco per sanctos Angelos tuos corpus beátæ Catharínæ Vírginis et Mártyris tuæ mirabíliter collocásti: præsta, qu\'{\ae}sumus; ut, ejus méritis et intercessióne, ad montem, qui Christus est, perveníre valeámus.
}{
	The just shall flourish like the palm tree.
	\Rbar.~And shall grow up like the cedar of Libanus.
	
	Let us pray.
	O God, Who didst give the law to Moses on the summit of Mount Sinai and by means of Thy holy angels didst miraculously place there the body of blessed Catherine, Thy Virgin and Martyr: grant we beseech Thee, that, by her merits and intercession, we may be able to come unto the mountain which is Christ.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of the Octave of the Dedication of our Parish Church (November 26)}

\textit{\textnormal{Ant.} O how terrible is this place! Indeed this is no other but the house of God and the gate of heaven.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/dedication-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Domum tuam Dómine decet sanctitúdo.\\
	\Rbar.~In longitúdinem diérum.
	
	Orémus.
	Deus, qui nobis per síngulos annos hujus sancti templi tui consecratiónis réparas diem,
	et sacris semper mystériis repræséntas incólumes:
	exáudi preces pópuli tui, et præsta;
	ut quisquis hoc templum benefícia petitúrus ingréditur,
	cuncta se impetrásse lætétur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	Holiness becometh Thy house, O Lord.
	\Rbar.~To the end of days.
	
	Let us pray.
	O God, who every year dost renew the day of the consecration of this Thy holy temple,
	and continuest to bring us to safety to Thy sacred mysteries,
	graciously hear the prayers of Thy people, and grant that all who enter this temple to implore Thy blessings,
	may rejoice in obtaining whatsoever they ask.
	Through our Lord Jesus Christ, Thy Son, Who lives and reigns with Thee in the unity of the Holy Spirit, God, forever and ever.
}

\end{latinenglishsection}

\section*{Commemoration of St. Sylvester, Abbot (November 26)}

\textit{\textnormal{Ant.} This man, despising the world, and triumphing over earthly things, hath laid up treasures in heaven by word and deed.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\
	\Rbar.~Et osténdit illi regnum Dei.  
	
	Orémus.
	Clementíssime Deus, qui sanctum Silvéstrum Abbátem, s\'{\ae}culi hujus vanitátem in apérto túmulo pie meditántem, ad erémum vocáre, et præcláris vitæ méritis decoráre dignátus es: te súpplices exorámus; ut, ejus exémplo terréna despiciéntes, tui consórtio perfruámur ætérno.
}{
	The Lord led the just in right paths.
	\Rbar.~And showed him the kingdom of God.
	
	Let us pray.
	Most merciful God, Who when the holy Abbot Sylvester was devoutly meditating upon the vanity of this world beside an open grave, graciously willed to call him into the desert and enrich him with unusual merits, we humbly pray that, following his example, despising the things of earth, we may thoroughly enjoy Thy everlasting presence.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of St. Peter of Alexandria, Bishop Martyr (November 26)}

\textit{\textnormal{Ant.} If any man will come after me, let him deny himself and take up his cross daily and follow me.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--qui_vult_venire--solesmes_1961}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Justus ut palma florébit.\\
	\Rbar.~Sicut cedrus Líbani multiplicábitur.
	
	Orémus.
	Infirmitátem nostram réspice, omnípotens Deus: et, quia pondus própria; actiónis gravat, beáti Petri Mártyris tui atque Pontíficis intercéssio gloriósa nos prótegat. 
	Per Dóminum nostrum Jesum Christum, Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	\Vbar.~The righteous shall flourish like the palm-tree.
	\Rbar.~He shall grow like a cedar in Lebanon.
	
	Let us pray.
	Look upon our infirmity, O almighty God, and, because the weight of our own evil deeds beareth us down, may the glorious intercession of blessed Peter, Thy martyr and bishop, protect us. 
	Through our Lord Jesus Christ, Thy Son, Who lives and reigns with Thee in the unity of the Holy Spirit, God, forever and ever.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

%% TODO: benedicamus in line with the season

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

%% TODO: proper Marian anthem for the seasons

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}