\documentclass{../vespers-sheet}

%% reduce the spacing before the title so the hymn can fit all on one page
\titlespacing*{\chapter}{0pt}{35pt}{40pt}

\begin{document}

\chapter*{Vespers Insert for Saturday, Nov. 27, 2021}

 \textit{Vespers today is based on the following Sunday, which is the First of Advent. All is said as in the main booklet until the Little Chapter.}
 
 \section*{Little Chapter (Romans 13:11)}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Hora est jam nos de somno súrgere:~\GreDagger\
	nunc enim própior est nostra salus,~*
	quam cum credídimus.
	\Rbar.~Deo grátias.
}{
	 Brethren: It is now the hour for us to rise from sleep:
	 for now our salvation is nearer than when we believed.
	 \Rbar.~Thanks be to God.
}
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Behold, the Name of the Lord comes from afar, and His glory fills all the earth.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Excita, qu\'{\ae}sumus, Dómine, poténtiam tuam, et veni:~\GreDagger\
	ut ab imminéntibus peccatórum nostrórum perículis, te mereámur protegénte éripi, te liberánte salvári.
	Qui vivis.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Show forth Thy power, O Lord, we beseech Thee, and come, that with Thee as our protector we may be rescued from the impending danger of our sins; and with Thee as our deliverer, we may obtain our salvation. Who liveth.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of the Octave of the Dedication of our Parish Church (Nov. 26)}

\textit{\textnormal{Ant.} O how terrible is this place! Indeed this is no other but the house of God and the gate of heaven.} %PT: alleluia

\begin{rubricbox}

After the cantor intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../commemorations/dedication-2v}

\begin{rubricbox}

All \textbf{stand} for the prayer.

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Domum tuam Dómine decet sanctitúdo.\\
	\Rbar.~In longitúdinem diérum.
	
	Orémus.
	Deus, qui nobis per síngulos annos hujus sancti templi tui consecratiónis réparas diem,
	et sacris semper mystériis repræséntas incólumes:
	exáudi preces pópuli tui, et præsta;
	ut quisquis hoc templum benefícia petitúrus ingréditur,
	cuncta se impetrásse lætétur.
	Per Dóminum.
}{
	Holiness becometh Thy house, O Lord.
	\Rbar.~To the end of days.
	
	Let us pray.
	O God, who every year dost renew the day of the consecration of this Thy holy temple,
	and continuest to bring us to safety to Thy sacred mysteries,
	graciously hear the prayers of Thy people, and grant that all who enter this temple to implore Thy blessings,
	may rejoice in obtaining whatsoever they ask.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}