\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, February 12, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Sixth after Epiphany (February 13, 2022). All is said as in the main booklet until the Magnificat.}

\section*{Magnificat}

%% all the Sundays after Epiphany share the same antiphon
\gresetinitiallines{1}
\gregorioscore{../post-epiph-2/magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{../post-epiph-2/magnificat-antiphon}

\textit{\textnormal{Ant.} God hath received Israel His servant according as He spoke to Abraham and to his seed, and hath exalted the humble forever.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Preces pópuli tui, quǽsumus, Dómine, cleménter exáudi: ut, qui juste pro peccátis nostris afflígimur, pro tui nóminis glória misericórditer liberémur.
Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	O Lord, we beseech thee favourably to hear the prayers of thy people; that we, who are justly punished for our offences, may be mercifully delivered by thy goodness, for the glory of thy Name.
Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of the Seven Holy Founders of the Servite Order, Confessors (February 12)}

\textit{\textnormal{Ant.} Theirs is the kingdom of heaven, who despising a worldy life have attained the rewards of the kingdom,
and have washed their robes in the Blood of the Lamb.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~ Hi viri misericórdiæ sunt, quorum pietátes non defuérunt..\\
	\Rbar.~Semen eórum et glória eórum non derelinquétur.
	
	Orémus.
	Dómine Jesu Christe, qui ad recoléndam memóriam dolórum sanctíssimæ Genetrícis tuæ, per septem beátos Patres nova Servórum ejus família Ecclésiam tuam fecundásti: concéde propítius; ita nos eórum consociári flétibus, * ut perfruámur et gáudiis.
Qui vivis et regnas cum Deo Patre, in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	These were merciful men whose righteousness hath not been forgotten.
	\Rbar.~Their seed and their glory shall not be blotted out.
	
	Let us pray.
	Lord Jesus Christ, who, that thou mightest recall to mind the woes of thy most holy Mother, didst through the Seven blessed Fathers make thy Church herself the mother of a new household of her servants, grant unto us in mercy that we may so share their tears as to share their blessedness also.
Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.
	Through Our Lord.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-sunday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

%% TODO: make sure this is correct
\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}