\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, April 30, 2022}

 \textit{Vespers today is based on the following Sunday, which is the second after Easter.
 All is said as in the Eastertide booklet until the Little Chapter.}

\section*{Little Chapter (1 Pet. 2:21-22)}

\begin{latinenglishsection}

 \latinenglish{
	 Caríssimi: Christus passus est pro nobis,~\GreDagger\
	 vobis relínquens exémplum, ut sequámini vestígia ejus.~*
	 Qui peccátum non fecit, nec invéntus est dolus in ore ejus.
 	\Rbar.~Deo gratias.
 }{
 	Beloved: Christ suffered for us, leaving you an example that you should follow his steps.
 	Who did no sin, neither was guile found in his mouth.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/ad-coenam-agni}

{\itshape
	1. The Lamb's high banquet we await
	in snow-white robes of royal state:
	and now, the Red Sea's channel past,
	to Christ our Prince we sing at last. 
	
	2. Upon the Altar of the Cross
	His Body hath redeemed our loss:
	and tasting of his roseate Blood,
	our life is hid with Him in God. 
	
	3. That Paschal Eve God's arm was bared,
	the devastating Angel spared:
	by strength of hand our hosts went free
	from Pharaoh's ruthless tyranny. 
	
	4. Now Christ, our Paschal Lamb, is slain,
	the Lamb of God that knows no stain,
	the true Oblation offered here,
	our own unleavened Bread sincere. 
	
	5. O Thou, from whom hell's monarch flies,
	O great, O very Sacrifice,
	Thy captive people are set free,
	and endless life restored in Thee.
	
	6. For Christ, arising from the dead,
	from conquered hell victorious sped,
	and thrust the tyrant down to chains,
	and Paradise for man regains. 
	
	7. We pray Thee, King with glory decked,
	in this our Paschal joy, protect
	from all that death would fain effect
	Thy ransomed flock, Thine own elect.
	
	8. To Thee who, dead, again dost live,
	all glory Lord, Thy people give;
	all glory, as is ever meet,
	to Father and to Paraclete. Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ma(h)ne(h) no(h)bís(h)cum(h) Dó(h)mi(h)ne,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Quó(h)ni(h)am(h) ad(h)ves(h)pe(h)rá(h)scit,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Stay with us, Lord, alleluia.
\Rbar.~For it is getting towards evening, alleluia.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} I am the Shepherd of the sheep; I am the Way, the Truth, and the Life. I am the Good Shepherd, and I know Mine and Mine know Me, alleluia, alleluia.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui in Fílii tui humilitáte jacéntem mundum erexísti:
	fidélibus tuis perpétuam concéde lætítiam;
	ut, quos perpétuæ mortis eripuísti cásibus, gaudiis fácias pérfrui sempitérnis.
	Per eúmdem Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, who, by the humility of Thy Son, didst lift up a fallen world, grant unending happiness to Thy faithful:
	that those whom Thou hast snatched from the perils of endless death, Thou mayest cause to rejoice in everlasting days.
	Through the same.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\pagebreak

\section*{Commemoration of the Second Sunday after Easter}

\textit{\textnormal{Ant.} I am the Shepherd of the Sheep: I am the Way, the Truth, and the Life: I am the Good Shepherd, and know My Sheep, and am known of Mine. Alleluia, Alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--ego_sum_pastor_ovium--solesmes}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Mane nobíscum, Dómine, allelúja.\\
	\Rbar.~Quóniam advesperáscit, allelúja.
	
	Orémus.
	Deus, qui in Fílii tui humilitáte jacéntem mundum erexísti: fidélibus tuis perpétuam concéde lætítiam; ut, quos perpétuæ mortis eripuísti cásibus, gáudiis fácias pérfrui sempitérnis.
}{
	\Vbar.~Stay with us, O Lord, alleluia.
	\Rbar.~Because it is towards evening, alleluia.
	
	Let us pray.
	O God, Whose Son hath humbled Himself, and Who hast through Him raised up the whole world, grant to thy faithful people everlasting joy; and as Thou hast delivered them from the bitter pains of eternal death, make them to be glad for ever in thy presence.
}

\end{latinenglishsection}

\section*{Commemoration St. Catherine of Siena (April 30)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Diffúsa est grátia in lábiis tuis, allelúia \\
	\Rbar.~Proptérea benedíxit te Deus in ætérnum, allelúia.
	
	Orémus.
	Da, quǽsumus, omnípotens Deus: ut, qui beátæ Catharínæ Vírginis tuæ natalícia cólimus; et ánnua solemnitáte lætémur, et tantæ virtútis proficiámus exémplo.
	Per Dóminum.
}{
	Grace is poured out upon thy lips, alleluia.
	\Rbar.~Therefore God hath blessed thee forever, alleluia.
	
	Let us pray.
	Grant, we beseech thee, O Almighty God, that we which do keep the birthday of thy blessed Virgin Katharine, and do year by year renew her memorial with solemn gladness in thy presence, may likewise be conformed to the pattern of her saintly walk with thee.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain with us always. \Rbar.~Amen.
}

\end{latinenglishsection}

\end{document}