\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Second Vespers of the Transfiguration of Our Lord Jesus Christ}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently.
Then, when the Officiant stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

%% FIRST PSALM

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} Jesus took Peter, * and James, and John his brother, and brought them up into an high mountain apart and was transfigured before them.
 \textnormal{Ps.} The Lord said to my Lord: * Sit thou at my right hand:}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk. This is done for the remaining Psalms.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-1} %%

}{
	\input{../psalms/english/ps109} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red} The antiphon is repeated: \textit{Assúmpsit Jesus\dots}} %%

\end{rubricbox}

%% SECOND PSALM

\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} His Face did shine as the sun, * and His raiment exceeding white as snow, alleluia.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; * in the council of the just, and in the congregation.}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps110-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps110-8} %%

}{
	\input{../psalms/english/ps110} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Resplénduit\dots}} %%

\end{rubricbox}

%% THIRD PSALM

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} And, behold, there appeared unto them Moses and Elias, * talking with Jesus.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: * he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps111-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-4-E-A-d} %%

}{
	\input{../psalms/english/ps111} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Et ecce\dots}, then all \textbf{stand}} %%

\end{rubricbox}

%% FOURTH PSALM

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} Then answered Peter, * and said unto Jesus: Lord, it is good for us to be here.
 \textnormal{Ps.} Praise the Lord, ye children: * praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-1} %%

}{
	\input{../psalms/english/ps112} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Respóndens autem Petrus\dots}} %%

\end{rubricbox}

\vfill\pagebreak

%% FIFTH PSALM

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} While he yet spake, * behold, a bright cloud overshadowed them.
 \textnormal{Ps.} Praise the Lord, all ye nations: * praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps116-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps116-6} %%

}{
	\input{../psalms/english/ps116} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Adhuc eo loquénte\dots}, then all \textbf{stand}}

\end{rubricbox}

\section*{Little Chapter (Phil 3:20-21)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Salvatórem exspectámus Dóminum nostrum Jesum Christum,\GreDagger\
	qui\\ reformábit corpus humilitátis nostræ configurátum córpori claritátis suæ.
	\Rbar.~Deo grátias.
}{
	We look for the Saviour, the Lord Jesus Christ, Who shall change our vile body, that it may be fashioned like unto His glorious Body.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/quicumque-christum--solesmes}

{\itshape

	1. All ye who would the Christ descry,
	Lift up your eyes to him on high:
	There mortal gaze hath strength to see
	The token of his majesty.
	
	2. A wondrous sign we there behold,
	That knows not death nor groweth old,
	Sublime, most high, that cannot fade,
	That was ere earth and heaven were made.
	
	3. Here is the King the gentiles fear,
	The Jews' most mighty King is here
	Promised to Abraham of yore,
	And to his seed forevermore.
	
	4. 'Tis he the prophets' words foretold,
	And by their signs shown forth of old;
	The Father's witness hath ordained
	That we should hear with faith unfeigned.
	
	5. Jesu, to thee our praise we pay,
	To little ones revealed today,
	With Father and blest Spirit One
	Until the ages' course is done.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Glo(h)ri(h)ó(h)sus(h) ap(h)pa(h)ru(h)í(h)sti(h) in(h) con(h)spéc(h)tu(h) Dó(h)mi(f)ni(f.).(::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Prop(h)té(h)re(h)a(h) de(h)có(h)rem(h) ín(h)du(h)it(h) te(h) Dó(h)mi(h)nus(f.). (::)
}

\textit{\Vbar.~Thou wast manifested in thy glory in the Presence of the Lord.
\Rbar.~Therefore the Lord hath clothed thee with majesty.}

\vfill\pagebreak

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} And when the disciples heard it, they fell on their face, * and were sore afraid and Jesus came and touched them, and said unto them: Arise, and be not afraid, alleluia.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

{\color{red}After the leader intones up to the asterisk, all \textbf{sit} and join in singing.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only-2v}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation-2v}

 \begin{latinenglishsection}

\latinenglish{	
3. \textit{Quia} respéxit humilitátem ancíllæ \textbf{su}æ:~*
	ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.

4. \textit{Quia} fecit mihi magna qui \textbf{pot}ens est:~* 
	et sanctum \textit{no}\textit{men} \textbf{e}jus.

5. \textit{Et\ mi}sericórdia ejus a progénie in pro\textbf{gé}nies~*
	timén\textit{ti}\textit{bus} \textbf{e}um.

6. \textit{Fecit} poténtiam in bráchio \textbf{su}o:~*
	dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.

7. \textit{Depó}suit poténtes de \textbf{se}de,~*
	et exal\textit{tá}\textit{vit} \textbf{hú}miles.

8. \textit{Esu}riéntes implévit \textbf{bo}nis:~*
	et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.

9. \textit{Suscé}pit Israël púerum \textbf{su}um,~*
	recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.

10. \textit{Sicut} locútus est ad patres \textbf{nos}tros,~*
	Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~* 
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.


\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Et audiéntes\dots}, %%
then \textbf{stand} for the prayer.}

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui fídei sacraménta in Unigéniti tui gloriósa Transfiguratióne patrum testimónio roborásti, et adoptiónem filiórum perféctam, voce delápsa in nube lúcida, mirabíliter præsignásti: concéde propítius; ut ipsíus Regis glóriæ nos coherédes effícias, et ejúsdem glóriæ tríbuas esse consórtes.
	Per eúmdem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	\Vbar.~Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who, in the glorious Transfiguration of thine Only Begotten Son didst attest the mysteries of the Faith by the witness of the Fathers, and didst wonderfully signify by a voice out of a bright cloud, the adoption of sons, mercifully grant unto us to be made co-heirs with the very King of glory, and bestow upon us a partaking of His glory.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of St. Cajetan, Confessor (August 7)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\
	\Rbar.~Stolam glóriæ induit eum. 
	
	Orémus.
	Deus, qui beáto Cajetáno Confessóri tuo apostólicam vivéndi formam imitári tribuísti: da nobis, ejus intercessióne et exémplo, in te semper confídere et sola cæléstia desideráre.
}{
	\Vbar.~The Lord loved him and adorned him.
	\Rbar.~He clothed him with a robe of glory.
	
	Let us pray.
	O God, who didst give grace unto thy blessed confessor Cajetan to order his life on the pattern of thine Apostles, grant unto us through his prayers, and after his example ever to put all our trust in thee, and to seek only heavenly things.
}

\end{latinenglishsection}

\section*{Commemoration of the Ninth Sunday after Pentecost}

\textit{\textnormal{Ant.}  I dwell in the highest places, and my throne is in a cloudy pillar.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--ego_in_altissimis_habito--solesmes_1934}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Vespertína orátio ascéndat ad te, Dómine.\\
	\Rbar.~Et descéndat super nos misericórdia tua.
	
	Orémus.
	Páteant aures misericórdiæ tuæ, Dómine, précibus supplicántium: et, ut peténtibus desideráta concédas; fac eos, quæ tibi sunt plácita, postuláre.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	\Vbar.~Let the evening prayer ascend unto thee, O Lord.
	\Rbar.~And let there descend upon us thy mercy.
	
	Let us pray.
	Let thy merciful ears, O Lord, be open to the prayers of thy humble servants; and, that they may obtain their petitions, make them to ask such things as shall please thee.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-double}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\section*{Marian Anthem}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/salve-regina-simple}

{\itshape
	Hail, holy Queen, Mother of mercy, our life, our sweetness and our hope. 
	To thee do we cry, poor banished children of Eve. 
	To thee do we send up our sighs, mourning and weeping in this valley of tears. 
	Turn, then, most gracious advocate, thine eyes of mercy toward us, 
	and after this, our exile, show unto us the blessed fruit of thy womb, Jesus. 
	O clement, O loving, O sweet Virgin Mary.
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Ora pro nobis, sancta Dei Genitrix.\\
	\Rbar. Ut digni efficamur promissionibus Christi.
	
	Orémus.
	Omnipotens sempiterne Deus, qui gloriosae Virginis Matris Mariae corpus et animam, ut dignum Filii tui habitaculum effici mereretur, Spiritu Sancto cooperante, praeparasti, da, ut cuius commemoratione laetamur; eius pia intercessione, ab instantibus malis et a morte perpetua liberemur.
	Per eúmdem Christum Dóminum nostrum.
	\Rbar.~Amen.
}{
	\Vbar.~Pray for us, O holy Mother of God.
	\Rbar.~That we may be made worthy of the promises of Christ.
	
	Let us pray.
	Almighty and everlasting God, Who by the working of the Holy Spirit didst prepare both body and soul of the glorious Virgin Mother, Mary, that she might deserve to be made a worthy dwelling for Thy Son, grant that we who rejoice in her memory, may, by her loving intercession, be delivered from present evils and from lasting death, through the same Christ our Lord.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}