\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, July 9, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fifth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 11:33)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
 	O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei, ~\GreDagger\
 	quam incomprehensibília sunt judí\textit{cia} \textbf{e}jus, et\\ investigábiles viæ ejus!
	 
 	\Rbar.~Deo grátias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) as(h)cén(h)dat(h) ad(h) te(h), Dó(h)mi(f)ne(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a(f.). (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} All generations shall call me blessed, * for God hath regarded the lowliness of His hand-maiden.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui diligéntibus te bona invisibília præparásti:~\GreDagger infúnde córdibus nostris tui amóris afféctum; ut te in ómnibus et super ómnia \textit{dili}\textbf{gén}tes, promissiónes tuas, quæ omne desidérium súperant, consequámur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	O God, Who hast prepared for them that love thee such good things as pass man's understanding; pour into our hearts such love toward thee, that we, loving thee in all things and above all things, may obtain thy promises, which exceed all that we can desire.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of The Seven Holy Brothers, Marytrs, and Sts. Rufina and Secunda, Virgins and Martyrs (July 10)}

\textit{\textnormal{Ant.} Theirs is the kingdom of heaven, who despising a worldy life have attained the rewards of the kingdom,
and have washed their robes in the Blood of the Lamb.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/many-martyrs-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Lætámini in Dómino, et exsultáte justi.\\
	\Rbar.~Et gloriámini omnes recti corde.
	
	Orémus.
	Præsta, qu\'{\ae}sumus, omnípotens Deus: ut, qui gloriósos Mártyres fortes in sua confessióne cognóvimus, pios apud te in nostra intercessióne sentiámus.
}{
	Be glad in the Lord and rejoice, ye just.
	\Rbar.~And exult, all ye upright of heart.
	
	Let us pray.
	Grant, we beseech thee, almighty God, that we, who know how courageously thy holy Martyrs have confessed the faith, may experience their goodness as they intercede for us with thee.
}

\end{latinenglishsection}

\section*{Suffrage of the Saints}

\textit{\textnormal{Ant.} May the Blessed Mother of God, the Virgin Mary, and all the Saints, intercede for us before the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/suffrage-saints}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Mirificávit Dóminus Sanctos suos.\\
	\Rbar. Et exaudívit eos clamántes ad se.
	
	Orémus.
	A cunctis nos, qu\'{\ae}sumus Dómine, mentis et córporis defénde perículis:
	et intercedénte beáta et gloriósa semper Vírgine Dei Genitríce María,
	cum beáto Joseph, beátis Apóstolis tuis Petro et Paulo, atque beáto Thoma apóstolo et ómnibus Sanctis,
	salútem nobis tríbue benígnus et pacem;
	ut destrúctis adversitátibus et erróribus univérsis,
	Ecclésia tua secúra tibi sérviat libertáte.
	Per eúmdem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	The Lord hath made His holy ones wonderful.
	\Rbar.~And He hath heard them when they cried unto Him.
	
	Let us pray.
	From all perils of soul and body defend us, O Lord, we beseech Thee,
	and by the intercession of blessed and glorious ever Virgin Mary, Mother of God,
	of blessed Joseph, of thy blessed apostles Peter and Paul, and of blessed apostle Thomas and all the Saints,
	graciously grant us safety and peace that all adversities and errors being overcome,
	thy Church may serve Thee in security and freedom. Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}