\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of Trinity Sunday}

\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} O wondrous exchange! The Creator of mankind, having assumed a living body, deigned to be born of a Virgin, and having become man without man's aid, enriched us with His divinity.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-1}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Gloria\dots}, then all \textbf{stand}.

\end{rubricbox}


\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} By Thy ineffable birth of a Virgin, the Scriptures were fulfilled; like rain upon the fleece didst Thou descend to save mankind: we praise Thee, our God.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-2}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Laus et\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} In the bush which Moses saw burning yet unconsumed, we recognize the preservation of thy glorious virginity; O Mother of God, intercede for us.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-3}
	
}{
	\input{../psalms/english/ps111}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{Gloria laudis\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} The Root of Jesse hath blossomed; the Star hath risen out of Jacob; the Virgin hath borne the Savior; we praise Thee, our God.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-4-E-A-d}
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Laus Deo Patri\dots}, then all \textbf{stand}.

\end{rubricbox}

\pagebreak

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} Behold, Mary hath given birth to the Savior, Whom John seeing, exclaimed: Behold the Lamb of God; behold Him who taketh away the sins of the world, alleluia.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps116-5}
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Ex quo omnia\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Wisdom 5:1)}

\begin{latinenglishsection}

\latinenglish{
	Stabunt justi in magna constántia~\GreDagger\
	advérsus eos qui se angustiavérunt,~*
	et qui abstulérunt labóres eórum.
	\Rbar.~Deo grátias.
}{
	The just shall stand with great constancy against those that have afflicted them, and taken away their labours.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.

	2. At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	3. To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Be(h)ne(h)di(h)ca(h)mus(h) Pa(h)trem(h) et(h) Fí(h)li(h)um(h) cum(h) San(h)cto(h) Spí(h)ri(f)tu(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Lau(h)dé(h)mus(h) et(h) su(h)pe(h)rex(h)al(h)té(h)mus(h) eum(h) in(h) sæ(h)cu(h)la(f.). (::)
}

\textit{\Vbar.~Bless we the Father, and the Son, and the Holy Ghost.
\Rbar.~Let us praise Him and exalt Him above all for ever..}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Thanks be * to thee, O God, thanks be unto thee, O True and One Trinity, O one almighty Deity, Holy and One Unity.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-1}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon: \textit{Gratias\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens sempitérne Deus, qui dedísti fámulis tuis in confessióne veræ fídei, ætérnæ Trinitátis glóriam agnóscere, et in poténtia majestátis adoráre Unitátem: quǽsumus; ut, ejúsdem fídei firmitáte, ab ómnibus semper muniámur advérsis.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{æ}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and everlasting God, you have given thy servants in the confession of the true faith, to acknowledge the glory of the Eternal Trinity, and to adore the unity in the power of thy Majesty; grant, that by steadfastness in this faith we may ever be defended from all adversities.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\section*{Commemoration of the Sunday}

\textit{\textnormal{Ant.} But the father said to his servants: bring forth quickly the first robe, and put it on him, and put a ring on his finger, and shoes on his feet.}

\gresetinitiallines{1}
\gregorioscore{an--loquere_domine--solesmes_1961}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Vespertína orátio ascéndat ad te, Dómine.\\
	\Rbar.~Et descéndat super nos misericórdia tua.
	
	Orémus.
	Deus, in te sperántium fortitúdo, adésto propítius invocatiónibus nostris: et, quia sine te nihil potest mortális infírmitas, præsta auxílium grátiæ tuæ; ut, in exsequéndis mandátis tuis, et voluntáte tibi et actióne placeámus.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{æ}cula sæculórum.
	\Rbar.~Amen.
}{
	\Vbar.~Let the evening prayer ascend unto thee, O Lord.
	\Rbar.~And let there descend upon us thy mercy.
	
	Let us pray.
	O God, the Strength of all them that put their trust in thee, mercifully accept our prayers, and because through the weakness of our mortal nature we can do no good thing without thee, grant us the help of thy grace, that in keeping of thy commandments we may please thee, both in will and deed.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen.  \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\pagebreak

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solem}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 224 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}