\documentclass{../../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Friday, February 3, 2023}

 \textit{Vespers today is that of St. Casimir, of semidouble rank.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Sirach 31:8-9)}

\begin{latinenglishsection}

 \latinenglish{
 	Beátus vir, qui invéntus est sine mácula, et qui post aurum non ábiit,~\GreDagger\ 
 	nec sperávit in pecúnia et thesáuris.~*
 	Quis est hic, et laudábimus eum? Fecit enim mirabília in vita sua.
 	\Rbar.~Deo gratias.
 }{
	Blessed is the man that is found without blemish: and that hath not gone after gold, nor put his trust in money nor in treasures. Who is he, and we will praise him? for he hath done wonderful things in his life.
 	\Rbar.~Thanks be to God.
 }
 
 \end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../../hymns/iste_confessor}

%\vfill
%\pagebreak

{\itshape

	1. This the confessor of the Lord, whose triumph
	Now all the faithful celebrate, with gladness
	On this feast day year by year receiveth
	Merited honors.
	
	2. Saintly and prudent, modest in behaviour,
	Peaceful and sober, chaste was he, and lowly.
	While that life’s vigour, coursing through his members,
	Quickened his being.
	
	3. Sick ones of old time, to his tomb resorting,
	Sorely by ailments manifold afflicted,
	Oft-times have welcomed health and strength returning,
	At his petition.
	
	4. Whence we in chorus gladly do him honour,
	Chanting his praises with devout affection,
	That in his merits we may have a portion,
	Now and forever.
	
	5. His be the glory, power and salvation,
	Who over all things reigneth in the highest,
	Earth’s mighty fabric ruling and directing,
	Onely and Trinal.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>.  A(h)má(h)vit(h) e(h)um(h) Dó(h)mi(h)nus(h), et(h) or(h)ná(h)vit(h) e(h)um.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Sto(h)lam(h) gló(h)ri(h)æ(h) ín(h)du(h)it(h) e(h)um.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~The Lord loved him and beautified him.
\Rbar.~He clothed him with a robe of glory.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../../psalms/magnificat-1}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.}  I will liken him unto a wise man, who built his house upon a rock.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the Officiant.

\end{rubricbox}

\section*{Collect and conclusion}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui inter regáles delícias et mundi illécebras sanctum Casimírum virtúte constántiæ roborásti:
	qu\'{\ae}sumus; ut ejus intercessióne fidéles tui terréna despíciant,
	et ad cæléstia semper aspírent.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.\\
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, who amidst the delights of royalty and the blandishments of the world, didst strengthen holy Casimir with the virtue of constancy, grant, we beseech Thee, that by his intercession Thy faithful may despise earthly things, and ever aspire to those of heaven.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of the Ember Day}

\textit{\textnormal{Ant.} He that made me whole, He it is that said to me: Take up thy bed and walk in peace.}

\gresetinitiallines{1}
\gregorioscore{comm-ember-friday}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Angelis suis Deus mandávit de te.\\
	\Rbar.~Ut custódiant te in ómnibus viis tuis.
	
	Orémus.
	Esto, Dómine, propítius plebi tuæ: et quam tibi facis esse devótam, benígno réfove miserátus auxílio.
}{
	God has given His Angels charge over thee.
	\Rbar.~To keep thee in all thy ways.
	
	Let us pray.
	Be gracious, O Lord, unto Thy people, and even as Thou makest them devoted to Thee,
	so mercifully revive them with Thy kind assistance.
}

\end{latinenglishsection}

%\vfill\pagebreak

\section*{Commemoration of St Lucius, Pope and Martyr}

\textit{\textnormal{Ant.} This Saint fought even unto death for the law of his God, and feared not the words of the wicked; for he was founded upon a firm rock.}

\gresetinitiallines{1}
\gregorioscore{../../commemorations/one-martyr-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Glória et honóre coronásti eum, Dómine.\\
	\Rbar.~Et constituísti eum super ópera mánuum tuárum.
	
	Orémus.
	Gregem tuum, Pastor ætérne, placátus inténde:
	et per beátum Lúcium Mártyrem tuum atque Summum Pontíficum,
	perpétua protectióne custódi; quem totíus Ecclésiæ præstitísti esse pastórem.
}{
	Thou hast crowned him with glory and honor, O Lord.
	\Rbar.~And hast set him over the works of Thy hands.
	
	Let us pray.
	Eternal shepherd, look kindly upon Thy flock, and preserve it with a continual protection, by the blessed Lucius
	Thy martyr and Sovereign Pontiff, whom Thou hast made the Shepherd of the whole Church.
}

\end{latinenglishsection}

\section*{Suffrage of the Saints}

\textit{\textnormal{Ant.} May the Blessed Mother of God, the Virgin Mary, and all the Saints, intercede for us before the Lord.}

\gresetinitiallines{1}
\gregorioscore{../../commemorations/suffrage-saints}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Mirificávit Dóminus Sanctos suos.\\
	\Rbar. Et exaudívit eos clamántes ad se.
	
	Orémus.
	A cunctis nos, qu\'{\ae}sumus Dómine, mentis et córporis defénde perículis:
	et intercedénte beáta et gloriósa semper Vírgine Dei Genitríce María,
	cum beáto Joseph, beátis Apóstolis tuis Petro et Paulo, atque beáto Thoma apóstolo et ómnibus Sanctis,
	salútem nobis tríbue benígnus et pacem;
	ut destrúctis adversitátibus et erróribus univérsis,
	Ecclésia tua secúra tibi sérviat libertáte.
	Per Dóminum nostrum Jesum Christum, Fílium tuum:
	qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.\\
	\Rbar.~Amen.
}{
	The Lord hath made His holy ones wonderful.
	\Rbar.~And He hath heard them when they cried unto Him.
	
	Let us pray.
	From all perils of soul and body defend us, O Lord, we beseech Thee,
	and by the intercession of blessed and glorious ever Virgin Mary, Mother of God,
	of blessed Joseph, of thy blessed apostles Peter and Paul, and of blessed apostle Thomas and all the Saints,
	graciously grant us safety and peace that all adversities and errors being overcome,
	thy Church may serve Thee in security and freedom. Through our Lord. \Rbar.~Amen.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../../common/benedicamus-semidouble}

%\vfill\pagebreak

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	\Vbar. May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	\Vbar. May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian Anthem \textit{Ave Regina Cælorum} follows: No. 188 in the \textit{Traditional Roman Hymnal}.

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Vbar.~Amen.
}
\end{latinenglishsection}

%\vfill\pagebreak

%\section*{Marian Anthem}
%
%\gresetinitiallines{1}
%\gregorioscore{../../marian-anthems/ave-regina-caelorum-simple}
%
%{\itshape
%	Hail, O Queen of Heaven enthroned.
%	Hail, by angels mistress owned.
%	Root of Jesse, Gate of Morn
%	whence the world's true light was born.
%	
%	Glorious Virgin, Joy to thee,
%	loveliest whom in heaven they see;
%	fairest thou, where all are fair,
%	plead with Christ our souls to spare.
%}
%
%\begin{latinenglishsection}
%
%\latinenglish{
%	\Vbar. Dignáre me laudáre te, Virgo sacráta.\\
%	\Rbar. Da mihi virtútem contra hostes tuos.
%	
%	Orémus.
%	Concéde, miséricors Deus, fragilitáti nostræ præsídium:
%	ut qui sanctæ Dei Genitrícis memóriam ágimus;
%	intercessiónis ejus auxílio, a nostris iniquitátibus resurgámus.
%	Per eúmdem Christum Dóminum nostrum.
%	\Rbar.~Amen.
%}{
%	Vouchsafe that I may praise thee, O sacred Virgin.
%	\Rbar.~Give me stength against thine enemies.
%	
%	Let us pray.
%	Grant, O merciful God, to our weak natures Thy protection,
%	that we who commemorate the holy Mother of God may,
%	by the help of her intercession, arise from our iniquities.
%	Through the same Christ our Lord.
%	\Rbar.~Amen.
%}
%\end{latinenglishsection}
%
%\begin{latinenglishsection}
%
%\latinenglish{
%	\Vbar. Divínum auxílium máneat semper nobíscum.\\
%	\Rbar. Amen.
%}{
%	May the divine assistance remain always with us.
%	\Vbar.~Amen.
%}
%\end{latinenglishsection}
%
%\begin{rubricbox}
%
%After the Office, all \textbf{kneel} and pray in silence for a time.
%
%\end{rubricbox}

\end{document}