\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, March 12, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Second of Lent.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (1 Thess. 4:2)}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Rogámus vos et obsecrámus in Dómino Jesu:~\GreDagger\
	 ut, quemádmodum accepístis a nobis, quómodo opórteat vos ambuláre et placére Deo:~*
	 sic et ambulétis, ut abundétis magis.
 	\Rbar.~Deo gratias.
 }{
	Brethren: we pray and beseech you in the Lord Jesus,
 	that as you have received from us, how you ought to walk, and to please God,
 	so also you would walk, that you may abound the more. 
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/audi-benigne-conditor}

\vfill
\pagebreak

{\itshape

1. O merciful Creator, hear!
To us in pity bow Thine ear:
accept the tearful prayer we raise
in this our fast of forty days.

2. Our hearts are open, Lord, to Thee:
Thou knowest our infirmity;
pour out on all who seek Thy face
abundance of Thy pardoning grace.

3. Our sins are many, this we know;
spare us, good Lord, Thy mercy show;
and for the honor of Thy name
our fainting souls to life reclaim.

4. Give us self-control that springs
from discipline of outward things,
that fasting inward secretly
the soul may purely dwell with Thee.

5. We pray Thee, Holy Trinity,
one God, unchanging Unity,
that we from this our abstinence
may reap the fruits of penitence. Amen.

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. An(h)ge(h)lis(h) su(h)is(h) De(h)us(h) man(h)dá(h)vit(h) de(h) te.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Ut(h) cus(h)tó(h)di(h)ant(h) te(h) in(h) óm(h)ni(h)bus(h) vi(h)is tu(h)is.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~He hath given his angels charge over thee.
\Rbar.~To keep thee in all thy ways.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Tell the vision that ye have seen to no man, until the Son of man be risen again from the dead.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui cónspicis omni nos virtúte destítui: intérius exteriúsque custódi;
	ut ab ómnibus adversitátibus muniámur in córpore,
	et a pravis cogitatiónibus mundémur in mente.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who seest how we are deprived of all strength, guard us inwardly and outwardly,
	that in body we may be protected against all misfortunes, and in mind cleansed of evil thoughts.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of Pope St. Gregory the Great, Dialogist and Doctor of the Church (March 12)}

\textit{\textnormal{Ant.} O right excellent Teacher, light of the Holy Church, St. Gregory, Lover of God's law, pray for us to the Son of God} 

\gresetinitiallines{1}
\gregorioscore{an--o_doctor_gregorii_optime--solesmes_1961}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ 
	\Rbar.~Et osténdit illi regnum Dei. 
	
	Orémus.
	Deus, qui ánimæ fámuli tui Gregórii ætérnæ beatitúdinis pr\'{\ae}mia contulísti: concéde propítius; ut, qui peccatórum nostrórum póndere prémimur, ejus apud te précibus sublevémur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	The Lord guided the just in right paths. 
	\Rbar.~And showed him the kingdom of God. 
	
	Let us pray.
O God, who hast blessed the soul of thy servant Gregory with an everlasting blessing, mercifully grant that we, who groan under the burden of our sins, may by his prayers be relieved.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}