\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Second Vespers of St. Joseph}

%% this heading doesn't quite fit with the "Laus tibi Domine" version of Deus in adjutorium
%\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-lent}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} The parents of Jesus were wont to go every year to Jerusalem at the feast of the Passover.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. Donec ponam ini\textbf{mí}cos \textbf{tu}os,~*
	scabéllum pe\textit{dum} \textit{tu}\textbf{ó}rum.

3. Virgam virtútis tuæ emíttet Dómi\textbf{nus} ex \textbf{Si}on:~*
	domináre in médio inimicó\textit{rum} \textit{tu}\textbf{ó}rum.

4. Tecum princípium in die virtútis tuæ in splendóri\textbf{bus} sanc\textbf{tó}rum:~*
	ex útero ante lucíferum \textit{gé}\textit{nu}\textbf{i} te.

5. Jurávit Dóminus, et non p{\oe}ni\textbf{té}\-bit \textbf{e}um:~*
	Tu es sacérdos in ætérnum secúndum órdi\textit{nem} \textit{Mel}\textbf{chí}se\-dech.

6. Dóminus a \textbf{dex}tris \textbf{tu}is,~*
	confrégit in die iræ \textit{su}\textit{æ} \textbf{re}ges.

7. Judicábit in natiónibus, im\textbf{plé}bit ru\textbf{í}nas:~*
	conquassábit cápita in ter\textit{ra} \textit{mul}\textbf{tó}rum.

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

8. De torrénte in \textbf{vi}a \textbf{bi}bet:~*
	proptérea exal\textit{tá}\textit{bit} \textbf{ca}put.

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~*
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Ibant\dots}, then all \textbf{stand}.

\end{rubricbox}


\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} When they were returning, the Boy Jesus remained in Jerusalem, and His parents did not know it.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
2. Magna ópera \textbf{Dó}mini:~* exquisíta in omnes voluntá\textit{tes} \textbf{e}jus.

3. Conféssio et magnificéntia opus \textbf{e}jus:~* et justítia ejus manet in s\'{\ae}cu\textit{lum} \textbf{s\'{\ae}}culi.

4. Memóriam fecit mirabílium suórum,~\GreDagger\ miséricors et miserátor \textbf{Dó}minus:~* escam dedit timén\textit{ti}\textbf{bus} se.

5. Memor erit in s\'{\ae}culum testaménti \textbf{su}i:~* virtútem óperum suórum annuntiábit pópu\textit{lo} \textbf{su}o:

6. Ut det illis hereditátem \textbf{gén}ti\-um:~* ópera mánuum ejus véritas, et \textit{ju}\textbf{dí}cium.

7. Fidélia ómnia mandáta ejus:~\GreDagger\ confirmáta in s\'{\ae}culum \textbf{s\'{\ae}}culi,~* facta in veritáte et æ\textit{qui}\textbf{tá}te.

8. Redemptiónem misit pópulo \textbf{su}o:~* mandávit in ætérnum testamén\textit{tum} \textbf{su}um.

9. Sanctum, et terríbile nomen \textbf{e}jus:~* inítium sapiéntiæ ti\textit{mor} \textbf{Dó}\-mini.

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

10. Intelléctus bonus ómnibus faciéntibus \textbf{e}um:~* laudátio ejus manet in s\'{\ae}cu\textit{lum} \textbf{s\'{\ae}}culi.

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~* et Spirítu\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~* et in s\'{\ae}cula sæculó\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Cum redirent\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} Not finding Jesus, they returned to Jerusalem seeking Him, and after three days they found Him in the temple, sitting in the midst of the doctors, hearing them and asking them questions.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. Potens in terra erit \textbf{se}men \textbf{e}jus:~* generátio rectórum be\textbf{ne}di\textbf{cé}tur.

3. Glória, et divítiæ in \textbf{do}mo \textbf{e}jus:~* et justítia ejus manet in \textbf{s\'{\ae}}culum \textbf{s\'{\ae}}culi.

4. Exórtum est in ténebris \textbf{lu}men \textbf{rec}tis:~* miséricors, et mise\textbf{rá}tor, et \textbf{jus}tus.

5. Jucúndus homo qui miserétur et cómmodat,~\GreDagger\ dispónet sermónes suos \textbf{in} ju\textbf{dí}\textbf{ci}o:~* quia in ætérnum non \textbf{com}mo\textbf{vé}bitur.

6. In memória ætérna \textbf{e}rit \textbf{jus}tus:~* ab auditióne mala \textbf{non} ti\textbf{mé}bit.

7. Parátum cor ejus speráre in Dómino,~\GreDagger\ confirmátum \textbf{est} cor \textbf{e}jus:~* non commovébitur donec despíciat ini\textbf{mí}cos \textbf{su}os.

8. Dispérsit, dedit paupéribus:~\GreDagger\ justítia ejus manet in \textbf{s\'{\ae}}culum \textbf{s\'{\ae}}\-\textbf{cu}li,~* cornu ejus exaltábi\textbf{tur} in \textbf{gló}ria.

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}\textbf{li}o,~* et Spi\textbf{rí}tui \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~* et in s\'{\ae}cula sæcu\textbf{ló}rum. \textbf{A}men.
}{
2. His seed shall be mighty upon earth: the generation of the righteous shall be blessed.

3. Glory and wealth shall be in his house: and his justice remaineth for ever and ever.

4. To the righteous a light is risen up in darkness: he is merciful, and compassionate and just.

5. Acceptable is the man that sheweth mercy and lendeth: he shall order his words with judgment:
because he shall not be moved for ever.

6. The just shall be in everlasting remembrance: he shall not fear the evil hearing.

7. His heart is ready to hope in the Lord: his heart is strengthened, he shall not be moved until he look over his enemies.

8. He hath distributed, he hath given to the poor: his justice remaineth for ever and ever: his horn shall be exalted in glory.

9. The wicked shall see, and shall be angry, he shall gnash with his teeth and pine away: the desire of the wicked shall perish. 

Glory be.
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Non invenientes\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} And His Mother said to Him: Son, why hast Thou done so to us? Behold, Thy father and I have been seeking Thee sorrowing.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{

2. \textit{(bow)} Sit nomen Dómini \textit{be}\textit{ne}\-\textbf{díc}tum,~* ex hoc nunc, et \textit{us}\textit{que} \textit{in} \textbf{s\'{\ae}}\textbf{cu}lum.

3. A solis ortu usque \textit{ad} \textit{oc}\textbf{cá}sum,~* laudábi\textit{le} \textit{no}\textit{men} \textbf{Dó}\textbf{mi}ni.

4. Excélsus super omnes \textit{gen}\textit{tes} \textbf{Dó}minus,~* et super cælos \textit{gló}\textit{ri}\textit{a} \textbf{e}jus.

5. Quis sicut Dóminus, Deus noster, qui in \textit{al}\textit{tis} \textbf{há}bitat,~* et humília réspicit in cæ\textit{lo} \textit{et} \textit{in} \textbf{ter}ra?

6. Súscitans a \textit{ter}\textit{ra} \textbf{ín}opem,~* et de stércore \textit{é}\textit{ri}\textit{gens} \textbf{páu}\textbf{pe}rem:

7. Ut cóllocet eum \textit{cum} \textit{prin}\textbf{cí}pibus,~* cum princípibus \textit{pó}\textit{pu}\textit{li} \textbf{su}i.

8. Qui habitáre facit stéri\textit{lem} \textit{in} \textbf{do}mo,~* matrem fili\textit{ó}\textit{rum} \textit{læ}\textbf{tán}tem.
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}


\textit{(bow)} Glória Pa\textit{tri}, \textit{et} \textbf{Fí}lio,~* et Spi\textit{rí}\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textit{nunc}, \textit{et} \textbf{sem}per,~* et in s\'{\ae}cula sæ\textit{cu}\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Dixit Mater ejus ad illum\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} Jesus went down with them and came to Nazareth, and was subject to them.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon-only}

\vfill\pagebreak

\gresetinitiallines{0}
\gregorioscore{ps116-intonation}

 \begin{latinenglishsection}

\latinenglish{

2. Quóniam confirmáta est super nos misericórdia \textbf{e}jus:~*
	et véritas Dómini manet \textit{in} \textit{æ}\textbf{tér}num.

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~*
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Descendit\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Proverbs 28:20, 27:18)}

\begin{latinenglishsection}

\latinenglish{
	Vir fidélis multum laudábitur.~*
	Et qui custos est Dómini sui, glorificábitur.
	\Rbar.~Deo grátias.
}{
	A faithful man shall be much praised, and he that is the keeper of his Lord, shall be glorified.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/te-joseph-celebrent}

{\itshape

   1.  Let Angels chant thy praise, pure spouse of purest Bride,
    While Christendom’s sweet choirs the gladsome strains repeat,
    To tell thy wondrous fame, to raise the pealing hymn,
    Wherewith we all they glory greet.
    
   2.  When doubts and bitter fears thy heavy heart oppressed,
    And filled thy righteous soul with sorrow and dismay,
    An Angel quickly came, the wondrous secret told,
    And drove thy anxious griefs away.
    
    3. Thy arms thy new-born Lord, with tender joy embrace;
    Him then to Egypt’s Land thy watchful care doth bring;
    Him in the Temple’s courts once lost thou dost regain,
    And 'mid thy tears dost greet thy King.
    
   4.  Not till death’s pangs are o’er do others gain their crown,
    But, Joseph, unto thee the blessed lot was given
    While life did yet endure, thy God to see and know,
    As do the Saints above in heaven.
    
    5. Grant us, great Trinity, for Joseph’s holy sake,
    In highest bliss and love, above the stars to reign,
    That we in joy with him may praise our loving God,
    And sing our glad eternal strain. 
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Gló(h)ri(h)a(h) et(h) di(h)ví(h)ti(h)æ(h) in(h) do(h)mo(h) e(h)jus.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) jus(h)tí(h)ti(h)a(h) e(h)jus(h) ma(h)net(h) in(h) s<sp>'ae</sp>(h)cu(h)lum(h) s<sp>'ae</sp>(h)cu(h)li.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Glory and riches are in his house.
\Rbar.~And his jutice endureth forever.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Behold a faithful and prudent servant, whom the Lord has set over His household.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
3. Quia respéxit humilitátem ancíllæ \textbf{su}æ:~*
	ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.

4. Quia fecit mihi magna qui \textbf{pot}ens est:~* 
	et sanctum \textit{no}\textit{men} \textbf{e}jus.

5. Et misericórdia ejus a progénie in pro\textbf{gé}nies~*
	timén\textit{ti}\textit{bus} \textbf{e}um.

6. Fecit poténtiam in bráchio \textbf{su}o:~*
	dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.

7. Depósuit poténtes de \textbf{se}de,~*
	et exal\textit{tá}\textit{vit} \textbf{hú}miles.

8. Esuriéntes implévit \textbf{bo}nis:~*
	et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.

9. Suscépit Israël púerum \textbf{su}um,~*
	recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.

10. Sicut locútus est ad patres \textbf{nos}tros,~*
	Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~* 
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.


\begin{rubricbox}

All \textbf{sit} and repeat the antiphon: \textit{Non turbetur\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Sanctíssimæ Genitrícis tuæ Sponsi, qu\'{\ae}sumus Dómine, méritis adjuvémur:
	ut quod possibílitas nostra non óbtinet, ejus nobis intercessióne donétur.
	Qui vivis.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	We beseech Thee, O Lord, that we may be heped by the merits of the Spouse of Thy most holy Mother,
	so that what we cannot obtain of ourselves, may be given to us through his intercession.
	Who livest.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\section*{Commemoration of the Sunday}

\textit{\textnormal{Ant.} But the father said to his servants: bring forth quickly the first robe, and put it on him, and put a ring on his finger, and shoes on his feet.}

\gresetinitiallines{1}
\gregorioscore{comm-lent-3}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Angelis suis Deus mandávit de te.\\
	\Rbar.~Ut custódiat te in ómnibus viis tuis.
	
	Orémus.
	Qu\'{\ae}sumus, omnípotens Deus, vota humílium réspice:
	atque ad defensiónem nostram,
	déxteram tuæ majestátis exténde.
	Per Dóminum.
}{
	He hath given his angels charge over thee.
	\Rbar.~To keep thee in all thy ways.
	
	Let us pray.
	We beseech You, almighty God, hear the prayers of the humble and stretch forth the right hand of Your majesty to protect us.
	Through our Lord.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-2v-solem}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

%\begin{rubricbox}
%
%The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).
%
%\end{rubricbox}

\section*{Marian Anthem}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/ave-regina-caelorum-simple}

{\itshape
	Hail, O Queen of Heaven enthroned.
	Hail, by angels mistress owned.
	Root of Jesse, Gate of Morn
	whence the world's true light was born.
	
	Glorious Virgin, Joy to thee,
	loveliest whom in heaven they see;
	fairest thou, where all are fair,
	plead with Christ our souls to spare.
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Dignáre me laudáre te, Virgo sacráta.\\
	\Rbar. Da mihi virtútem contra hostes tuos.
	
	Orémus.
	Concéde, miséricors Deus, fragilitáti nostræ præsídium:
	ut qui sanctæ Dei Genitrícis memóriam ágimus;
	intercessiónis ejus auxílio, a nostris iniquitátibus resurgámus.
	Per eúmdem Christum Dóminum nostrum.
	\Rbar.~Amen.
}{
	Vouchsafe that I may praise thee, O sacred Virgin.
	\Rbar.~Give me stength against thine enemies.
	
	Let us pray.
	Grant, O merciful God, to our weak natures Thy protection,
	that we who commemorate the holy Mother of God may,
	by the help of her intercession, arise from our iniquities.
	Through the same Christ our Lord.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}