\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, August 20, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Eleventh Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romas 11:33)}

\begin{latinenglishsection}

 \latinenglish{
 	O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\ 
 	quam incomprehensibília sunt judí\textit{cia} \textbf{e}jus, et investigábiles viæ ejus!
	 
 	\Rbar.~Deo grátias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape
	1. As fades the glowing orb of day,
	To thee, great source of light, we pray;
	Blest Three in One, to every heart
	Thy beams of life and love impart.
	
	At early dawn, at close of day,
	To thee our vows we humbly pay;
	May we, mid joys that never end,
	With thy bright saints in homage bend.
	
	To God the Father, and the Son,
	And Holy Spirit, Three in One,
	Be endless glory, as before
	The world began, so evermore.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) des(h)cén(h)d(h)at(h) su(h)p(h)er(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Wisdom crieth in the streets: If any man love wisdom, let him come unto me, and he shall find her and when once he hath found her, blessed shall he be if he hold her.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens sempitérne Deus, qui, abundántia pietátis tuæ, et mérita súpplicum excédis et vota:~\GreDagger effúnde super nos misericór\textit{diam} \textbf{tu}am; ut dimíttas quæ consciéntia métuit, et adícias quod orátio non præsúmit.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and everlasting God, Who art always more ready to hear than we to pray, and art wont to give more than either we desire or deserve, pour down upon us the abundance of thy mercy, forgiving us those things whereof our conscience is afraid, and giving us those good things which we are not worthy to ask.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of the Octave of the Assumption of the Blessed Virgin Mary}

\textit{\textnormal{Ant.}  Today the Blessed Virgin Mary * ascended to heaven, rejoice, she reigns with Christ forever.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--hodie_maria_virgo_caelos--solesmes_1934}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Exaltáta est sancta Dei Génitrix..\\
	\Rbar.~Super choros Angelórum ad cæléstia regna.
	
	Orémus.
Omnípotens sempitérne Deus, qui Immaculátam Vírginem Maríam, Fílii tui Genetrícem, córpore et ánima ad cæléstem glóriam assumpsísti: concéde, quǽsumus; ut ad supérna semper inténti, ipsíus glóriæ mereámur esse consórtes.
}{
	\Vbar.~Let the evening prayer ascend unto thee, O Lord.
	\Rbar.~And let there descend upon us thy mercy.
	
	Let us pray.
	Almighty everlasting God, who hast taken body and soul into heaven the Immaculate Virgin Mary, Mother of thy Son: grant, we beseech thee, that by steadfastly keeping heaven as our goal we may be counted worthy to join her in glory.
}
\end{latinenglishsection}

\section*{Commemoration of St. Bernard of Clairvaux, Doctor of the Church (August 20)}

\textit{\textnormal{Ant.} O right excellent Teacher, light of the Holy Church, St. N, Lover of God's law, pray for us to the Son of God.} 

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--o_doctor_optime--solesmes_1961}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ 
	\Rbar.~Et osténdit illi regnum Dei. 
	
	Orémus.
	Deus, qui pópulo tuo ætérnæ salútis beátum Bernárdum minístrum tribuísti: præsta, quǽsumus; ut quem Doctórem vitæ habúimus in terris, intercessórem habére mereámur in cælis.
}{
	\Vbar.~The Lord guided the just in right paths. 
	\Rbar.~And showed him the kingdom of God. 
	
	Let us pray.
	O God, who didst give blessed Bernard to be thy people’s minister in eternal salvation; grant, we pray, that we who have him for teacher of life here on earth, may also deserve now that he is in heaven, to have him for an advocate.
}

\end{latinenglishsection}

\section*{Commemoration of St. Jeanne-Françoise Frémiot (Augsut 21)}

\textit{\textnormal{Ant.} The kingdom of heaven is like a merchant seeking good pearls, who, when he had found one of great price,
gave all that he had and bought it.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/holy-woman-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Spécie tua et pulchritúdine tua.\\
	\Rbar.~Inténde, próspere procéde, et regna. 
	
	Orémus.
	Omnípotens et miséricors Deus, qui beátam Joánnam Francíscam, tuo amóre succénsam, admirábili spíritus fortitúdine per omnes vitæ sémitas in via perfectiónis donásti, quique per illam illustráre Ecclésiam tuam nova prole voluísti: ejus méritis et précibus concéde; ut, qui infirmitátis nostræ cónscii de tua virtúte confídimus, cæléstis grátiæ auxílio cuncta nobis adversántia vincámus.
	Per eúndem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	In thy glory and thy splendor.
	\Rbar.~Go forth, advance with victory and reign.
	
	Let us pray.
	O Almighty and merciful God, Who didst enkindle the love of thyself in the blessed Jeanne Frances, Who didst give unto her the grace to make every path of life temporal the straight and narrow way which leadeth unto life eternal, and Who wast pleased to use her as a mean whereby to adorn thy Church with a new sisterhood, grant unto us for that thine handmaid's sake, and at her prayers, that we who know that we have no strength as of ourselves to help ourselves, and therefore do put all our trust in thine Almighty power, may by the assistance of thy heavenly grace, always prevail in all things against whatsoever shall arise to fight against us.
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}