\documentclass{../vespers-sheet}

%% reduce the spacing before the title so the hymn can fit all on one page
\titlespacing*{\chapter}{0pt}{35pt}{40pt}

\begin{document}

\chapter*{Vespers Insert for Saturday, 11 Dec. 2021}

 \textit{Vespers today is based on the following Sunday, which is the Third of Advent, also known as Gaudete Sunday. All is said as in the main booklet until the Little Chapter.}
 
 \section*{Little Chapter (Philippians 4:4)}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Gaudéte in Dómino semper: íterum dico, gaudéte.~\GreDagger\
	Modéstia vestra nota sit ómnibus homínibus:~*
	Dóminus enim prope est.
	\Rbar.~Deo grátias.
}{
	 Brethren: Rejoice in the Lord always; again, I say, rejoice. Let your modesty be known to all men, for the Lord is nigh. 
	 \Rbar.~Thanks be to God.
}
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Neither before Me nor after Me shall there be any other God; for to Me shall every knee bow
and every tongue confess.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Aurem tuam, qu\'{\ae}sumus Dómine, précibus nostris accómmoda:
	et mentis nostræ ténebras grátia tuæ visitatiónis illústra.
	Qui vivis.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Incline Thine ear to our prayers, we beseech Thee, O Lord;
	and enlighten the darkness of our minds by the grace of Thy visitation:
	Who livest.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of Our Lady of Guadalupe (Dec. 12)}

\textit{\textnormal{Ant.} I have chosen and sanctified this place, that my name may reside there, and my eyes and my heart remain there for all days.}

\gresetinitiallines{1}
\gregorioscore{comm-olog}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Non fecit táliter omni natióni.\\
	\Rbar.~Et judícia sua non manifestávit eis.\\
	
	Orémus.
	Deus, qui sub beatíssimæ Vírginis Maríæ singulári patrocínio constitútos, perpétuis benefíciis nos cumulári voluísti:
	præsta supplícibus tuis; ut cujus hódie commemoratióne lætámur in terris,
	ejus conspéctu perfruámur in cælis.
}{
	He hath not done in like manner to every nation.
	\Rbar.~And his judgments he hath not made manifest to them.
	
	Let us pray.
	O God, Who hast placed us under the special patronage of the Blessed Virgin Mary,
	and through her hast favored us with endless blessings: may we, who joyfully honor her this day on earth,
	enjoy her company forever in heaven. 
}

\end{latinenglishsection}

\section*{Commemoration of Pope St. Damasus (Dec. 11)}

\textit{\textnormal{Ant.} The Lord loved him, and adorned him; He clothed him with a robe of glory, and crowned him at the gates of Paradise.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-pont-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ %PT: allelúia
	\Rbar.~Et osténdit illi regnum Dei.\\  %PT: allelúia
	
	Orémus.
	Gregem tuum, Pastor ætérne, placátus inténde:
	et per beátum Dámasum Summum Pontíficem, perpétua protectióne custódi;
	quem totíus Ecclésiæ præstitísti esse pastórem.
	Per Dóminum.
}{
	The Lord led the just in right paths. %PT: alleluia
	\Rbar.~And showed him the kingdom of God. %PT: alleluia
	
	Let us pray.
	Eternal shepherd, look kindly upon Thy flock, and preserve it with a continual protection, by the blessed Damasus, sovereign pontiff,
	whom Thou hast made the shepherd of the whole Church.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{The Marian anthem \textnormal{Alma Redemptoris Mater} follows (no. 181 in the \textnormal{Traditional Roman Hymnal}).}

\end{rubricbox}

\end{document}