\documentclass{../vespers-sheet}
\usepackage{multicol}

\begin{document}

\chapter*{Ember Friday of Pentecost}

\section*{Beginning of the Office}

\begin{rubricbox}

{\color{red}When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and make the sign of the cross with the leader as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
{\color{red}\Vbar.}~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\section*{Psalm 109}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-intonation} %TODO: look for alternate 3rd mode melody

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-3}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps109-antiphon}

\textit{\textnormal{Ant. 1.} When the days of Pentecost were drawing to a close, they were all saying together, alleluia.}

\section*{Psalm 110}

\gresetinitiallines{1}
\gregorioscore{ps110-intonation}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-8}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\vfill\pagebreak

\gresetinitiallines{0}
\gregorioscore{ps110-antiphon}

\textit{\textnormal{Ant. 2.} The Spirit of the Lord has filled the whole world, alleluia.}


\section*{Psalm 111}

\gresetinitiallines{1}
\gregorioscore{ps111-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-8}
	
}{
	\input{../psalms/english/ps111}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps111-antiphon}

\textit{\textnormal{Ant. 3.} They were all filled with the Holy Spirit, and began to speak, alleluia.}

\vfill\pagebreak

\section*{Psalm 112}

\gresetinitiallines{1}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-1}
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{ps112-antiphon}

\textit{\textnormal{Ant. 4.} Ye fountains, and all things that move in the waters, sing a hymn to God, alleluia.}

\section*{Psalm 113}

\textit{\textnormal{Ant. 5.} The Apostles were speaking with diverse tongues the wonderful works of God, alleluia, alleluia, alleluia.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps113-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps113-7}
	
}{
	\input{../psalms/english/ps113}
}

\end{latinenglishsection}

\vfill\pagebreak

\gresetinitiallines{1}
\gregorioscore{ps113-antiphon}

\section*{Little Chapter (Acts 2:1-2)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	 Cum compleréntur dies Pentecóstes, erant omnes discípuli pariter in eódem loco:~\GreDagger\
	 et factus est repénte de c{\oe}lo sonus, tamquam adveniéntis spíritus veheméntis:~*
	 et replévit totam domum, ubi erant sedentes.
	{\color{red}\Rbar.}~Deo grátias.
}{
 	When the days of the Pentecost were accomplished, they were all together in one place:
 	and suddenly there came a sound from heaven, as of a mighty wind coming,
 	and it filled the whole house where they were sitting. 
	 {\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/veni-creator-spiritus_solemn_vespers}

{\itshape

	1. Come, Holy Spirit, Creator blest,
	and in our souls take up Thy rest;
	come with Thy grace and heavenly aid
	to fill the hearts which Thou hast made.
	
	2. O comforter, to Thee we cry,
	O heavenly gift of God Most High,
	O fount of life and fire of love,
	and sweet anointing from above.
	
	3. Thou in Thy sevenfold gifts are known;
	Thou, finger of God's hand we own;
	Thou, promise of the Father, Thou
	Who dost the tongue with power imbue. 
	
	4. Kindle our sense from above,
	and make our hearts o'erflow with love;
	with patience firm and virtue high
	the weakness of our flesh supply.
	
	5. Far from us drive the foe we dread,
	and grant us Thy peace instead;
	so shall we not, with Thee for guide,
	turn from the path of life aside.
	
	6. Oh, may Thy grace on us bestow
	the Father and the Son to know;
	and Thee, through endless times confessed,
	of both the eternal Spirit blest.
	
	7. Now to the Father and the Son,
	Who rose from death, be glory given,
	with Thou, O Holy Comforter,
	henceforth by all in earth and heaven.
	Amen. 
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>V/</sp>.</c> Re(h)plé(h)ti(h) sunt(h) om(h)nes(h) Spí(h)ri(h)tu(h) San(h)cto,(h) (,)
al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>R/</sp>.</c> Et(h) c<sp>oe</sp>(h)pé(h)runt(h) lo(h)qui,(h) (,) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.)(::)
}

\textit{{\color{red}\Vbar.}~They were all filled with the Holy Spirit, alleluia.
{\color{red}\Rbar.}~And began to speak, alleluia.}

\vfill\pagebreak

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-8}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\section*{Collect}

\begin{rubricbox}

{\color{red} All \textbf{stand} while the Officiant chants the collect prayer.}

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui hodiérna die corda fidélium Sancti Spíritus illustratióne docuísti:
	da nobis in eódem Spíritu recta sápere;
	et de ejus semper consolatióne gaudére.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	{\color{red}\Rbar.}~Amen. 
}{
	Lord, hear my prayer.
	{\color{red}\Rbar.}~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who on this day hast taught the hearts of the faithful by the light of the Holy Spirit,
	grant us by that same Spirit, to know what is right, and ever to rejoice in His consolation.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	{\color{red}\Rbar.}~Amen. 
}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
}{
	{\color{red}\Vbar.}~Lord, hear my prayer. {\color{red}\Rbar.}~And let my cry come unto Thee.
	{\color{red}\Vbar.}~Let us bless the Lord. {\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tp}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	{\color{red}\Rbar.} Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. {\color{red}\Rbar.}~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	{\color{red}\Vbar.} Dóminus det nobis suam pacem. \\
	{\color{red}\Rbar.} Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. {\color{red}\Rbar.}~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The Marian anthem \textit{Regina Caeli} follows, as is fitting for the season.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/regina-caeli-simple}

\begin{latinenglishsection}
\latinenglish{
	{\color{red}\Vbar.}~Gaude et lætáre, Virgo María, allelúia.\\
	{\color{red}\Rbar.}~Quia surréxit Dóminus vere, allelúia.
	
	Orémus.
	Deus, qui per resurrectiónem Fílii tui, Dómini nostri Jesu Christi, mundum lætificáre dignátus es: præsta, qu\'{\ae}sumus; ut per ejus Genetrícem Vírginem Maríam, perpétuæ capiámus gáudia vitæ. Per eúndem Christum Dóminum nostrum.
	{\color{red}\Rbar.}~Amen. 
}{
	{\color{red}\Vbar.}~Rejoice and be glad, O Virgin Mary, alleluia. 
	{\color{red}\Rbar.}~For the Lord has truly risen, alleluia.

	Let us pray.
	O God, who gave joy to the world through the resurrection of Thy Son, our Lord Jesus Christ, grant we beseech Thee, that through the intercession of the Virgin Mary, His Mother, we may obtain the joys of everlasting life. Through the same Christ our Lord. 
	{\color{red}\Rbar.}~Amen. 
}
\end{latinenglishsection}

\textit{\color{red}The Officiant says the following slowly, in a low recto tono:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Divínum auxílium máneat semper nobíscum.\\
	{\color{red}\Rbar.} Amen.
}{
	May the divine assistance remain always with us.
	{\color{red}\Rbar.}~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

{\color{red}After the Office, all \textbf{kneel} and pray in silence for a time.}

\end{rubricbox}

\end{document}