\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of the Circumcision of Our Lord}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently.
Then, when the Officiant stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\vfill\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} O wondrous interchange! * the Creator of mankind, taking upon him a living body, vouchsafed to be born of a pure Virgin: and by his Humanity, which was begotten in no earthly wise, hath made us partakers of his Divinity..
 \textnormal{Ps.} The Lord said to my Lord: * Sit thou at my right hand.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-6} %%

}{
	\input{../psalms/english/ps109} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{O admirábile commércium\dots}} %%

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 112}

\textit{\textnormal{Ant. 2.} When thou wast born * all ineffably of a Virgin, then were the Scriptures fulfilled; thou camest down like the dew upon the fleece of wool, to bring salvation unto all mankind; we praise thee, O our God.
 \textnormal{Ps.} Praise the Lord, ye children: * praise ye the name of the Lord.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-3a2} %%

}{
	\input{../psalms/english/ps112} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Quando natus es\dots}} %%

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 121}

\textit{\textnormal{Ant. 3.}  In the bush which Moses saw * unconsumed, we recognize the preservation of thy glorious virginity: holy Mother of God, intercede for us.
 \textnormal{Ps.} I rejoiced at the things that were said to me: * We shall go into the house of the Lord.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps121-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps121-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps121-4-E-A-d} %%

}{
	\input{../psalms/english/ps121} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Rubum, quem víderat Móyses\dots}} %%

\end{rubricbox}

\vfill\pagebreak

\section*{Psalm 126}

\textit{\textnormal{Ant. 4.} The Root of Jesse hath budded, * the Star hath come out of Jacob, the Virgin hath borne the Saviour: we praise thee, O our God.
 \textnormal{Ps.} Unless the Lord build the house, * they labour in vain that build it.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps126-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps126-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps126-1} %%

}{
	\input{../psalms/english/ps126} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Germinávit radix Jesse\dots}} %%

\end{rubricbox}

\section*{Psalm 147}

\textit{\textnormal{Ant. 5.} Lo, Mary hath brought forth * the Saviour of whom when John saw him, he did proclaim: Behold the Lamb of God that taketh away the sins of the world, alleluia.
 \textnormal{Ps.} Praise the Lord, O Jerusalem: * praise thy God, O Sion.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps147-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps147-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps147-2} %%

}{
	\input{../psalms/english/ps147} %%
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The antiphon is repeated: \textit{Ecce María génuit \dots}} %%

\end{rubricbox}

\section*{Little Chapter (Titus 2:11-12)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Appáruit grátia Dei Salvatóris nostri ómnibus homínibus,\GreDagger\ erúdiens nos, ut abnegántes imp\textit{ie}\textbf{tá}tem, et\\ sæculária desidéria,~* sóbrie, et juste, et pie vivámus in hoc s\'{\ae}culo.
	\Rbar.~Deo grátias.
}{
	The grace of God that bringeth salvation hath appeared to all men, teaching us that, denying ungodliness and worldly lusts, we should live soberly, righteously, and godly in this present world.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jesu-redemptor-omnium}

{\itshape

	1.  Jesus, the Ransomer of man,
	Who, ere created light began,
	Didst from the sovereign Father spring,
	His power and glory equalling.
	
	2. The Father’s Light and Splendor Thou,
	Their endless hope to Thee that bow;
	Accept the prayers and praise to-day
	That through the world Thy servants pay.
	
	3. Salvation’s Author, call to mind
	How, taking form of humankind,
	Born of a Virgin undefiled,
	Thou in man’s flesh becam’st a Child.
	
	4. Thus testifies the present day,
	Through every year in long array,
	That Thou, salvation’s source alone,
	Proceedest from the Father’s throne.
	
	5. The heavens above, the rolling main
	And all that earth’s wide realms contain,
	With joyous voice now loudly sing
	The glory of their new-born King.
	
	6. And we who, by Thy precious Blood
	From sin redeemed, are marked for God,
	On this the day that saw Thy birth,
	Sing the new song of ransomed earth.
	
	7. O Lord, the Virgin-born, to Thee
	Eternal praise and glory be,
	Whom with the Father we adore
	And Holy Ghost forevermore.
	Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ver(h)bum(h) ca(h)ro(h) fac(h)tum(h) est(h), al(h)le(h)lú(h)ja.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) ha(h)bi(h)tá(h)vit(h) in(h) no(h)bis(h), al(h)le(h)lú(h)ja.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~The Word was made flesh, alleluia.
\Rbar.~And dwelt among us, alleluia.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Behold a faithful and prudent servant, whom the Lord has set over His household.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.}

\begin{rubricbox}

{\color{red}The Cantor leads by intoning the antiphon and the first verse.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

{\color{red}All \textbf{stand} and make the sign of the cross with the Cantor.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
3. Quia respéxit humilitátem ancíllæ \textbf{su}æ:~*
	ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.

4. Quia fecit mihi magna qui \textbf{pot}ens est:~* 
	et sanctum \textit{no}\textit{men} \textbf{e}jus.

5. Et misericórdia ejus a progénie in pro\textbf{gé}nies~*
	timén\textit{ti}\textit{bus} \textbf{e}um.

6. Fecit poténtiam in bráchio \textbf{su}o:~*
	dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.

7. Depósuit poténtes de \textbf{se}de,~*
	et exal\textit{tá}\textit{vit} \textbf{hú}miles.

8. Esuriéntes implévit \textbf{bo}nis:~*
	et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.

9. Suscépit Israël púerum \textbf{su}um,~*
	recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.

10. Sicut locútus est ad patres \textbf{nos}tros,~*
	Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\textit{\color{red}(bow)}Glória Patri, et \textbf{Fí}lio,~* 
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{\color{red}(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.


\begin{rubricbox}

{\color{red}All \textbf{sit} and repeat the antiphon: \textit{Propter nímiam\dots}, %%
then \textbf{stand} for the prayer.}

\end{rubricbox}

\section*{Collect}

\textit{\color{red}The Officiant leads the collect:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui salútis ætérnæ, beátæ Maríæ virginitáte fecúnda, humáno géneri pr\'{\ae}mia præstitísti: tríbue, qu\'{\ae}sumus; ut ipsam pro nobis intercédere sentiámus, per quam merúimus auctórem vitæ\\ suscípere, Dóminum nostrum Jesum Christum Fílium tuum:
Qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who, by the fruitful virginity of the Blessed Mary, hast given unto mankind the rewards of everlasting life; grant, we beseech thee, that we may continually feel the might of her intercession, through whom we have worthily received the Author of our life, our Lord Jesus Christ, thy Son:
Who with thee liveth and reigneth, in the unity of the Holy Spirit, God, world without end.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of Pope St. Slyvester I, Confessor (December 31)}

\textit{\textnormal{Ant.} While he was supreme Pontiff, he feared not earthly powers, but gloriously went his way to the heavenly kingdom.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an-dum_esset_summus_pontifex}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ 
	\Rbar.~Stolam glóriæ índuit eum. 
	
	Orémus.
	Gregem tuum, Pastor ætérnæ placátus inténde: et per beátum Silvéstrum, Summum\\ Pontíficem tuum, perpétua\\ protectióne custódi; quem totíus Ecclésiæ præstitísti esse pastórem. 
	Per Dóminum nostrum Jesum\\ Christum, Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar. Amen.
}{
	\Vbar.~The Lord loved him and beautified him. 
	\Rbar.~He clothed him with a robe of glory. 
	
	Let us pray.
	Eternal Shepherd, direct gently Thy flock, and guard it with an everlasting protection, by the intercession of blessed Sylvester, Thy Sovereign Pontiff, whom Thou hast constituted Shepherd of the whole Church. 
	Through our Lord Jesus Christ, Thy Son, Who lives and reigns with Thee in the unity of the Holy Spirit, God, forever and ever.
	\Rbar. Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}
\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solem}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\section*{Marian Anthem}

\textit{\color{red}The Cantor leads the Marian anthem and responses afterwards; the Officiant leads the ending collect:}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/alma-redemptoris-monasticum}

{\itshape
	Holy mother of our Redeemer,
	thou gate leading to heaven and star of the sea; 
	help the falling people who seek to rise, 
	thou who, all nature wondering, didst give birth to thy holy Creator. 
	Virgin always, hearing the greeting from Gabriel’s lips, take pity on sinners.
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Post partum, Virgo, inviolata permansisti.\\
	\Rbar. Dei Genitrix, intercede pro nobis.
	
	Orémus.
	Deus, qui salutis aeternae, beatae Mariae virginitate fecunda, humano generi praemia praestitisti: \GreDagger tribue, quaesumus, ut ipsam pro nobis intercedere \textit{senti}\textbf{a}mus, per quam meruimus auctorem vitae\\ suscipere, ~* Dominum nostrum Iesum Christum, Filium tuum.\\
	\Rbar.~Amen.
}{
	Vouchsafe that I may praise thee, O sacred Virgin.
	\Rbar.~Give me stength against thine enemies.
	
	Let us pray.
	O God, Who by the fruitful virginity of blessed Mary hast offered unto the human race the rewards of eternal salvation, grant, we beseech thee, that we may know the effects of her intercession, through whom we have deserved to receive the author of life, our Lord Jesus Christ, Thy Son. 
	\Rbar.~Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant says the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}