\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers of the Circumcision}

\section*{Before the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently, making the sign of the cross first over the lips, then over the heart at the marked spots:

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	Aperi, + Dómine, os meum ad benedicéndum nomen sanctum tuum:
	+ munda quoque cor meum ab ómnibus vanis, pervérsis et aliénis cogitatiónibus;
	intelléctum illúmina, afféctum inflámma, ut digne, atténte ac devóte hoc Offícium recitáre váleam,
	et exaudíri mérear ante conspéctum divínæ Majestátis tuæ.
	Per Christum Dóminum nostrum. 
	Amen.
}{
	Open, + O Lord, my mouth to bless Thy holy Name; + cleanse also my heart from all vain, evil, and wandering thoughts; enlighten my understanding and kindle my affections; that I may worthily, attentively, and devoutly say this Office, and so merit to be heard before the presence of Thy divine Majesty.  Through Christ our Lord.  Amen.
}

\latinenglish{
	Domine, in unióne illíus divínæ intentiónis, qua ipse in terris laudes Deo persolvísti, hanc tibi Horam persólvo.
}{
	O Lord, in union with that divine intention wherewith Thou, whilst here on earth, didst render praises unto God, I desire to offer this Hour of prayer unto Thee.
}

\end{latinenglishsection}

%\section*{Beginning of the Office}

\begin{rubricbox}

When the leader stands, all \textbf{stand} and pray silently:

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	Pater Noster\dots
}{
	Our Father\dots
}

\latinenglish{
	Ave María\dots
 }{
 	Hail Mary\dots
 }


 \end{latinenglishsection}

\vfill

\pagebreak

\begin{rubricbox}

All make the sign of the cross with the leader as he intones:

\end{rubricbox}
 
 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} O wondrous exchange! The Creator of mankind, having assumed a living body, deigned to be born of a Virgin, and having become man without man's aid, enriched us with His divinity.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

%\gresetinitiallines{1}
%\gregorioscore{ps109-antiphon-alone}

%\vfill
%\pagebreak
%
% \begin{rubricbox}
%
%After the psalm is intoned, all \textbf{sit} at the asterisk.
%
%\end{rubricbox}
%
%\gresetinitiallines{0}
%\gregorioscore{ps109-psalm-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-6}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{O admirabile commercium\dots}, then all \textbf{stand}.

\end{rubricbox}


\section*{Psalm 112}

\textit{\textnormal{Ant. 2.} By Thy ineffable birht of a Virgin, the Scriptures were fulfilled; like rain upon the fleece didst Thou descend to save mankind: we praise Thee, our God.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}


\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps112-3}

}{
 	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Quando natus es\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 121}

\textit{\textnormal{Ant. 3.} In the bush which Moses saw burning yet unconsumed, we recognize the preservation of thy glorious virginity; O Mother of God, intercede for us.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps121-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps121-4-E-A-d}
	
}{
	\input{../psalms/english/ps121}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{Rubum quem viderat\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 126}

\textit{\textnormal{Ant. 4.} The Root of Jesse hath blossomed; the Star hath risen out of Jacob; the Virgin hath borne the Savior; we praise Thee, our God.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps126-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps126-1}
	
}{
	\input{../psalms/english/ps126}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Germinavit\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 147}

\textit{\textnormal{Ant. 5.} Behold, Mary hath given birth to the Savior, Whom John seeing, exclaimed: Behold the Lamb of God; behold Him who taketh away the sins of the world, alleluia.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps147-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps147-2}
	
}{
	\input{../psalms/english/ps147}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Ecce Maria\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Titus 2:11)}

\begin{latinenglishsection}

\latinenglish{
	
	Appáruit grátia Dei Salvatóris nostri ómnibus homínibus, \GreDagger\
	erúdiens nos, ut abnegántes impietátem et sæculária desidéria,~*
	sóbrie, et juste, et pie vivámus in hoc s\'{\ae}culo.
	\Rbar.~Deo grátias.
}{
	 The grace of God our Saviour hath appeared to all men; instructing us, that,
	 denying ungodliness and worldly desires, we should live soberly, and justly, and godly in this world.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/christe-redemptor-omnium-ex-patre}

{\itshape

Jesu, the Father's only Son,
whose death for all redemption won,
before the worlds, of God most high,
begotten all ineffably. 

 The Father's Light and Splendor Thou
their endless Hope to Thee that bow:
accept the prayers and praise today
that through the world Thy servants pay. 

 Salvation's author, call to mind
how, taking the form of humankind,
born of a Virgin undefiled,
Thou in man's flesh becamest a Child. 

Thus testifies the present day
Through every year in long array,
that Thou, salvation's source alone
proceedest from the Father's Throne. 

Whence sky, and stars, and sea's abyss,
and earth, and all that therein is,
shall still, with laud and carol meet,
the Author of thine Advent greet. 

And we who, by Thy precious Blood
from sin redeemed, are marked for God,
on this, the day that saw Thy Birth,
sing the new song of ransomed earth. 

All honor, laud, and glory be,
O Iesu, Virgin-born, to Thee;
whom with the Father we adore,
and Holy Ghost forevermore. Amen. 
}

\begin{rubricbox}

The Office continues on the accompanying insert with the \Rbar. and \Vbar.

\end{rubricbox}

\end{document}