\documentclass{../vespers-booklet}
\usepackage{multicol}

\titlespacing*{\chapter}{0pt}{0pt}{40pt}

\begin{document}

\chapter*{Second Vespers of the Circumcision}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. No(h)tum(h) fe(h)cit(h) Dó(h)mi(h)nus,(h) al(h)le(h)lu(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Sa(h)lu(h)tá(h)re(h) su(h)um,(h) al(h)le(h)lu(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} Great is the mystery of our inheritance; the womb of her who knew not man has become the temple of God; by taking flesh of her he was not defiled; all nations shall come and say: Glory be to Thee, O Lord.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-2v-antiphon-alone.gabc}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-2v-intonation.gabc}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-2}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon (on the first page): \textit{Magnum\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\vfill

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus. Omnípotens sempitérne Deus, qui in dilécto Fílio tuo, universórum Rege, ómnia instauráre voluísti:
	concéde propítius; ut cunctæ famíliæ Géntium, peccáti vúlnere disgregátæ,
	ejus suavíssimo subdántur império. Qui tecum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and everlasting God, who in Thy beloved Son, King of the whole world, hast willed to restore all things anew;
	grant in Thy mercy that all the families of nations, rent asunder by the wound of sin,
	may be subjected to His most gentle rule. Who with Thee liveth.
	\Rbar.~Amen.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{0}
\gregorioscore{../common/benedicamus-bvm}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\rubric{If no other service follows:}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{The Marian anthem \textnormal{Alma Redemptoris Mater} follows (no. 181 in the \textnormal{Traditional Roman Hymnal}).}

\end{rubricbox}

\end{document}