\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of the Apostles Philip and James}

\section*{Beginning of the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently.
Then, when the leader stands, all \textbf{stand} and say silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the leader as he intones:

\end{rubricbox}

 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\vfill
\pagebreak

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} O wondrous exchange! The Creator of mankind, having assumed a living body, deigned to be born of a Virgin, and having become man without man's aid, enriched us with His divinity.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.
After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-7}

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Domine\dots}, then all \textbf{stand}.

\end{rubricbox}


\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} By Thy ineffable birth of a Virgin, the Scriptures were fulfilled; like rain upon the fleece didst Thou descend to save mankind: we praise Thee, our God.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
	\input{../psalms/ps110-7}

}{
 	\input{../psalms/english/ps110}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Philippe\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} In the bush which Moses saw burning yet unconsumed, we recognize the preservation of thy glorious virginity; O Mother of God, intercede for us.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-3}
	
}{
	\input{../psalms/english/ps111}
}

\end{latinenglishsection}
\begin{rubricbox}

The antiphon is repeated: \textit{Tanto tempore\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} The Root of Jesse hath blossomed; the Star hath risen out of Jacob; the Virgin hath borne the Savior; we praise Thee, our God.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-7}
	
}{
	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Si cognovissetis me\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} Behold, Mary hath given birth to the Savior, Whom John seeing, exclaimed: Behold the Lamb of God; behold Him who taketh away the sins of the world, alleluia.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps116-3}
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

\begin{rubricbox}

The antiphon is repeated: \textit{Si diligitis me\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Wisdom 5:1)}

\begin{latinenglishsection}

\latinenglish{
	Stabunt justi in magna constántia~\GreDagger\
	advérsus eos qui se angustiavérunt,~*
	et qui abstulérunt labóres eórum.
	\Rbar.~Deo grátias.
}{
	Tthe just shall stand with great constancy against those that have afflicted them, and taken away their labours.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/tristes-erant-apostoli}

{\itshape

%% TODO
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. San(h)cti(h) et(h) jus(h)ti(h) in(h) Dó(h)mi(h)no(h) gau(h)dé(h)te,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Vos(h) e(h)lé(h)git(h) De(h)us(h) in(h) he(h)re(h)di(h)tá(h)tem(h) si(h)bi,(h) al(h)le(h)lú(h)ia.(g'_/hvGF'E/fgf.) (::)
}

%% TODO
\textit{\Vbar.~Blessed are they who dwell in Thy house, O Lord.
\Rbar.~They will praise Thee for ever and ever.}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} The Word was made flesh, and dwelt among us, full of grace and truth; of Whose fullness we have all received, and grace for grace, alleluia.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-6}

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon: \textit{Non turbetur\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui nos ánnua Apostolórum tuórum Philíppi et Jacóbi solemnitáte lætíficas:
	præsta, qu\'{\ae}sumus; ut quorum gaudémus méritis,
	instruámur exémplis.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	%% TODO
	\Rbar.~Amen.
}
\end{latinenglishsection}

\section*{Commemoration of St. Catherine of Sienna (April 30)}

\textit{\textnormal{Ant.} Come, Spouse of Christ, receive the crown which the Lord hath prepared for thee for all eternity, alleluia.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/virgin-2v-pt}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Diffúsa est grátia in lábiis tuis, allelúia.\\
	\Rbar.~Proptérea benedíxit te Deus in ætérnum, allelúia.
	
	Orémus.
	Da qu\'{\ae}sumus omnípotens Deus:
	ut qui beátæ Catharínæ Vírginis tuæ natalítia cólimus;
	et ánnua solemnitáte lætémur, et tantæ virtútis proficiámus exemplo.
}{
	Grace is poured out upon thy lips, alleluia.
	\Rbar.~Therefore God hath blessed thee forever, alleluia.
	
	Let us pray.
	%% translation of the prayer
}

\end{latinenglishsection}

\section*{Commemoration of the Sunday}

\textit{\textnormal{Ant.} I am the Shepherd of the sheep; I am the Way, the Truth, and the Life. 
I am the Good Shepherd, and I know Mine and Mine know Me, alleluia, alleluia.}

\gresetinitiallines{1}
\gregorioscore{comm-2-post-pascha}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Mane nobíscum Dómine, allelúia.\\
	\Rbar.~Quóniam advesperáscit, allelúia.
	
	Orémus.
	Deus, qui in Fílii tui humilitáte jacéntem mundum erexísti:
	fidélibus tuis perpétuam concéde lætítiam;
	ut, quos perpétuæ mortis eripuísti cásibus, gaudiis fácias pérfrui sempitérnis.
	Per eúmdem Dóminum.
}{
	Stay with us, Lord, alleluia.
	\Rbar.~For it is getting towards evening, alleluia.
	
	Let us pray.
	O God, who, by the humility of Thy Son, didst lift up a fallen world, grant unending happiness to Thy faithful:
	that those whom Thou hast snatched from the perils of endless death, Thou mayest cause to rejoice in everlasting days.
	Through the same.
	\Rbar.~Amen.
}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-double}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Regina Caeli} follows (no. 221 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}