\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, November 5, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Twenty-second Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Rom 11:33)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	 O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,\GreDagger\
	 quam incomprehensibília sunt ju\textit{dícia} \textbf{e}jus, et investigábiles viæ ejus!
 	\Rbar.~Deo gratias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape

1. As fades the glowing orb of day,
To thee, great source of light, we pray;
Blest Three in One, to every heart
Thy beams of life and love impart.

2. At early dawn, at close of day,
To thee our vows we humbly pay;
May we, mid joys that never end,
With thy bright saints in homage bend.

3. To God the Father, and the Son,
And Holy Spirit, Three in One,
Be endless glory, as before
The world began, so evermore.
Amen.

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h)scén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Hedge us about with thy wall that cannot be broken down, O Lord, and shield us continually with the arms of thy might.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, refúgium nostrum et virtus: adésto piis Ecclésiæ tuæ précibus, auctor ipse pietátis, et præsta; ut, quod fidéliter pétimus, efficáciter consequámur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, our Refuge and Strength, Who art the author of all godliness, be ready, we beseech thee, to hear the devout prayers of thy Church, and grant that those things which we ask faithfully, we may obtain effectually.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of the Fifth Day of the Octave of All Saints}

\textit{\textnormal{Ant.} I saw the Lord sitting upon a throne high and lifted up, and the whole earth was full of His glory, and His train filled the Temple.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--o_quam_gloriosum_est_regnum--solesmes_1934}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Exsultábunt sancti in glória.\\
	\Rbar.~Lætabúntur in cubílibus suis.
	
	Orémus.
	Omnípotens sempitérne Deus, qui nos ómnium Sanctórum tuórum mérita sub una tribuísti celebritáte venerári: quǽsumus; ut desiderátam nobis tuæ propitiatiónis abundántiam, multiplicátis intercessóribus, largiáris.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
}{
	Let the Saints be joyful in glory.
	\Rbar.~Let them sing aloud upon their beds.
	
	Let us pray.
	Almighty, eternal God, Who granted us to honor the merits of all Thy Saints in a single solemn festival, bestow on us, we beseech Thee, through their manifold intercession, that abundance of Thy mercy for which we yearn.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}