\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, April 9, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Palm Sunday, the beginning of Holy Week.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Heb. 9:11-12)}

\begin{latinenglishsection}

 \latinenglish{
	 Fratres: Hoc enim sentíte in vobis, quod et in Christo Jesu:
	 qui, cum in forma Dei esset, non rapínam arbitrátus est esse se æqualem Deo:~\GreDagger\
	 sed semetípsum exinanívit, formam servi accípiens, in similitúdinem hóminum factus,~*
	 et hábitu invéntus ut homo.
 	\Rbar.~Deo gratias.
 }{ 	
	Brethren: Let this mind be in you, which was also in Christ Jesus:
	Who though he was in the form of God, did not regard equality with God as something to be grasped:
	But he emptied himself, taking the form of a servant, being made in the likeness of men, and in habit found as a man. 
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\begin{rubricbox}

All \textbf{kneel} during the verse of this hymn beginning \textit{O Crux Ave}.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../hymns/vexilla-regis-passiontide}

{\itshape

1. Abroad the regal banners fly,
now shines the Cross's mystery:
upon it Life did death endure,
and yet by death did life procure.

2. Who, wounded with a direful spear,
did purposely to wash us clear
from stain of sin, pour out a flood
of precious water mixed with blood. 

3. That which the prophet-king of old
hath in mysterious verse foretold,
is now accomplished, whilst we see
God ruling the nations from a Tree. 

4. O lovely and refulgent Tree,
adorned with purpled majesty;
culled from a worthy stock, to bear
those limbs which sanctified were.

5. Blest Tree, whose happy branches bore
the wealth that did the world restore;
the beam that did that Body weigh
which raised up Hell's expected prey.

6. Hail Cross, of hopes the most sublime!
Now, in the mournful Passion time;
grant to the just increase of grace,
and every sinner's crimes efface. 

7. Blest Trinity, salvation's spring
may every soul Thy praises sing;
to those Thou grantest conquest by
the Holy Cross, rewards supply. Amen. 

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. E(h)ri(h)pe(h) me(h) Dó(h)mi(h)ne(h) ab(h) hó(h)mi(h)ne(h) ma(h)lo.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h) vi(h)ro(h) i(h)ní(h)quo(h) é(h)ri(h)pe(h) me.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~	Deliver me, O Lord, from the evil man.
\Rbar.~Rescue me from the unjust man.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-4-E-A-d}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} O Righteous Father, the world hath not known Thee: but I have known Thee, because Thou has sent me.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens sempitérne Deus, qui humáno generi, ad imitandum humilitátis exémplum,
	Salvatórem nostrum carnem súmere et crucem subíre fecísti:
	concéde propítius; ut et patiéntiæ ipsíus habére documénta,
	et resurrectiónis consórtia mereámur.
	Per Dóminum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty, eternal God, Who, to provide mankind an example of humility for it to imitate,
	willedst that the Saviour should assume our flesh and suffer death upon the Cross,
	mercifully grant that we may be found worthy of the lesson of His endurance and the fellowship of His resurrection.
	Through Our Lord.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Ave Regina Caelorum} follows (no. 188 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}