\documentclass{../vespers-sheet}

\begin{document}

%% TODO: correct date
\chapter*{Vespers Insert for Friday, June 2, 2023}

 \textit{Vespers today of Semidouble rank and is that of the Sixth Feria in the Octave of Pentecost.}

 \section*{Little Chapter (2 Acts 2:1-2)}
 
 \begin{rubricbox}

{\color{red} The Officiant leads the Little Chapter.}

\end{rubricbox}

\begin{latinenglishsection}

 \latinenglish{
	 Cum compleréntur dies Pentecóstes, erant omnes discípuli páriter in eódem loco:~{\color{red}\GreDagger}\ et factus est repénte de cælo sonus, tamquam adveniéntis spíritus veheméntis,~{\color{red}*} et replévit totam domum, ubi erant sedéntes.
	 
 	{\color{red}\Rbar.}~Deo gratias.
 }{
	When the day of Pentecost was fully come, the disciples were all with one accord in one place and suddenly there came a sound from heaven, as of a rushing mighty wind and it filled all the house where they were sitting.
 	{\color{red}\Rbar.}~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/veni-creator-spiritus}


{\itshape

1. Creator-Spirit, all-divine,
Come, visit every soul of thine,
And fill with thy celestial flame
The hearts which thou thyself didst frame.

2. O gift of God, thine is the sweet
Consoling name of Paraclete
And spring of life and fire and love
And unction flowing from above.

3. The mystic sevenfold gifts are thine,
Finger of God's right hand divine;
The Father's promise sent to teach
The tongue a rich and heavenly speech.

4. Kindle with fire brought from above
Each sense, and fill our hearts with love;
And grant our flesh, so weak and frail,
The strength of thine which cannot fail.

5. Drive far away our deadly foe,
And grant us thy true peace to know;
So we, led by thy guidance still,
May safely pass through every ill.

6. To us, through thee, the grace be shown
To know the Father and the Son;
And Spirit of them both, may we
Forever rest our faith in thee.

7. To Sire and Son be praises meet,
And to the Holy Paraclete;
And may Christ send us from above
That Holy Spirit's gift of love.
Amen

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>V/</sp></c>. Lo(h)que(h)bán(h)tur(h) vá(h)ri(h)is(h) lin(h)gu(h)is(h) A(h)pós(h)to(h)li(h), al(h)le(h)lú(h)ja.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>R/</sp></c>. Ut(h) cus(h)tó(h)di(h)ant(h) te(h) in(h) óm(h)ni(h)bus(h) vi(h)is tu(h)is.(g'_/hvGF'E/fgf.) (::)
}

\textit{{\color{red}\Vbar.}~The Apostles spoke in diverse tongues, alleluia.
{\color{red}\Rbar.}~The wonderful works of God, alleluia.}

\vfill
\pagebreak

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

{\color{red} All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} But the Comforter which is the Holy Ghost, Whom the Father will send in My Name, He shall teach you all things, and bring all things to your remembrance, whatsoever I have said unto you. Alleluia.}

\begin{rubricbox}

{\color{red} All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\vfill\pagebreak

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui hodiérna die corda fidélium Sancti Spíritus illustratióne docuísti: da nobis in eódem Spíritu recta sápere; et de ejus semper consolatióne gaudére.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte ejúsdem Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	{\color{red}\Rbar.}~Amen.
}{
	{\color{red}\Vbar.}~Lord, hear my prayer.
	{\color{red}\Rbar.}~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who on this day didst teach the hearts of thy faithful people, by the sending to them the light of thine Holy Spirit, grant us by the same Spirit to have a right judgment in all things, and evermore to rejoice in His holy comfort.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the same Holy Ghost, God, world without end.
	{\color{red}\Rbar.}~Amen.
}

\end{latinenglishsection}

%% TODO: commemorations based on the date

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
}{
	{\color{red}\Vbar.}~Lord, hear my prayer. {\color{red}\Rbar.}~And let my cry come unto Thee.
	{\color{red}\Vbar.}~Let us bless the Lord.{\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-2v-semidouble}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	{\color{red}\Rbar.}~Amen.
}{
	{\color{red}\Vbar.}~May the souls of the faithful departed, through the mercy of God, rest in peace. {\color{red}\Rbar.}~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	{\color{red}\Vbar.}~Dóminus det nobis suam pacem. \\
	{\color{red}\Rbar.}~Et vitam ætérnam. Amen.
}{
	{\color{red}\Vbar.}~May the Lord grant us his peace.{\color{red}\Rbar.}~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The Marian anthem \textit{Salve Regina} follows, as is proper to the season.}

\end{rubricbox}

\end{document}