\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, June 10, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fourteenth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Rom 11:33)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
	 O Altitúdo divitiárum sapiéntiæ, et sciéntiæ Dei,~\GreDagger\
	 quam incomprehensibília sunt ju\textit{dícia} \textbf{e}jus, et investigábiles viæ ejus!
 	\Rbar.~Deo gratias.
 }{
	O the depth of the riches of the wisdom and of the knowledge of God! How incomprehensible are his judgments, and how unsearchable his ways!
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/jam-sol-recedit}

\vfill
\pagebreak

{\itshape

1. As fades the glowing orb of day,
To thee, great source of light, we pray;
Blest Three in One, to every heart
Thy beams of life and love impart.

2. At early dawn, at close of day,
To thee our vows we humbly pay;
May we, mid joys that never end,
With thy bright saints in homage bend.

3. To God the Father, and the Son,
And Holy Spirit, Three in One,
Be endless glory, as before
The world began, so evermore.
Amen.

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ves(h)per(h)tí(h)na(h) o(h)rá(h)ti(h)o(h) a(h)scén(h)dat(h) ad(h) te(h), Dó(h)mi(h)ne.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) de(h)scén(h)dat(h) su(h)per(h) nos(h) mi(h)se(h)ri(h)cór(h)di(h)a(h) tu(h)a.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Let the evening prayer ascend unto thee, O Lord.
\Rbar.~And let there descend upon us thy mercy.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} In all these things Job sinned not with his lips, nor charged God with foolishness.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Custódi, Dómine, qu\'{\ae}sumus, Ecclésiam tuam propitiatióne perpétua: et quia sine te lábitur humána mortálitas; tuis semper auxíliis et abstrahátur a nóxiis et ad salutária dirigátur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Keep, we beseech thee, O Lord, thy Church with thy perpetual mercy, and because the frailty of man without thee cannot but fall, keep us ever by thy help from all things hurtful, and lead us to all things profitable to our salvation.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of St. Peter Claver, Confessor (September 9)}

\textit{\textnormal{Ant.} This man, despising the world, and triumphing over earthly things, hath laid up treasures in heaven by word and deed.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ 
	\Rbar.~Et osténdit illi regnum Dei.  
	
	Orémus.
	Deus, qui abréptos in servitutem Nigritas, ad agnitiónem tui nominis vocaturus, beatum Petrum mira in eis juvándis caritate et patientia roborásti: ejus nobis intercessióne concéde; ut, quæ Jesu Christi sunt quæréntes, proximos opere et veritate diligamus.
}{
	The Lord led the just in right paths. 
	\Rbar.~And showed him the kingdom of God. 
	
	Let us pray.
	O God, Who, when about to call the enslaved negroes unto the knowledge Thy name, didst stremgthen blessed Peter with a wondrous charity and patience for their help: grant, through his intercession, that we, seeking the things of Christ Jesus, may love our neighbors in deeed and in truth.
}

\end{latinenglishsection}

\section*{Commemoration of St. Gorgonius of Rome, Martyr (September 9)}

\textit{\textnormal{Ant.} He that will come after Me, let him deny himself, and take up his cross, and follow Me.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/one-martyr-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justus ut palma florébit.\\
	\Rbar.~Sicut cedrus Líbani sic multiplicábitur.
	
	Orémus.
	Sanctus tuus, Dómine, Gorgónius sua nos intercessióne lætíficet: et pia fáciat sollemnitáte gaudére.
}{
	The just shall flourish like the palm tree.
	\Rbar.~And shall grow up like the cedar of Libanus.
	
	Let us pray.
	May Gorgonius, Your Saint, O Lord, gladden us by his intercession, and cause us to rejoice on his festival which we are devoutly keeping.
}

\end{latinenglishsection}

\section*{Commemoration of St. Nicholas of Tolentino, Confessor (September 10)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.} 

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ 
	\Rbar.~Stolam glóriæ induit eum.
	
	Orémus.
	Adésto, Dómine, supplicatiónibus nostris, quas in beáti Nicolái Confessóris tui solemnitáte deférimus: ut, qui nostræ justítiæ fidúciam non habémus, ejus qui tibi plácuit, précibus adjuvémur.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen. 
}{
	The Lord loved him and adorned him. 
	\Rbar.~He clothed him with a robe of glory. 
	
	Let us pray.
	O Lord, mercifully hear the supplications, which we offer unto thee on this solemn feast day of thy blessed Confessor, Nicolaus, and forasmuch as we put no trust in our own righteousness, grant that we may be helpen by his prayers who walked with thee
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen. 
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-tone-8}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}