\documentclass{../vespers-sheet}

%% reduce the spacing before the title so the hymn can fit all on one page
\titlespacing*{\chapter}{0pt}{35pt}{40pt}

\begin{document}

\chapter*{Vespers Insert for Saturday, 4 December 2021}

 \textit{Vespers today is based on the following Sunday, which is the Second of Advent. All is said as in the main booklet until the Little Chapter.}
 
 \section*{Little Chapter (Romans 15:4)}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Quæcúmque scripta sunt, ad nostram doctrínam scripta sunt:~\GreDagger\
	ut per patiéntiam, et consolatiónem Scripturárum,~*
	spem habeámus.
	\Rbar.~Deo grátias.
}{
	Brethren: What things soever were written, were written for our learning:
	that through patience and the comfort of the scriptures, we might have hope. 
	 \Rbar.~Thanks be to God.
}
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Come, O Lord, visit us in peace, that we may rejoice before Thee with a perfect heart.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Excita Dómine corda nostra ad præparándas Unigéniti tui vias:
	ut per ejus adventum, purificátis tibi méntibus servíre mereámur.
	Qui tecum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Stir up our hearts, O Lord, to prepare the ways of Thine only-begotten Son;
	that through His coming we may attain to serve Thee with purified minds.
	Who liveth.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Comm. of St. Peter Chrysologus, Bishop, Confessor, and Doctor (Dec. 4)}

\textit{\textnormal{Ant.} O blest teacher, light of holy Church, blessed Peter Chrysologus, thou lover of God's law, plead with the Son of God for us.}

\gresetinitiallines{1}
\gregorioscore{comm-peter-chrysologus}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\
	\Rbar.~Et osténdit illi regnum Dei.\\ 
	
	Orémus.
	Deus, qui beátum Petrum Chrysólogum Doctórem egrégrium, divínitus præmonstrátum,
	ad regéndam et instruéndam Ecclésiam tuam éligi voluísti:
	præsta, qu\'{\ae}sumus; ut quem Doctórem vitæ habúimus in terris,
	intercessórem habére mereámur in cælis.
	Per Dóminum.
}{
	The Lord led the just in right paths.
	\Rbar.~And showed him the kingdom of God.
	
	Let us pray.
	O God, who didst elect the blessed and illustrious Doctor, Peter Chrysologus, by a divine direction, to govern and instruct Thy Church;
	grant, we beseech Thee, that we may desere to have him for our intercessor in heaven,
	whom we have had as our teacher of the supernatural life on earth.
	Through Our Lord.
}

\end{latinenglishsection}

%\vfill\pagebreak

\section*{Commemoration of St. Sabbas, Abbot (Dec. 5)}

\textit{\textnormal{Ant.} I will liken him to a wise man that built his house upon a rock.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\
	\Rbar.~Stolam glóriæ induit eum.\\
	
	Orémus.
	Intercéssio nos, qu\'{\ae}sumus Dómine, beáti Sabbæ Abbátis comméndet:
	ut quod nostris méritis non valémus,
	ejus patrocínio assequámur.
	Per Dóminum.
}{
	The Lord loved him and adorned him. %PT: alleluia
	\Rbar.~He clothed him with a robe of glory. %PT: alleluia
	
	Let us pray.
	May the intercession of the blessed abbot Sabbas, we beseeech Thee, O Lord, commend us unto Thee,
	that what we cannot have through our won merits, we may obtain through his patronage.
	Through Our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}