\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, December 3, 2022}

 \textit{Vespers today is based on the following Sunday, which is Second Sunday of Advent.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Romans 15:4)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Quæcúmque scripta sunt, ad nostram doctrínam scripta sunt:~\GreDagger\
	ut per patiéntiam, et consolatiónem Scripturárum,~*
	spem habeámus.
	\Rbar.~Deo grátias.
}{
	Brethren: What things soever were written, were written for our learning:
	that through patience and the comfort of the scriptures, we might have hope. 
	 \Rbar.~Thanks be to God.
}
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-7}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} Come, O Lord, visit us in peace, that we may rejoice before thee with all our heart.}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Excita, Dómine, corda nostra ad præparándas Unigéniti tui vias: ut, per ejus advéntum, purificátis tibi méntibus servíre mereámur:
	Qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Stir up our hearts, O Lord, to make ready the ways of thine Only-begotten Son, that by His coming our minds being purified, we may the more worthily give up ourselves to thy service:
	Who with thee liveth and reigneth, in the unity of the Holy Spirit, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

%% TODO: commemorations based on the date
\textit{\color{red}For commemorations, the Cantor intones the antiphon and says the responsorial prayer afterwards. The Officiant prays the associated collect.}

\section*{Commemoration of St. Bibiana, Virgin Martyr (December 2)}

\textit{\textnormal{Ant.} He that will come after Me, let him deny himself, and take up his cross, and follow Me.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/one-martyr-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justus ut palma florébit.\\
	\Rbar.~Sicut cedrus Líbani sic multiplicábitur.
	
	Orémus.
	Deus, ómnium largítor bonórum, qui in fámula tua Bibiána cum virginitátis flore martýrii palmam conjunxísti: mentes nostras ejus intercessióne tibi caritáte conjúnge; ut, amótis perículis, pr\'{\ae}mia consequámur ætérna.
}{
	The just shall flourish like the palm tree.
	\Rbar.~And shall grow up like the cedar of Libanus.
	
	Let us pray.
	O God, the bestower of all good gifts, who in Thy handmaid Bibiana didst to the flower of virginity join the palm of martyrdom: do Thou, through her intercession, so unite our minds in the bond of charity to Thee, that all dangers being done away, we may obtain everlasting rewards.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of St. Francis Xavier, Priest and Confessor (December 3)}

\textit{\textnormal{Ant.} This man, despising the world, and triumphing over earthly things, hath laid up treasures in heaven by word and deed.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/conf-non-pont-2v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Justum dedúxit Dóminus per vias rectas.\\ 
	\Rbar.~Et osténdit illi regnum Dei. 
	
	Orémus.
	Deus, qui Indiárum gentes beáti Francísci prædicatióne et miráculis Ecclésiæ tuæ aggregáre voluísti: concéde propítius; ut, cujus gloriósa mérita venerámur, virtútum quoque imitémur exémpla. 
}{
	The Lord led the just in right paths. 
	\Rbar.~And showed him the kingdom of God. 
	
	Let us pray.
	O God, Who wast pleased to gather into Thy Church the nations of the Indies by the preaching and miracles of blessed Francis: mercifully grant that we, who venerate his glorious merits, may also imitate the example of his virtues. 
}

\end{latinenglishsection}

\section*{Commemoration of St. Peter Chrysologus, Doctor of the Church (December 3)}

\textit{\textnormal{Ant.} O right excellent Teacher, light of the Holy Church, St. N, Lover of God's law, pray for us to the Son of God.} %PT: alleluia

\gresetinitiallines{1}
\gregorioscore{../commemorations/an--o_doctor_optime--solesmes_1961}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ 
	\Rbar.~Stolam glóriæ índuit eum.
	
	Orémus.
	Deus, qui beátum Petrum Chrysólogum Doctórem egrégium, divínitus præmonstrátum, ad regéndam et instruéndam Ecclésiam tuam éligi voluísti: præsta, qu\'{\ae}sumus; ut, quem Doctórem vitæ habúimus in terris, intercessórem habére mereámur in cælis.
}{
	\Vbar.~The Lord loved him and beautified him.
	\Rbar.~He clothed him with a robe of glory.
	
	Let us pray.
	O God, who didst miraculously point out the glorious doctor Peter Chrysologus, and choose him to be a ruler and teacher of thy Church; grant, we beseech thee, that as on earth he taught us the way of life, so in heaven he may be our continual intercessor with thee.
}

\end{latinenglishsection}

\vfill\pagebreak

\section*{Commemoration of St. Barbara, Virgin Martyr (December 3)}

\textit{\textnormal{Ant.} This Saint fought even unto death for the law of his God, and feared not the words of the wicked; for he was founded upon a firm rock.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/one-martyr-1v}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Glória et honóre coronásti eum, Dómine.\\
	\Rbar.~Et constituísti eum super ópera mánuum tuárum.
	
	Orémus.
	Deus, qui inter cétera poténtiæ tuæ mirácula étiam in sexu frágili victóriam martýrii contulísti: concéde propítius; ut, qui beátæ Bárbaræ Vírginis et Martyris tuæ natalícia cólimus, per ejus ad te exémpla gradiámur. 
	Per Dóminum nostrum Jesum Christum, Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
}{
	Thou hast crowned him with glory and honor, O Lord.
	\Rbar.~And hast set him over the works of Thy hands.
	
	Let us pray.
	O God, Who, among other miracles of Thy power, hast conferred the victory of martyrdom even on the weaker sex, mercifully grant that we, who celebrate the natal feast of blessed Barbara, Thy virgin and martyr, may advance toward Thee through her example. 
	Through our Lord Jesus Christ, Thy Son, Who lives and reigns with Thee in the unity of the Holy Spirit, God, forever and ever
}

\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

%% TODO: benedicamus in line with the season

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

%% TODO: proper Marian anthem for the seasons

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}