\documentclass{../vespers-sheet}
\usepackage{multicol}

\begin{document}

\chapter*{Second Vespers of the Epiphany of Our Lord}

%\section*{Beginning of the Office}

\begin{rubricbox}

{\color{red}When the Officiant kneels, all \textbf{kneel} and pray silently one \textit{Pater noster} (Our Father) and \textit{Ave Maria} (Hail Mary).
Then all make the sign of the cross with the Officiant as he intones:}

\end{rubricbox}

% TODO: Make sure that the tone of the deus adjutorium matches the season primarily and the solemnity of the feast secondarily
 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Praise to Thee, O Lord, King of endless glory.}

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} The Lord our Saviour, begotten before the day-star, and before the ages, is this day made manifest in the world.}
 
 \begin{rubricbox}

{\color{red}All remain standing throughout the first antiphon.
After the psalm is intoned by the Cantor, all \textbf{sit} at the asterisk.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps109-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps109-2} %%

}{
	\input{../psalms/english/ps109} %%
}

\end{latinenglishsection}

 \begin{rubricbox}

{{\color{red} Repeat the antiphon:} Ante luciferum\dots}

\end{rubricbox}

%%

\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} O Jerusalem, thy light is come, and the glory of the Lord is risen upon thee, and the Gentiles shall walk in thy light. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps110-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps110-1} %%

}{
	\input{../psalms/english/ps110} %%
}

\end{latinenglishsection}

%\vfill\pagebreak

 \begin{rubricbox}

{{\color{red} Repeat the antiphon:} Venit lumen tuum\dots}

\end{rubricbox}

%%

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} When the wise men had opened their treasures, they presented unto the Lord gold, frankincense, and myrrh. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon}

%\vfill\pagebreak

\gresetinitiallines{0}
\gregorioscore{ps111-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps111-1} %%

}{
	\input{../psalms/english/ps111} %%
}

\end{latinenglishsection}

 \begin{rubricbox}

{{\color{red} Repeat the antiphon:} Apertis thesauris suis\dots}

\end{rubricbox}
%%

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} O ye seas and floods, bless ye the Lord. O ye wells, bless ye the Lord. Alleluia.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps112-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps112-4-E-A-d} %%

}{
	\input{../psalms/english/ps112} %%
}

\end{latinenglishsection}

 \begin{rubricbox}

{{\color{red} Repeat the antiphon:} Maria et flumina\dots}

\end{rubricbox}

\vfill\pagebreak

%%

\section*{Psalm 113}

\textit{\textnormal{Ant. 5.} This star shineth * as a flame of fire, and pointeth out God, the King of kings: the Wise Men beheld it, and offered gifts unto the great King.}

\gresetinitiallines{1}
\gregorioscore{ps113-antiphon}

\gresetinitiallines{0}
\gregorioscore{ps113-intonation}

 \begin{latinenglishsection}

\latinenglish{

	\input{../psalms/ps113-7} %%

}{
	\input{../psalms/english/ps113} %%
}

\end{latinenglishsection}

%\vfill\pagebreak

 \begin{rubricbox}

{{\color{red} Repeat the antiphon:} Stella ista\dots}

\end{rubricbox}

\section*{Little Chapter (Isaiah 60:1)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	Surge, illumináre Jerúsalem, quia venit \textit{lumen} \textbf{tu}um,~* et glória Dómini super te orta est.
	\Rbar.~Deo grátias.
}{
	Arise, be enlightened, O Jerusalem: for thy light is come, and the glory of the Lord is risen upon thee.
	\Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/hostis-herodes}

{\itshape
1. Why, impious Herod, vainly fear
That Christ the Saviour cometh here?
He takes no earthly realms away
Who gives the crown that lasts for aye.

2. To greet His birth the Wise Men went,
Led by the star before them sent;
Called on by light, towards light they pressed,
And by their gifts their God confessed.

3. In holy Jordan's purest wave
The heavenly Lamb vouchsafed to lave;
That He, to whom was sin unknown,
Might cleanse His people from their own.

4. New miracle of power divine!
The water reddens into wine:
He spake the word: and poured the wave
In other streams than nature gave.

5. All glory, Lord, to thee we pay
For thine Epiphany to-day:
All glory, as is ever meet,
To Father and to Paraclete.
Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Re(h)ges(h) Thar(h)sis(h) et(h) ín(hi)su(h)læ(h_') (,)
mú(h)ne(f)ra(e) óf(f_h)fe(h)rent.(hiH'Ghih.ghG'FE'fggf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Re(h)ges(h) A(h)ra(h)bum(h) et(h) Sa(hi)ba(h_') (,) do(h)na(f) ad(e)dú(f_h)cent.(hiH'Ghih.ghG'FE'fggf.)(::)
}

\textit{\Vbar.~The kings of Tarshish and of the isles shall bring presents.
\Rbar.~The kings of Arabia and Saba shall offer gifts..}

%TODO: Verify that the magnificat antiphon is correct and match the mangificat intonation and pointed text with the tone

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} We celebrate a holy day adorned with three mysteries: this day a star led the Magi the manger; this day wine was made from water at the wedding; this day Christ willed to be baptized in the Jordan by John in order to save us, alleluia.}

\begin{rubricbox}

The Cantor leads by intoning the antiphon; all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-antiphon-only}

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the Cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	\input{../psalms/magnificat-1} %%
}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

 \begin{rubricbox}

Repeat the antiphon: \textit{Tribus miraculis\dots}

\end{rubricbox}

\vfill\pagebreak

\section*{Collect and Conclusion}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Deus, qui hodiérna die Unigénitum tuum géntibus stella duce revelásti: concéde propítius; ut qui jam te ex fide cognóvimus, usque ad contemplándam spéciem tuæ celsitúdinis perducámur.
	Per eúmdem Dóminum nostrum Jesum Christum Fílium tuum, qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	\Vbar. Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O God, Who by the leading of a star didst, as on this day, manifest thine Only-begotten Son to the Gentiles, mercifully grant that we, which know thee now by faith, may after this life have the fruition of thy glorious Godhead:
	Through the same Jesus Christ, thy Son, Our Lord, Who liveth and reigneth with thee in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-2v-solem}

%\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\section*{Marian Anthem}

\gresetinitiallines{1}
\gregorioscore{../marian-anthems/alma-redemptoris-mater-solemn}

{\itshape
O loving Mother of our Redeemer, gate of heaven, star of the sea,
Hasten to aid thy fallen people who strive to rise once more.
Thou who brought forth thy holy Creator, all creation wond'ring, 
Yet remainest ever Virgin, taking from Gabriel's lips
that joyful "Hail!": be merciful to us sinners.
}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Post partum, Virgo, invioláta permansísti.\\
	\Rbar.~Dei Génitrix, intercéde pro nobis
	
	Deus, qui salútis ætérnae, beátæ Maríæ virginitáte f{\oe}cúnda, humáno géneri pr\'{\ae}mia præstitísti:~\GreDagger\ tríbue, qu\'{\ae}sumus; ut ipsam pro nobis intercédere sentiámus, * per quam merúimus auctórem vitæ suscípere, Dóminum nostrum Jesum Christum Filium tuum.\\
	\Rbar.~Amen.
}{
	\Vbar.~After child-birth thou didst remain a pure virgin. \\
	\Rbar.~Intercede for us, O Mother of God.
	
	Let us pray.
	O God, who, by the fruitful virginity of the blessed Mary, hast given to mankind the rewards of eternal  salvation;
	grant, we beseech thee, that we may experience her intercession for us, through whom we have received the author of life,
	Thy Son Jesus Christ, our Lord.
	\Rbar.~Amen.
}

\latinenglish{
	\Vbar.~Divínum auxílium máneat semper nobíscum.\\
	\Rbar.~Amen.
}{
	\Vbar.~May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\begin{rubricbox}

After the Office, all \textbf{kneel} and pray in silence for a time.

\end{rubricbox}

\end{document}