\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, July 2, 2022}

 \textit{Vespers today is based on the following Sunday, which is the Fourth Sunday after Pentecost.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (Sirach 24:14)}

\begin{latinenglishsection}

 \latinenglish{
 	Ab inítio et ante s\'{æ}cula creáta sum,~\GreDagger\
 	et usque ad futúrum s\'{æ}culum \textit{non dési}nam, et in habitatióne sancta coram ipso ministrávi.
	 
 	\Rbar.~Deo grátias.
 }{
	From the beginning, and before the world, was I created, and unto the world to come I shall not cease to be, and in the holy dwelling place I have ministered before him.
 	\Rbar.~Thanks be to God.
 }
 
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/ave_maris_stella}

\vfill
\pagebreak

{\itshape
	1. Hail, star of ocean,
	Child divine who barest,
	Mother, ever-Virgin,
	Heaven's portal fairest.
	
	2. Taking that sweet Ave
	Erst by Gabriel spoken,
	Eva's name reversing,
	Be of peace the token.
	
	3. Break the sinners' fetters,
	Light to blind restoring,
	All our ills dispelling,
	Every boon imploring.
	
	4. Show thyself a mother
	In thy supplication;
	He will hear who chose thee
	At his incarnation.
	
	5. Maid all maids excelling,
	Passing meek and lowly,
	Win for sinners pardon,
	Make us chaste and holy.
	
	6. As we onward journey
	Aid our weak endeavour,
	Till we gaze on Jesus
	And rejoice forever.
	
	7. Father, Son, and Spirit,
	Three in One confessing,
	Give we equal glory
	Equal praise and blessing.
	Amen.
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Be(h)ne(h)dí(h)cta(h) tu(h) in(h) mu(h)li(h)é(h)ri(f)bus(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. Et(h) be(h)ne(h)díc(h)tus(h) fru(h)ctus(h) ven(h)tris(h) tu(h)i(f.). (::)
}

\textit{\Vbar.~Blessed art thou among women..
\Rbar.~And blessed is the Fruit of thy womb.}

\section*{Magnificat}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-8}

\begin{rubricbox}

All \textbf{sit} and sing the antiphon together.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} All generations shall call me blessed, * for God hath regarded the lowliness of His hand-maiden.}

\begin{rubricbox}

All \textbf{stand} for the reading of the collect prayer by the leader.

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Fámulis tuis, qu\'{\ae}sumus, Dómine, cæléstis grátiæ munus impertíre: ut, quibus beátæ Vírginis partus éxstitit salútis exórdium; Visitatiónis ejus votíva solémnitas, pacis tríbuat increméntum.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Grant, O Lord, we beseech thee, unto all thy servants, the gift of thine heavenly grace, that even as the Blessed Virgin being made a Mother hath been unto them the first step unto their salvation, so the godly and solemn memorial of her Visitation may be the bringer of an increase of peace.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of Pope St. Leo, Confessor (June 3)}

\textit{\textnormal{Ant.} While he was supreme Pontiff, he feared not earthly powers, but gloriously went his way to the heavenly kingdom.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/an-dum_esset_summus_pontifex}
%PT: an-dum_esset_summus_pontifex_pt

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Amávit eum Dóminus, et ornávit eum.\\ %PT: allelúia
	\Rbar.~Stolam glóriæ índuit eum. %PT: allelúia
	
	Orémus.
	Gregem tuum, Pastor ætérne, placátus inténde: et per beátum Leónem Summum Pontíficem perpétua protectióne custódi; quem totíus Ecclésiæ præstitísti esse pastórem.
Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia sǽcula sæculórum.
	\Rbar. Amen.
}{
	\Vbar.~The Lord loved him and beautified him. %PT: allelúia
	\Rbar.~He clothed him with a robe of glory. %PT: allelúia
	
	Let us pray.
	Look forgivingly on thy flock, Eternal Shepherd, and keep it in thy constant protection, by the intercession of blessed Leo thy Sovereign Pontiff, whom thou didst constitute Shepherd of the whole Church.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.
	\Rbar. Amen.
}
\end{latinenglishsection}

\section*{Commemoration of the Fourth Sunday after Pentecost}

\textit{\textnormal{Ant.} So David prevailed over the Philistine with a sling, and with a stone, in the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{../commemorations/praevaluit_david}

\begin{latinenglishsection}
\latinenglish{
	\Vbar.~Vespertína orátio ascéndat ad te, Dómine.\\
	\Rbar.~Et descéndat super nos misericórdia tua.
	
	Orémus.
	Da nobis, quǽsumus, Dómine: ut et mundi cursus pacífice nobis tuo órdine dirigátur; et Ecclésia tua tranquílla devotióne lætétur.
}{
	\Vbar.~Let the evening prayer ascend unto thee, O Lord.
	\Rbar.~And let there descend upon us thy mercy.
	
	Let us pray.
	Grant, O Lord, we beseech thee, that the course of this world may be so peaceably ordered by thy governance, that thy Church may joyfully serve thee in all godly quietness.
}
\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-2v-double}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Salve Regina} follows (no. 244 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}