\documentclass{../../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Friday, July 7, 2023}

 \textit{Vespers today is that of Ss. Cyril and Methodius, and is of Double Rite.}

 \section*{Little Chapter (Sirach 44:16-17)}
 
 \textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

 \latinenglish{
 	Ecce sacérdos magnus, qui in diébus suis plácuit Deo, et invéntus est justus: ~{\color{red}\GreDagger}\ et in témpore iracúndiæ factus est reconciliátio.
 	{\color{red}\Rbar.}~Deo grátias.
 }{
	Behold an high priest, who in his days pleased God, and was found righteous, and in the time of wrath he was made a propitiation.
 	{\color{red}\Rbar.}~Thanks be to God.
 }
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../../hymns/sedibus-caeli}

{\itshape
1. Sing, O ye faithful, sing two athlete brothers,
Welcomed above to thrones of light supernal;
Sing, of Slavonia's race, the glory twofold
And strength eternal.

2. One love these brethren bound in sweetest union,
By pity same their solitude is broken;
Forth they would hasten, unto many bearing
Life's blessed token.

3. Soon o'er Bulgaria, Moravia, Bohemia,
Light from the heavenly temple shineth glorious.
Once savage hordes, now countless flocks, to Peter
Lead they victorious.

4. Now with the well-earned crown your brow encircled,
Hear, blessèd ones, to suppliant tears bending;
Unto the Slavs, erst by your gifts enriched,
Protection lending.

5. Oh, may each gen'rous land, your aid imploring,
Keep bright the faith through every generation;
Rome to that land first gave, and ever guardeth,
Life and salvation.

6. Lord, of our race Creator and Redeemer,
By nature good, all goods on us bestowing,
Glory to thee through ages all, from grateful
Hearts overflowing.
Amen.
}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>V/</sp>.</c> Sa(h)cer(h)dó(h)tes(h) tu(h)i(h) in(h)du(h)án(h)tur(h) ju(h)stí(h)ti(f)am.(f.). (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <c><sp>R/</sp>.</c>  Et(h) Sanc(h)ti(h) tu(h)i(h) ex(h)súl(h)tent.(f.). (::)
}

\textit{{\color{red}\Vbar.}~Let thy Priests be clothed with righteousness.
{\color{red}\Rbar.}~And let thy Saints shout for joy.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../../psalms/magnificat-1}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} These are * the holy men become the friends of God and made glorious by the preaching of divine truth; their tongues have become the keys of heaven.}

\vfill\pagebreak

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
	
	Orémus.
	Omnípotens sempitérne Deus, qui Slavóniæ gentes per beátos Confessóres tuos atque Pontífices Cyríllum et Methódium ad agnitiónem tui nóminis veníre tribuísti: præsta; ut, quorum festivitáte gloriámur, eórum consórtio copulémur.
}{
	Lord, hear my prayer.
	{\color{red}\Rbar.}~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and everlasting God, Who hast granted unto the Slavic peoples the knowledge of thy Name through the mean of thy blessed Confessors and Bishops Cyril and Methodius, grant that we, who here keep gladly the festival of the same thy Saints, may hereafter be gathered unto their company.
}

\end{latinenglishsection}

\begin{rubricbox}

{\color{red}The Officiant leads the commemoration in the same manner of the collect:}

\end{rubricbox}

\section*{Commemoration of St. Elizabeth of Hungary, Widow (July 8)}

\textit{\textnormal{Ant.} And now, O ye kings, understand: receive instruction, you that judge the earth.}

\gresetinitiallines{1}
\gregorioscore{../../commemorations/an--et_nunc_reges--solesmes_1961}

\begin{latinenglishsection}
\latinenglish{
	{\color{red}\Vbar.}~Ora pro nobis, beáta Elísabeth.\\
	{\color{red}\Rbar.}~Ut digni efficiámur promissiónibus Christi.
	
	Orémus.
	Clementíssime Deus, qui beátam Elísabeth regínam, inter céteras egrégias dotes, béllici furóris sedándi prærogatíva decorásti: da nobis, ejus intercessióne; post mortális vitæ, quam supplíciter pétimus, pacem, ad ætérna gáudia perveníre.
	Per Dóminum nostrum Jesum Christum, Fílium tuum: qui tecum vivit et regnat in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.\\
	{\color{red}\Rbar.} Amen.
}{
	{\color{red}\Vbar.}~Pray for us, blessed Elizabeth.
	{\color{red}\Rbar.}~That we may be made worthy of the promises of Christ.
	
	Let us pray.
	O most merciful God, Who didst ennoble the blessed Queen Elizabeth with diverse excellent graces, and withal with a gift of stilling the tempest of war, grant unto us through her pleading, in this dying life that peace for which we humbly pray, and hereafter to attain unto everlasting joy in thy Presence.
	Through Jesus Christ, thy Son our Lord, Who liveth and reigneth with thee, in the unity of the Holy Ghost, God, world without end.\\
	{\color{red}\Rbar.} Amen.
}
\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.}~Dómine exáudi oratiónem meam.\\
	{\color{red}\Rbar.}~Et clamor meus ad te véniat.
}{
	{\color{red}\Vbar.}~Lord, hear my prayer. {\color{red}\Rbar.}~And let my cry come unto Thee.
	{\color{red}\Vbar.}~Let us bless the Lord. {\color{red}\Rbar.}~Thanks be to God.
}

\end{latinenglishsection}

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../../common/benedicamus-2v-double}

\vfill\pagebreak

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	{\color{red}\Rbar.} Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. {\color{red}\Rbar.}~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	{\color{red}\Vbar.} Dóminus det nobis suam pacem. \\
	{\color{red}\Rbar.} Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. {\color{red}\Rbar.}~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem as fitting for the season, which in this case, is the Salve Regina. The Officiant leads the concluding collect.}

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../../marian-anthems/salve-regina-simple}

\begin{latinenglishsection}
\latinenglish{
	{\color{red}\Vbar.}~Ora pro nobis, sancta Dei Genitrix.\\
	{\color{red}\Rbar.}~Ut digni efficamur promissionibus Christi.
	
	Orémus.
	Omnipotens sempiterne Deus, qui gloriosae Virginis Matris Mariae corpus et animam, ut dignum Filii tui habitaculum effici mereretur, Spiritu Sancto cooperante, praeparasti, da, ut cuius commemoratione laetamur; eius pia intercessione, ab instantibus malis et a morte perpetua liberemur. Per eundem Christum Dominum nostrum.\\
	{\color{red}\Rbar.}~Amen. 
}{
	{\color{red}\Vbar.}~Pray for us, O holy Mother of God.
	{\color{red}\Rbar.}~That we may be made worthy of the promises of Christ.

	Let us pray. 
	Almighty and everlasting God, Who by the working of the Holy Spirit didst prepare both body and soul of the glorious Virgin Mother, Mary, that she might deserve to be made a worthy dwelling for Thy Son, grant that we who rejoice in her memory, may, by her loving intercession, be delivered from present evils and from lasting death, through the same Christ our Lord. Amen.\\
	{\color{red}\Rbar.}~Amen. 
}
\end{latinenglishsection}

\textit{\color{red}The Officiant leads the following in a low recto tono:}

\begin{latinenglishsection}

\latinenglish{
	{\color{red}\Vbar.} Divínum auxílium máneat semper nobíscum.\\
	{\color{red}\Rbar.} Amen.
}{
	May the divine assistance remain always with us.
	{\color{red}\Rbar.}~Amen.
}
\end{latinenglishsection}

\end{document}