\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{First Vespers of Christ the King\\Insert for 30 Oct. 2021}

\section*{Magnificat}

\textit{\textnormal{Ant Magn.} The Lord God shall give unto him the throne of David His father;
and He shall reign in the house of Jacob for ever, and of His Kingdom there shall be no end, alleluia.
\textnormal{Cant.} My soul doth magnify the Lord: and my spirit hath rejoiced in God my Saviour.
}

\begin{rubricbox}

After the leader intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{magnificat-1v-antiphon-alone}

\vfill
\pagebreak

\begin{rubricbox}

All \textbf{stand} and make the sign of the cross with the cantor.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-1v-cant-intonation}

 \begin{latinenglishsection}

\latinenglish{	
	3. Quia respéxit humilitátem an\textbf{cíl}læ \textbf{su}æ:~*
		ecce enim ex hoc beátam me dicent omnes gene\textit{ra}\textit{ti}\textbf{ó}nes.
	
	4. Quia fecit mihi \textbf{ma}gna qui \textbf{pot}ens est:~*
		et sanctum \textit{no}\textit{men} \textbf{e}jus.
	
	5. Et misericórdia ejus a progénie \textbf{in} pro\textbf{gé}nies~*
		timén\textit{ti}\textit{bus} \textbf{e}um.
	
	6. Fecit poténtiam in \textbf{brá}chio \textbf{su}o:~*
		dispérsit supérbos mente \textit{cor}\textit{dis} \textbf{su}i.
	
	7. Depósuit pot\textbf{én}tes de \textbf{se}de,~*
		et exal\textit{tá}\textit{vit} \textbf{hú}miles.
	
	8. Esuriéntes im\textbf{plé}vit \textbf{bo}nis:~*
		et dívites dimí\textit{sit} \textit{in}\textbf{á}nes.
	
	9. Suscépit Israël \textbf{pú}erum \textbf{su}um,~*
		recordátus misericór\textit{di}\textit{æ} \textbf{su}æ.
	
	10. Sicut locútus est ad \textbf{pa}tres \textbf{nos}tros,~*
		Abraham et sémini e\textit{jus} \textit{in} \textbf{s\'{\ae}}cula.
	
	\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~*
		et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

}{	
	\input{../psalms/english/magnificat}
}

\end{latinenglishsection}

	
\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

All \textbf{sit} and repeat the antiphon (starts on the first page): \textit{Dabit illi\dots},
then \textbf{stand} for the prayer.

\end{rubricbox}

\vfill

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus. Omnípotens sempitérne Deus, qui in dilécto Fílio tuo, universórum Rege, ómnia instauráre voluísti:
	concéde propítius; ut cunctæ famíliæ Géntium, peccáti vúlnere disgregátæ,
	ejus suavíssimo subdántur império. Qui tecum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Almighty and everlasting God, who in Thy beloved Son, King of the whole world, hast willed to restore all things anew;
	grant in Thy mercy that all the families of nations, rent asunder by the wound of sin,
	may be subjected to His most gentle rule. Who with Thee liveth.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\section*{Commemoration of the Sunday}

%% First of November, 23rd after Pentecost

\textit{\textnormal{Ant.} I saw the Lord sitting on a high throne, and the whole earth was full of His glory;
and His train filled the temple.}

\begin{rubricbox}

After the cantor intones up to the asterisk, all \textbf{sit} and join in singing.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{../commemorations/sat-nov-1}

\begin{rubricbox}

All \textbf{stand} for the prayer.

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Vespertína orátio ascéndat ad te Dómine.\\
	\Rbar.~Et descéndat super nos misericórdia tua.
	
	Orémus.
	Absólve, qu\'{\ae}sumus Dómine, tuórum delícta populórum:
	ut a peccatórum néxibus, quæ pro nostra fragilitáte contráximus,
	tua benignitáte liberémur. Per Dominum.
}{
	May our evening prayer ascend to Thee, O Lord.
	\Rbar.~And may Thy mercy descend upon us.
	\pagebreak
	
	Let us pray.
	Absolve, we beseech Thee, O Lord, the sins of Thy people; that we may be delivered by Thy goodness, from the bonds of sin
	which by our frailty we have committed. Through our Lord.
}

\end{latinenglishsection}

\begin{latinenglishsection}

\latinenglish{
	\Rbar.~Amen. \\
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-1v-solem}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\rubric{If no other service follows:}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\textit{The Marian anthem \textnormal{Salve Regina} follows (no. 224 in the \textnormal{Traditional Roman Hymnal}).}

\end{document}