\documentclass{../vespers-booklet}
\usepackage{multicol}

\begin{document}

\chapter*{Vespers of Christ the King}

\section*{Before the Office}

\begin{rubricbox}

When the leader kneels, all \textbf{kneel} and pray silently, making the sign of the cross first over the lips, then over the heart at the marked spots:

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	Aperi, + Dómine, os meum ad benedicéndum nomen sanctum tuum:
	+ munda quoque cor meum ab ómnibus vanis, pervérsis et aliénis cogitatiónibus;
	intelléctum illúmina, afféctum inflámma, ut digne, atténte ac devóte hoc Offícium recitáre váleam,
	et exaudíri mérear ante conspéctum divínæ Majestátis tuæ.
	Per Christum Dóminum nostrum. 
	Amen.
}{
	Open, + O Lord, my mouth to bless Thy holy Name; + cleanse also my heart from all vain, evil, and wandering thoughts; enlighten my understanding and kindle my affections; that I may worthily, attentively, and devoutly say this Office, and so merit to be heard before the presence of Thy divine Majesty.  Through Christ our Lord.  Amen.
}

\latinenglish{
	Domine, in unióne illíus divínæ intentiónis, qua ipse in terris laudes Deo persolvísti, hanc tibi Horam persólvo.
}{
	O Lord, in union with that divine intention wherewith Thou, whilst here on earth, didst render praises unto God, I desire to offer this Hour of prayer unto Thee.
}

\end{latinenglishsection}

%\section*{Beginning of the Office}

\begin{rubricbox}

When the leader stands, all \textbf{stand} and pray silently:

\end{rubricbox}

\begin{latinenglishsection}

\latinenglish{
	Pater Noster\dots
}{
	Our Father\dots
}

\latinenglish{
	Ave María\dots
 }{
 	Hail Mary\dots
 }


 \end{latinenglishsection}

\vfill

\pagebreak

\begin{rubricbox}

All make the sign of the cross with the leader as he intones:

\end{rubricbox}
 
 \gresetinitiallines{1}
\gregorioscore{../common/deus-in-adjutorium-solemn}

\textit{
O God, come to my assistance.
\Vbar.~O Lord, make haste to help me.
Glory be to the Father, and to the Son, and to the Holy Spirit,
as it was in the beginning, is now, and ever shall be, world without end. Amen.
Alleluia.}

\section*{Psalm 109}

\textit{\textnormal{Ant. 1.} He shalled be called the peaceful One, and His throne shall be established for ever.
 \textnormal{Ps.} The Lord said to my Lord: Sit thou at my right hand.}
 
 \begin{rubricbox}

All remain standing throughout the first antiphon.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps109-antiphon-alone}

\vfill
\pagebreak

 \begin{rubricbox}

After the psalm is intoned, all \textbf{sit} at the asterisk.

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{ps109-psalm-intonation}

 \begin{latinenglishsection}

\latinenglish{

2. Donec ponam inimícos \textbf{tu}os,~*
	scabéllum pe\textit{dum} \textit{tu}\textbf{ó}rum.
	
3. Virgam virtútis tuæ emíttet Dóminus ex \textbf{Si}on:~*
	domináre in médio inimicó\textit{rum} \textit{tu}\textbf{ó}rum.
	
4. Tecum princípium in die virtútis tuæ in splendóribus sanc\textbf{tó}rum:~*
	ex útero ante lucíferum \textit{gé}\textit{nu}\textbf{i} te.
	
5. Jurávit Dóminus, et non p{\oe}nitébit \textbf{e}um:~*
	Tu es sacérdos in ætérnum secúndum órdi\textit{nem} \textit{Mel}\textbf{chí}se\-dech.
	
6. Dóminus a dextris \textbf{tu}is,~* 
	confrégit in die iræ \textit{su}\textit{æ} \textbf{re}ges.
	
7. Judicábit in natiónibus, implébit ru\textbf{í}nas:~*
	conquassábit cápita in ter\textit{ra} \textit{mul}\textbf{tó}rum.
	
8. De torrénte in via \textbf{bi}bet:~*
	proptérea exal\textit{tá}\textit{bit} \textbf{ca}put.

}{
	\input{../psalms/english/ps109}
}

\end{latinenglishsection}

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~*
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.
	
\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Pacificus\dots}

\end{rubricbox}

\vfill
\pagebreak

\section*{Psalm 110}

\textit{\textnormal{Ant. 2.} His kingdom is a kingdom of all ages, and all kings shall serve and obey Him.
 \textnormal{Ps.} I will praise thee, O Lord, with my whole heart; in the council of the just: and in the congregation.}
 
 \begin{rubricbox}

All \textbf{stand} as the cantor intones the antiphon, then at the asterisk \textbf{sit} and join in singing.
The same postures are followed for each of the following psalms.

\end{rubricbox}

\gresetinitiallines{1}
\gregorioscore{ps110-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	2. Magna ópera \textbf{Dó}mini:~*
		exquisíta in omnes volun\textit{tá}\textit{tes} \textbf{e}jus.
	
	3. Conféssio et magnificéntia opus \textbf{e}jus:~*
		et justítia ejus manet in s\'{\ae}\textit{cu}\-\textit{lum} \textbf{s\'{\ae}}culi.
	
	4. Memóriam fecit mirabílium suórum,~\GreDagger\
		miséricors et miserátor \textbf{Dó}\-minus:~*
		escam dedit ti\textit{mén}\textit{ti}\textbf{bus} se.
	
	5. Memor erit in s\'{\ae}culum testaménti \textbf{su}i:~*
		virtútem óperum suórum annuntiábit pó\textit{pu}\textit{lo} \textbf{su}o:
		
	\pagebreak
	
	6. Ut det illis hereditátem \textbf{gén}ti\-um:~*
		ópera mánuum ejus véritas, \textit{et} \textit{ju}\textbf{dí}\-cium.
	
	7. Fidélia ómnia mandáta ejus:~\GreDagger\
		confirmáta in s\'{\ae}culum \textbf{s\'{\ae}}culi,~*
		facta in veritáte et \textit{æ}\textit{qui}\textbf{tá}te.
	
	8. Redemptiónem misit pópulo \textbf{su}o:~*
		mandávit in ætérnum testa\textit{mén}\textit{tum} \textbf{su}um.
	
	9. Sanctum, et terríbile nomen \textbf{e}jus:~*
		inítium sapiéntiæ \textit{ti}\textit{mor} \textbf{Dó}\-mini.
	
	10. Intelléctus bonus ómnibus faciéntibus \textbf{e}um:~*
		laudátio ejus manet in s\'{\ae}\textit{cu}\textit{lum} \textbf{s\'{\ae}}culi.
}{	
	 2. Great are the works of the Lord: sought out according to all his wills.
	 
	 3. His work is praise and magnificence: and his justice continueth for ever and ever.
	 
	 4.  He hath made a remembrance of his wonderful works, being a merciful and gracious Lord: he hath given food to them that fear him. 	
	 
	 5. He will be mindful for ever of his covenant: he will shew forth to his people the power of his works.
	 
	 \pagebreak
	 
	 6. That he may give them the inheritance of the Gentiles: the works of his hands are truth and judgment.
	 
	 7. All his commandments are faithful: confirmed for ever and ever, made in truth and equity.
	 
	 8. He hath sent redemption to his people: he hath commanded his covenant for ever.
	 
	 9. Holy and terrible is his name: the fear of the Lord is the beginning of wisdom.
	 
	 10. A good understanding to all that do it: his praise continueth for ever and ever. 
	 
	Glory be.
}

\end{latinenglishsection}

\textit{(bow)} Glória Patri, et \textbf{Fí}lio,~*
	et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Regnum ejus\dots}

\end{rubricbox}

\section*{Psalm 111}

\textit{\textnormal{Ant. 3.} Behold a Man, the Orient is His name; He shall sit and rule, and shall speak peace unto the nations.
 \textnormal{Ps.} Blessed is the man that feareth the Lord: he shall delight exceedingly in his commandments.}

\gresetinitiallines{1}
\gregorioscore{ps111-antiphon-alone}

\vfill
\pagebreak

\gresetinitiallines{0}
\gregorioscore{ps111-psalm-intonation}

 \begin{latinenglishsection}

\latinenglish{
	
	2. Potens in terra erit \textbf{se}men \textbf{e}jus:~*
		generátio rectórum be\textbf{ne}di\textbf{cé}tur.
	
	3. Glória, et divítiæ in \textbf{do}mo \textbf{e}jus:~*
		et justítia ejus manet in \textbf{s\'{\ae}}culum \textbf{s\'{\ae}}culi.
	
	4. Exórtum est in ténebris \textbf{lu}men \textbf{rec}tis:~*
		miséricors, et mise\textbf{rá}tor, et \textbf{jus}tus.
	
	5. Jucúndus homo qui miserétur et cómmodat,~\GreDagger\
		dispónet sermónes suos \textbf{in} ju\textbf{dí}cio:~*
		quia in ætérnum non \textbf{com}mo\textbf{vé}bitur.
	
	6. In memória ætérna \textbf{e}rit \textbf{jus}tus:~*
		ab auditióne mala \textbf{non} ti\textbf{mé}bit.
	
	7. Parátum cor ejus speráre in Dómino,~\GreDagger\
		confirmátum \textbf{est} cor \textbf{e}jus:~*
		non commovébitur donec despíciat ini\textbf{mí}cos \textbf{su}os.
	
	8. Dispérsit, dedit paupéribus:~\GreDagger\
		justítia ejus manet in \textbf{s\'{\ae}}culum \textbf{s\'{\ae}}\-culi,~*
		cornu ejus exaltábi\textbf{tur} in \textbf{gló}ria.
	
	9. Peccátor vidébit, et irascétur,~\GreDagger\
		déntibus suis fremet \textbf{et} ta\textbf{bé}scet:~*
		desidérium pecca\textbf{tó}rum per\textbf{í}bit.
}{
	2. His seed shall be mighty upon earth: the generation of the righteous shall be blessed.
	
	3. Glory and wealth shall be in his house: and his justice remaineth for ever and ever.
	
	4. To the righteous a light is risen up in darkness: he is merciful, and compassionate and just.
	
	5. Acceptable is the man that sheweth mercy and lendeth: he shall order his words with judgment:
	because he shall not be moved for ever.
	
	6. The just shall be in everlasting remembrance: he shall not fear the evil hearing.
	
	7. His heart is ready to hope in the Lord: his heart is strengthened, he shall not be moved until he look over his enemies.
	
	8. He hath distributed, he hath given to the poor: his justice remaineth for ever and ever: his horn shall be exalted in glory.
	
	9. The wicked shall see, and shall be angry, he shall gnash with his teeth and pine away: the desire of the wicked shall perish. 
	
%	Glory be.
}

\end{latinenglishsection}

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}lio,~*
		et Spi\textbf{rí}tui \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textbf{ló}rum. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated (on the previous page): \textit{Ecce vir Oriens\dots}

\end{rubricbox}

\section*{Psalm 112}

\textit{\textnormal{Ant. 4.} The Lord is our Judge, the Lord is our Law-giver; the Lord is our King, he will save us.
 \textnormal{Ps.} Praise the Lord, ye children: praise ye the name of the Lord.}

\gresetinitiallines{1}
\gregorioscore{ps112-antiphon}

 \begin{latinenglishsection}

\latinenglish{
	
2. \textit{(bow)} Sit nomen Dómini \textbf{be}ne\-\textbf{díc}tum,~*
	ex hoc nunc, et \textbf{us}que in \textbf{s\'{\ae}}culum.

3. A solis ortu usque \textbf{ad} oc\textbf{cá}sum,~*
	laudábile \textbf{no}men \textbf{Dó}mini.

4. Excélsus super omnes \textbf{gen}tes \textbf{Dó}\textbf{mi}nus,~*
	et super cælos \textbf{gló}ria \textbf{e}jus.

5. Quis sicut Dóminus, Deus noster, qui in \textbf{al}tis \textbf{há}\textbf{bi}tat,~*
	et humília réspicit in cælo \textbf{et} in \textbf{ter}ra?

6. Súscitans a \textbf{ter}ra \textbf{ín}\textbf{o}pem,~*
	et de stércore \textbf{é}rigens \textbf{páu}perem:

7. Ut cóllocet eum \textbf{cum} prin\textbf{cí}\textbf{pi}bus, *
	cum princípibus \textbf{pó}puli \textbf{su}i.

8. Qui habitáre facit stéri\textbf{lem} in \textbf{do}mo,~*
	matrem fili\textbf{ó}rum læ\textbf{tán}tem.

}{
 	\input{../psalms/english/ps112}
}

\end{latinenglishsection}

\textit{(bow)} Glória \textbf{Pa}tri, et \textbf{Fí}\textbf{li}o,~*
	et Spi\textbf{rí}tui \textbf{Sanc}to.

\textit{(rise)} Sicut erat in princípio, et \textbf{nunc}, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textbf{ló}rum. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated (on the previous page): \textit{Dominus\dots}

\end{rubricbox}

\section*{Psalm 116}

\textit{\textnormal{Ant. 5.} Behold I have given thee for a light of the nations, that thou mayest be My salvation unto the ends of the earth.
 \textnormal{Ps.} O praise the Lord, all ye nations: praise him, all ye people.}

\gresetinitiallines{1}
\gregorioscore{ps116-antiphon}

 \begin{latinenglishsection}

\latinenglish{

	2. Quóniam confirmáta est super nos misericórdia \textbf{e}jus:~*
		et véritas Dómini manet \textit{in} \textit{æ}\textbf{tér}num.
	
	Glória Patri, et \textbf{Fí}lio,~*
		et Spirí\textit{tu}\textit{i} \textbf{Sanc}to.
	
}{
	\input{../psalms/english/ps116}
}

\end{latinenglishsection}

Sicut erat in princípio, et nunc, et \textbf{sem}per,~*
	et in s\'{\ae}cula sæcu\textit{ló}\textit{rum}. \textbf{A}men.

\begin{rubricbox}

The antiphon is repeated: \textit{Ecce dedi te\dots}, then all \textbf{stand}.

\end{rubricbox}

\section*{Little Chapter (Col. 1:12-13)}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Grátias ágimus Deo Patri, qui dignos nos fecit in partem sortis sanctórum in lúmine: \GreDagger\
	qui erípuit nos de potestáte tenebrárum, *
	et tránstulit in regnum Fílii dilectiónis suæ.
	\Rbar.~Deo grátias.
}{
	 Brethren: We give thanks to God the Father, who hath made us worthy to be partakers of the lot of the saints in light:
	 who hath delivered us from the power of darkness, and hath translated us into the kingdom of the Son of his love.
	 \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/te-saeculorum-principem}

\vfill
\pagebreak

{\itshape

1. Thee, Lord of every age, we sing:
Thee, Christ, we hail the nations’ King:
Confess Thy right Thy realm to find
Within the hearts of all mankind.

2. The hate-swayed mob cries, pride-enticed,
They will not have Thy kingdom, Christ;
But we exultant round Thy throne
Thy reign o’er all creation own.

3. O Christ, our Prince, that bringest peace,
Let every rebel impulse cease,
And gather into Thy one fold
The wanderers whom Thy love doth hold.

4. For this Thine arms wide-stretched in plea
Hung bleeding on the atoning Tree;
For this the spear’s revealing dart
Laid bare Thy love-enflamèd Heart.

5. For this Thou didst Thy glory hide,
Outpouring from Thy piercèd side
The riches of Thy love divine
Beneath the veils of bread and wine.

6. May realms and they that rule them vie
With solemn rites to raise Thee high;
May laws and arts Thy servants be,
All life be sanctified in Thee.

7. Their kingly gear and royal state
Kings to their King shall consecrate;
Subjects their all before Thee lay
In service of Thy gentle sway.

8. Jesu, to Thee, beneath whose sway
All earth shall bow, all praise we pay;
With Father and with Spirit be
All glory Thine eternally.

}

\gresetinitiallines{0}
\gabcsnippet{
	(c3) <sp>V/</sp>. Da(h)ta(h) est(h) mi(hi)hi(h_') (,) om(h)nis(f) po(e)tés(f_h)tas.(hiH'Ghih.ghG'FE'fggf.) (::)
}
	
\gresetinitiallines{0}
\gabcsnippet{
	(c3) <sp>R/</sp>. In(h) cæ(hi)lo(h_') (,) et(f') in(e) ter(f_h)ra.(hiH'Ghih.ghG'FE'fggf.) (::)
}

\textit{\Vbar.~All power has been given to Me.
\Rbar.~In heaven and on earth.}

\begin{rubricbox}

The Office continues on the accompanying insert with the \textit{Magnificat}.

\end{rubricbox}

\end{document}