\documentclass{../vespers-sheet}

\begin{document}

\chapter*{Vespers Insert for Saturday, December 17, 2022}

 \textit{Vespers today is based on the following Sunday, which is Fourth Sunday of Advent.
 All is said as given in the main booklet until the Little Chapter.}

 \section*{Little Chapter (1 Cor 4:1-2)}

\textit{\color{red}The Officiant leads the Little Chapter:}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Sic nos exístimet homo ut minístros Christi\GreDagger\, et dispensatóres myster\textit{iórum} \textbf{De}i. Hic jam qu\'{\ae}ritur inter dispensatóres,~* ut fidélis quis inveniátur.
	\Rbar.~Deo grátias.
}{
	Brothers: Let a man so account of us as of the ministers of Christ, and the dispensers of the mysteries of God. Here now it is required among the dispensers, that a man be found faithful.
	 \Rbar.~Thanks be to God.
}
 
\end{latinenglishsection}

\section*{Hymn}

\textit{\color{red}The Cantor leads the hymn:}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\textit{\color{red}The Cantor says the following before all reply afterwards:}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\textit{\color{red}The Cantor intones the Magnificat and leads the first verse:}

\gresetinitiallines{1}
\gregorioscore{magnificat-intonation}

\input{../psalms/magnificat-2}

\begin{rubricbox}

{\color{red}All \textbf{sit} and sing the antiphon together.}

\end{rubricbox}

\gresetinitiallines{0}
\gregorioscore{magnificat-antiphon}

\textit{\textnormal{Ant.} O Wisdom, that comest out of the mouth of the Most High, that reachest from one end to another, and dost mightily and sweetly order all things; come, to teach us the way of prudence!}

\begin{rubricbox}

{\color{red}All \textbf{stand} for the reading of the collect prayer by the Officiant.}

\end{rubricbox}

\section*{Collect}
\textit{\color{red}The Officiant leads the collect:}
\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Aurem tuam, qu\'{\ae}sumus, Dómine, précibus nostris accómmoda: et mentis nostræ ténebras, grátia tuæ visitatiónis illústra:
	Qui vivis et regnas cum Deo Patre, in unitáte Spíritus Sancti, Deus, per ómnia s\'{\ae}cula sæculórum.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	O Lord, we beseech thee, mercifully incline thine ears unto our prayers, and lighten the darkness of our minds by the grace of thy heavenly visitation.
	Who livest and reignest with God the Father in the unity of the Holy Spirit, God, world without end.
	\Rbar.~Amen.
}

\end{latinenglishsection}

\textit{\color{red} The weekday intercessions are omitted.}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\vfill\pagebreak

\textit{\color{red}The Cantor leads the Benedicamus:}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\textit{\color{red}The Officiant leads the following prayers:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

\textit{\color{red}The Cantor leads the Marian anthem:}
\textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\textit{\color{red}The Officiant leads the following:}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Divínum auxílium máneat semper nobíscum.\\
	\Rbar. Amen.
}{
	May the divine assistance remain always with us.
	\Rbar.~Amen.
}
\end{latinenglishsection}

\end{document}