\documentclass{../vespers-sheet}

%% reduce the spacing before the title so the hymn can fit all on one page
\titlespacing*{\chapter}{0pt}{25pt}{40pt}

\begin{document}

\chapter*{Vespers Insert for Saturday, Dec. 18, 2021}

 \textit{Vespers today is based on the following Sunday, which is the Fourth of Advent. All is said as in the main booklet until the Little Chapter.}
 
 \section*{Little Chapter (1 Cor. 4:1-2)}

\begin{latinenglishsection}

\latinenglish{
	
	Fratres: Sic nos exístimet homo ut minístros Christi,~\GreDagger\
	et dispensatóres mysteriórum Dei.~*
	Hic jam qu\'{\ae}ritur inter dispensatóres, ut fidélis quis inveniátur.
	\Rbar.~Deo grátias.
}{
	Brethren: Let a man so account of us as of the ministers of Christ, and the dispensers of the mysteries of God.
	Here now it is required among the dispensers, that a man be found faithful. 
	 \Rbar.~Thanks be to God.
}
\end{latinenglishsection}

\section*{Hymn}

\gresetinitiallines{1}
\gregorioscore{../hymns/conditor-alme-siderum}

\pagebreak

{\itshape

1. Creator of the stars of night,
Thy people's everlasting light,
Jesu, Redeemer, save us all,
and hear Thy servants when they call.

2. Thou, grieving that the ancient curse
should doom to death a universe,
hast found the medicine, full of grace,
to save and heal a ruined race.

3. Thou camest, the Bridegroom of the Bride,
as drew the world to evening tide,
proceeding from a virgin shrine,
the spotless Victim all divine.

4.  At whose dread Name, majestic now,
all knees must bend, all hearts must bow;
and things celestial Thee shall own,
and things terrestrial Lord alone.

5. O Thou whose coming is with dread,
to judge and doom the quick and dead,
preserve us, while we dwell below,
from every insult of the foe.

6. To God the Father, God the Son,
and God the Spirit, Three in One,
laud, honor, might, and glory be
from age to age eternally. Amen. 

}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>V/</sp>. Ro(h)rá(h)te(h) cæ(h)li(h) dé(h)su(h)per,(h) et(h) nu(h)bes(h) plu(h)ant(h) jus(h)tum.(g'_/hvGF'E/fgf.) (::)
}

\gresetinitiallines{0}
\gabcsnippet{
(c3) <sp>R/</sp>. A(h)pe(h)ri(h)á(h)tur(h) ter(h)ra,(h) et(h) gér(h)mi(h)net(h) Sal(h)va(h)tó(h)rem.(g'_/hvGF'E/fgf.) (::)
}

\textit{\Vbar.~Drop down dew, ye heavens, from above, and let the clouds rain the Just One.
\Rbar.~Let the earth be opened, and bud forth the Savior.}

\section*{Magnificat}

\begin{rubricbox}

The antiphon at the \textit{Magnificat} is one of the Great ``O'' antiphons. The antiphon is sung in full both before and after the canticle,
and all \textbf{remain standing} while singing it both times.

\end{rubricbox}

\textit{\textnormal{Ant.} O Adonai, and Leader of the house of Israel, Who didst appear to Moses in the flame of the burning bush and didst give him the law in Sinai: come, and with an outstretched arm, redeem us.}

\gresetinitiallines{1}
\gregorioscore{dec-18-magnificat}

\input{../psalms/magnificat-2}

\begin{rubricbox}

The antiphon is repeated: \textit{O Adonai\dots}

\end{rubricbox}

\section*{Collect}

\begin{latinenglishsection}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
	
	Orémus.
	Excita, qu\'{\ae}sumus, Dómine, poténtiam tuam, et veni, et magna nobis virtúte succúrre,
	ut per auxílium grátiæ tuæ, quod nostra peccáta præpédiunt, indulgéntia tuæ propitiatiónis accéleret.
	Qui vivis.
	\Rbar.~Amen.
}{
	Lord, hear my prayer.
	\Rbar.~And let my cry come unto Thee.
	
	Let us pray.
	Stir up Thy power and come, we pray Thee, O Lord, and with great might succour us;
	that our deliverance, which our sins impede, may be hastened by the help of Thy grace and the forgiveness of Thy mercy.
	Who livest.
	\Rbar.~Amen.
}

\latinenglish{
	\Vbar.~Dómine exáudi oratiónem meam.\\
	\Rbar.~Et clamor meus ad te véniat.
}{
	\Rbar.~Amen. \Vbar.~Lord, hear my prayer. \Rbar.~And let my cry come unto Thee.
	\Vbar.~Let us bless the Lord. \Rbar.~Thanks be to God.
}

\end{latinenglishsection}

\gresetinitiallines{1}
\gregorioscore{../common/benedicamus-advent-lent}

\begin{latinenglishsection}

\latinenglish{
	\Vbar. Fidélium ánimæ, per misericórdiam Dei, requiéscant in pace. \\
	\Rbar. Amen.
}{
	May the souls of the faithful departed, through the mercy of God, rest in peace. \Rbar.~Amen.
}

\latinenglish{
	Pater noster \textit{(silently)}.
}{
	Our Father\dots
}

\latinenglish{
	\Vbar. Dóminus det nobis suam pacem. \\
	\Rbar. Et vitam ætérnam. Amen.
}{
	May the Lord grant us his peace. \Rbar.~And life eternal. Amen.
}

\end{latinenglishsection}

\begin{rubricbox}

The Marian anthem \textit{Alma Redemptoris Mater} follows (no. 181 in the \textit{Traditional Roman Hymnal}).

\end{rubricbox}

\end{document}